---
params:
  year1: 2010
  year2: 2019
fig.caption: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE, 
  message = FALSE, 
  warning = FALSE,
  fig.pos = "h", 
  out.extra = "",
  out.width = '90%',
  fig.align = 'center',
  eval.after = "fig.cap")

pckgs <- c('tidyverse', 'knitr', 'scales','janitor',
           'kableExtra', 'data.table','reactablefmtr',
           'forcats','lubridate', 'anytime','reactable',
           'htmltools', 'plotly','purrr')
# for (pp in `pckgs`) { if (!require(pp)) install.packages(pp); library(pp, character.only = T)  }

if (!require("pacman")) install.packages("pacman")
pacman::p_load(pckgs, character.only = TRUE)
## data input

# add data
dta <- readr::read_csv("data/NSAtleePD.csv") %>%
  setDT() %>%
  .[data.table::between(BrthYear, params$year1, params$year2), ]
```
<!-- # (PART) Deliveries and Births {-} -->

# Deliveries and Births {#section-1}

## Number of deliveries (live births and stillbirths) to residents by year, Nova Scotia, `r paste0(params$year1,"-",params$year2)` {#section-11}

```{r tbl11}

frmt_ns <- dta[, .(BrthYear,
              DLNUMFET)] %>%
  .[,`:=` ( dlv = fifelse( DLNUMFET > 1, "Multiple", "Singleton"))] %>%
  # build 2x2 table
  janitor::tabyl(dlv, BrthYear) %>%
   # add column totals
janitor::adorn_totals(c("row")) %>%
dplyr::mutate_if(is.numeric, scales::comma_format())

  
tbl <- dta[, .(BrthYear,
              DLNUMFET)] %>%
  .[,`:=` ( dlv = fifelse( DLNUMFET > 1, "Multiple", "Singleton"))] %>%
  # build 2x2 table
  janitor::tabyl(dlv, BrthYear) %>%
   # add column totals
janitor::adorn_totals(c("row")) 
#dplyr::mutate_if(is.numeric, scales::comma_format()) %>%
  # formating the percentage value
#janitor::adorn_percentages("col") %>%
#janitor::adorn_pct_formatting(digits = 0) %>%
#janitor::adorn_ns(ns = frmt_ns) %>%
  # add titles
# janitor::adorn_title(row_name = "",
              # col_name = "")

# knitr::kable(tbl) %>%
#   # add an hline before the total
# kableExtra::row_spec(nrow(tbl) - 2, hline_after = TRUE) %>%  kableExtra::row_spec(nrow(tbl), bold = TRUE) %>%
# kableExtra::row_spec(0, bold = TRUE) %>%
# kableExtra::column_spec(1, border_right = TRUE, bold  = TRUE) %>%
# kableExtra::kable_styling(font_size = 12)

## using reactable package for table appearance

reactable(tbl,
          rownames = FALSE,
          resizable = TRUE, 
          onClick = "expand",
          highlight = TRUE, 
          compact = TRUE, 
          defaultColDef = colDef(align = "right",
                                 vAlign = "center",
                                 format = colFormat(locales = "en-CA", separators = TRUE),
                                 style = list(fontSize = 14),
                                 headerStyle = list(fontSize = 14),
                                 minWidth = 50),
          rowStyle = function(index){
            if (index == nrow(tbl)) list(color = "black",
                                         backgroundColor = "#e1e1e1",
                                         fontWeight = "bold", align = "center")
          },
          columns = list(
            dlv = colDef(name = "",
                         align = "left",
                         width = 180)
          )
)
```


## Number of births by year, Nova Scotia, `r paste0(params$year1,"-",params$year2)` {#section-12}

```{r plot11, fig.cap= paste('Number of births by year, Nova Scotia,', paste0(params$year1,"-",params$year2))}

dta1 <- dta[, .(BrthYear,
               DLNUMFET)] %>%
  .[, `:=` (count = .N), 
    by = list(BrthYear)] %>%
  .[, .(BrthYear,
        count)] %>%
  unique() %>%
  arrange(BrthYear) 

plot <- ggplot(data = dta1,
               aes(x = BrthYear,
                   y = count,
                   group = 1,
                   text = paste(BrthYear,
                                "<br>Total births:", scales::comma(count, accuracy = 1)
                                )
                   )
               ) +
  geom_point(color = "#1b9e77",
             show.legend = FALSE) +
  geom_line(color = "#1b9e77",
            show.legend = FALSE,
            alpha = 0.65) +
  # scale_x_date(date_breaks = "2 years",
  #              date_labels = "%Y") +
  scale_x_continuous(#trans = 'log2',
    breaks = pretty) +
  scale_y_continuous(#trans = 'log2',
    breaks = pretty,
    labels = scales::comma_format(accuracy = 1),
    limits = c(0, NA)) +
  labs(y = 'Number of births', x = '') +
  theme_classic()

plotly::ggplotly(plot, tooltip = c("text")) %>%
  plotly::style(hoverlabel = list(
    bgcolor  = "black",
    bordercolor = "transparent",
    align = "left",
    font = list(
      color = "white",
      size = 14,
      face = "bold"
    )
  )) %>%
  plotly::layout(legend = list(orientation = "h", y = -0.25, x = 0.5,
                               xanchor = "center"),
                 font = list(face = "bold", size = 10),
                 showlegend = FALSE)
```

## Number of births by outcome, sex, and year, Nova Scotia, `r paste0(params$year1,"-",params$year2)` {#section-13}

```{r tbl12}
tbl <- dta[, .(BrthYear, 
               BTOUTCOM, 
               BTSEX)] %>%
  .[,`:=` ( dlv = fifelse(!BTOUTCOM %in% "FTD" &
                             BTSEX %in% "M", "Male live births",
                  fifelse(!BTOUTCOM %in% "FTD" &
                             BTSEX %in% "F", "Female live births",
                  fifelse(BTOUTCOM %in% "FTD" &
                             BTSEX %in% "M", "Male stillbirths","Female stillbirths"))),
            dlv1 = fifelse(!BTOUTCOM %in% "FTD", "Total live births", "Total stillbirths"))] %>%
  # remove ambiguous phenotype sex
  .[!tolower(BTSEX) %in% "a"] %>%
  .[, `:=` (BTSEX = factor(BTSEX,
                           levels = c("F","M"),
                           labels = c("Female","Male")))]

  
tbl.f <- tbl[tolower(BTSEX) %in% "female",] %>% 
  janitor::tabyl(dlv, BrthYear) %>%
   # add column totals
janitor::adorn_totals(c("row"), name = "Total Female births")

tbl.m <- tbl[tolower(BTSEX) %in% "male",] %>% 
  janitor::tabyl(dlv, BrthYear) %>%
   # add column totals
janitor::adorn_totals(c("row"), name = "Total Male births")

tbl.t <- tbl %>% 
  janitor::tabyl(dlv1, BrthYear) %>%
   # add column totals
janitor::adorn_totals(c("row"), name = "Total births") %>%
data.table::setnames(old = "dlv1", new = "dlv")

tbl <- rbind(tbl.f, tbl.m, tbl.t) 
    # add titles
# janitor::adorn_title(row_name = "",
#               col_name = "")


# knitr::kable(tbl, 
#              format = "html",
#              booktabs = TRUE) %>%
#   # add an hline before the total
# kableExtra::row_spec(nrow(tbl) - 1, hline_after = TRUE) %>%  kableExtra::row_spec(nrow(tbl), bold = TRUE) %>%
# kableExtra::row_spec(0, bold = TRUE) %>%
# kableExtra::row_spec(nrow(tbl) - 6, bold = TRUE) %>%
# kableExtra::row_spec(nrow(tbl) - 3, bold = TRUE) %>%
# kableExtra::column_spec(1, border_right = TRUE, bold  = TRUE) %>%
# kableExtra::kable_styling(font_size = 12) %>%
# kableExtra::add_footnote(label = c("Sex could not be determined in some infants and these infants are not included in the male or female categories.",
# "Note: Stillbirth refers to the complete expulsion or extraction from its mother after at least 20 weeks pregnancy, or after attaining a weight of 500 g or more, of a fetus in whom, after such expulsion or extraction, there is no breathing, beating of the heart, pulsation of the umbilical cord, or unmistakable movement of voluntary muscle."))

## using reactable package for table appearance

reactable(tbl,
          rownames = FALSE,
          resizable = TRUE, 
          onClick = "expand",
          highlight = TRUE, 
          compact = TRUE, 
          defaultColDef = colDef(align = "right",
                                 vAlign = "center",
                                 format = colFormat(locales = "en-CA", separators = TRUE),
                                 style = list(fontSize = 14),
                                 headerStyle = list(fontSize = 14),
                                 minWidth = 50),
          rowStyle = function(index){
            if (index %in% (seq(1, nrow(tbl), by = 3) + 2)) list(color = "black",
                                         backgroundColor = "#e1e1e1",
                                         fontWeight = "bold", align = "center")
          },
          columns = list(
            dlv = colDef(name = "",
                         align = "left",
                         width = 180)
          )
)%>%
reactablefmtr::add_source("\u1d43 Sex could not be determined in some infants and these infants are not included in the male or female categories.",
                          font_size = 12,
                          #font_weight = "bold"
                          ) %>% 
  reactablefmtr::add_source("Note: Stillbirth refers to the complete expulsion or extraction from its mother after at least 20 weeks pregnancy, or after attaining a weight of 500 g or more, of a fetus in whom, after such expulsion or extraction, there is no breathing, beating of the heart, pulsation of the umbilical cord, or unmistakable movement of voluntary muscle.",
                            font_size = 12,
                            #font_weight = "bold"
                            )
```

## Deliveries resulting from assisted reproductive technology, Nova Scotia, `r paste0(params$year1,"-",params$year2)` {#section-14}

```{r plot12, fig.cap= paste('Deliveries resulting from assisted reproductive technology, Nova Scotia,', paste0(params$year1,"-",params$year2))}

dta1 <- dta[, .(BrthYear,
               DLNUMFET,
               # MZ311,
               # MZ312,
               # MZ313,
               MZ37001,
               MZ37101,
               MZ37201,
               MZ37301,
               MZ37501,
               MZ37611,
               MZ37911
                )] %>%
  .[, `:=` (art  = fifelse(
    MZ37001 >= 1 |
    MZ37101 >= 1 |
    MZ37201 >= 1 |
    MZ37301 >= 1 |
    MZ37501 >= 1 |
    MZ37611 >= 1 |
    MZ37911 >= 1 , 1, 0),
    dlv = fifelse( DLNUMFET >= 1, 1, 0)), 
    by = list(BrthYear)] %>%
  .[, .(BrthYear,
        dlv,
        # MZ311,
        # MZ312,
        # MZ313,
        art)] %>%
  arrange(BrthYear) %>%
  .[, `:=` (dlv = sum(dlv, na.rm = TRUE),
            art = sum(art, na.rm = TRUE)),
    by = list(BrthYear)] %>%
  .[, `:=` (prop = art/dlv)] %>%
  .[,.(BrthYear, prop, dlv)] %>%
  unique()

plot <- ggplot(data = dta1,
               aes(x = BrthYear,
                   y = prop,
                   group = 1,
                   text = paste(BrthYear,
                                "<br>Total births:", scales::comma(dlv, accuracy = 1),
                                "<br>Assisted reproduction (%):", scales::percent(prop, accuracy = 0.1)
                                )
                   )
               ) +
  
  geom_point(color = "#1b9e77",
             show.legend = FALSE) +
  geom_line(color = "#1b9e77",
            show.legend = FALSE,
            alpha = 0.65) +
  # scale_x_date(date_breaks = "2 years",
  #              date_labels = "%Y") +
  scale_x_continuous(#trans = 'log2',
    breaks = pretty) +
  scale_y_continuous(#trans = 'log2',
    breaks = pretty,
    labels = scales::percent_format(accuracy = 0.1),
    limits = c(0, NA)) +
  labs(y = 'Proportion of deliveries', x = '') +
  theme_classic()

plotly::ggplotly(plot, tooltip = c("text")) %>%
  plotly::style(hoverlabel = list(
    bgcolor  = "black",
    bordercolor = "transparent",
    align = "left",
    font = list(
      color = "white",
      size = 14,
      face = "bold"
    )
  )) %>%
  plotly::layout(legend = list(orientation = "h", y = -0.25, x = 0.5,
                               xanchor = "center"),
                 font = list(face = "bold", size = 10),
                 showlegend = FALSE)
```

```{r tbl13}
tbl <- dta[, .(BrthYear,
               DLNUMFET,
               # MZ311,
               # MZ312,
               # MZ313,
               MZ37001,
               MZ37101,
               MZ37201,
               MZ37301,
               MZ37501,
               MZ37611,
               MZ37911
                )] %>%
  .[, `:=` (art  = fifelse(
    MZ37001 >= 1 |
    MZ37101 >= 1 |
    MZ37201 >= 1 |
    MZ37301 >= 1 |
    MZ37501 >= 1 |
    MZ37611 >= 1 |
    MZ37911 >= 1 , 1, 0),
    dlv = fifelse( DLNUMFET >= 1, 1, 0)), 
    by = list(BrthYear)] %>%
  .[, .(BrthYear,
        dlv,
        # MZ311,
        # MZ312,
        # MZ313,
        art)] %>%
  arrange(BrthYear) %>%
  .[, `:=` (dlv = sum(dlv, na.rm = TRUE),
            art = sum(art, na.rm = TRUE)),
    by = list(BrthYear)] %>%
  .[, `:=` (prop = 100*art/dlv)] %>%
  unique() %>%
  melt(measure.vars = c("dlv", "art", "prop")) %>%
  dcast(variable ~ BrthYear)

# coln <- names(tbl)
# coln[1] <- ""

tbl <- as.data.frame(sapply(tbl, sprintf, fmt = c("%.0f", "%.0f", "%1.1f%%")))
#names(tbl) <- coln
tbl[,1] <- c("Total deliveries", "Total ART", "Assisted reproduction (%)")

tbl <- tbl[-2,]

# add comma separator to total count
tbl[1,-1] <- as.character(sapply(as.numeric(tbl[1,-1]), scales::comma_format(accuracy = 1)))

# knitr::kable(tbl, 
#              format = "html",
#              booktabs = TRUE) %>%
#   # add an hline before the total
# kableExtra::row_spec(nrow(tbl) - 1, hline_after = TRUE) %>%  #kableExtra::row_spec(nrow(tbl), bold = TRUE) %>%
# kableExtra::row_spec(0, bold = TRUE) %>%
# kableExtra::column_spec(1, border_right = TRUE, bold  = TRUE) %>%
# kableExtra::kable_styling(font_size = 12) %>%
# kableExtra::add_footnote(label = c("Note: Assisted reproductive technology can include ovulation induction, intracytoplasmic sperm injection (ICSI), embryo transfer, or in vitro fertilization (IVF)."))

reactable(tbl,
          rownames = FALSE,
          resizable = TRUE, 
          onClick = "expand",
          highlight = TRUE, 
          compact = TRUE, 
          # style = list(borderBottom = "#dddddd",
          #              headerStyle = list(borderColor = "#dddddd")),
          defaultColDef = colDef(align = "right",
                                 vAlign = "center",
                                 format = colFormat(locales = "en-CA", separators = TRUE),
                                 style = list(fontSize = 14),
                                 headerStyle = list(fontSize = 14),
                                 minWidth = 50,
                                 maxWidth = 180),
          # rowStyle = function(index){
          #   if (index %in% (seq(1, nrow(tbl), by = 3) + 2)) list(color = "black",
          #                                backgroundColor = "#e1e1e1",
          #                                fontWeight = "bold", align = "center")
          # },
          columns = list(
            variable = colDef(name = "",
                         align = "left",
                         width = 180)
          )
)%>%
reactablefmtr::add_source("Note: Assisted reproductive technology can include ovulation induction, intracytoplasmic sperm injection (ICSI), embryo transfer, or in vitro fertilization (IVF).",
                          font_size = 12,
                          #font_weight = "bold"
                          ) 
```

