---
params:
  year1: 2010
  year2: 2019
fig.caption: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE, 
  message = FALSE, 
  warning = FALSE,
  fig.pos = "h", 
  out.extra = "",
  out.width = "90%",
  fig.align = "center",
  eval.after = "fig.cap")

pckgs <- c("tidyverse", "knitr", "scales","janitor",
           "kableExtra", "formattable","data.table",
           "forcats","lubridate", "anytime","reactable",
           "htmltools", "plotly","purrr")
# for (pp in `pckgs`)00/870 { if (!require(pp)) install.packages(pp); library(pp, character.only = T)  }

if (!require("pacman")) install.packages("pacman")
pacman::p_load(pckgs, character.only = TRUE)
## data input

# add data
dta <- readr::read_csv("data/NSAtleePD.csv") %>%
  setDT() %>%
  .[data.table::between(BrthYear, params$year1, params$year2), ]
```

<!-- # (PART) Deliveries and Births {-} -->

# Labour and Birth Processes {#section-4}

##  Labour induction by parity and year, Nova Scotia, `r paste0(params$year1,"-",params$year2)` {#section-41}

```{r plot41, fig.cap= paste('Labour induction by parity and year, Nova Scotia,', paste0(params$year1,"-",params$year2))}

dta1 <- dta[, .(BrthYear,
               LABOUR,
               FAILED_INDUCTION,
               DLPARA)] %>%
  .[!is.na(DLPARA) |
    !is.na(LABOUR) |
    !is.na(FAILED_INDUCTION)] %>%
  .[, `:=` (ind = fcase(
    tolower(LABOUR) %in% "i" | tolower(FAILED_INDUCTION) %in% "y", 1,
    default = 0
    ),
    parity = fcase(
    DLPARA %in% 0 , 1,
    DLPARA %in% 1, 2,
    default = 3
    ))] %>%
  .[order(BrthYear, parity, ind)] %>%
  #unique() %>%
  .[, `:=` (count = .N), by = list(BrthYear, parity, ind)] %>%
  .[, `:=` (total = .N), by = list(BrthYear, parity)] %>%
  .[, `:=` (rate = count/total)] %>%
  unique()

# age category using UNICODE codes
dta1 <- dta1[, `:=` (ind = factor(ind,
                                   levels = c(0,1),
                                   labels = c("No",
                                              "Induction attempted")),
                     parity = factor(parity,
                                   levels = c(1,2,3),
                                   labels = c("0",
                                              "1",
                                              "\u2265 2")))]

## getting color for plotting

pal <- c("0" = "#1b9e77",
         "1" = "#d95f02",
         "\u2265 2" = "#7570b3")

plot <- ggplot(data = dta1[!ind %in% c("No")],
               aes(x = BrthYear,
                   y = rate,
                   group = parity,
                   colour = parity,
                   fill = parity,
                   shape = parity,
                   text =paste(parity,
                               "<br>Year:", BrthYear,
                               "<br>Total deliveries:", scales::comma(total, accuracy = 1),
                               paste0("<br>",parity," (%):") , scales::percent(rate, accuracy = .1)
                               )
                   )
               ) +
  geom_point(aes(color = parity, fill = parity, shape = parity)) +
  geom_line(aes(color = parity, fill = parity)) +
# Specifying colors by age group
  scale_color_manual(name = "",
                     values = pal,
                     limits = names(pal)) +
  scale_fill_manual(name = "",
                    values = pal,
                    limits = names(pal)) +
  scale_shape_manual(name = "",
                     values = c(21, 22, 23)) +
  # scale_x_date(date_breaks = "2 years",
  #              date_labels = "%Y") +
  scale_x_continuous(#trans = 'log2',
    breaks = pretty) +
  scale_y_continuous(#trans = 'log2',
    breaks = pretty,
    labels = scales::percent_format(accuracy = 1),
    limits = c(0, NA)) +
  labs(y = 'Proportion of deliveries', x = '') +
  theme_classic()

plotly::ggplotly(plot, tooltip = c("text")) %>%
  plotly::style(hoverlabel = list(
    bgcolor  = "black",
    bordercolor = "transparent",
    font = list(
      color = "white",
      size = 14,
      face = "bold"
    )
  )) %>%
  plotly::layout(legend = list(orientation = "h", y = -0.25, x = 0.5,
                               xanchor = "center"),
                 font = list(face = "bold", size = 10),
                 showlegend = TRUE)
```

```{r tbl41}
tbl <- dta[, .(BrthYear,
               LABOUR,
               FAILED_INDUCTION,
               DLPARA)] %>%
  .[!is.na(DLPARA) |
    !is.na(LABOUR) |
    !is.na(FAILED_INDUCTION)] %>%
  .[, `:=` (ind = fcase(
    tolower(LABOUR) %in% "i" | tolower(FAILED_INDUCTION) %in% "y", 1,
    default = 0
    ),
    parity = fcase(
    DLPARA %in% 0 , 1,
    DLPARA %in% 1, 2,
    default = 3
    ))] %>%
  .[order(BrthYear, parity, ind)] %>%
  #unique() %>%
  .[, `:=` (count = .N), by = list(BrthYear, parity, ind)] %>%
  .[, `:=` (total = .N), by = list(BrthYear, parity)] %>%
  .[, `:=` (rate = 100*count/total)] 

tbl1 <- tbl %>%
  .[parity %in% 1 & ind %in% 1] %>%
 melt(measure.vars = c("rate", "total", "count")) %>%
  .[!tolower(variable) %in% "count"] %>%
  .[, `:=` (ind = fifelse(tolower(variable) %in% "total",
                               0, ind))] %>%
  .[, .(BrthYear, ind, value)] %>%
  unique() %>%
  dcast(ind ~ BrthYear, value.var = "value")

# coln <- names(tbl1)
# coln[1] <- ""

tbl1 <- as.data.frame(sapply(tbl1, sprintf, fmt = c("%.0f", "%1.1f%%")))
#names(tbl1) <- coln

tbl1[,1] <- c("Total deliveries", "Induction attempted (%)")

# add comma separator to total count
tbl1[1,-1] <- as.character(sapply(as.numeric(tbl1[1,-1]), scales::comma_format(accuracy = 1)))

temp <- as.data.frame(t(c("0", rep(NA, ncol(tbl1)-1))))
names(temp) <- names(tbl1)

tbl1 <- rbind(temp,tbl1)

tbl2 <- tbl %>%
  .[parity %in% 2 & ind %in% 1] %>%
 melt(measure.vars = c("rate", "total", "count")) %>%
  .[!tolower(variable) %in% "count"] %>%
  .[, `:=` (ind = fifelse(tolower(variable) %in% "total",
                               0, ind))] %>%
  .[, .(BrthYear, ind, value)] %>%
  unique() %>%
  dcast(ind ~ BrthYear, value.var = "value")

# coln <- names(tbl2)
# coln[1] <- ""

tbl2 <- as.data.frame(sapply(tbl2, sprintf, fmt = c("%.0f", "%1.1f%%")))
#names(tbl2) <- coln

tbl2[,1] <- c("Total deliveries", "Induction attempted (%)")

# add comma separator to total count
tbl2[1,-1] <- as.character(sapply(as.numeric(tbl2[1,-1]), scales::comma_format(accuracy = 1)))

temp <- as.data.frame(t(c("1", rep(NA, ncol(tbl2)-1))))
names(temp) <- names(tbl2)

tbl2 <- rbind(temp,tbl2)

tbl3 <- tbl %>%
  .[parity %in% 3 & ind %in% 1] %>%
 melt(measure.vars = c("rate", "total", "count")) %>%
  .[!tolower(variable) %in% "count"] %>%
  .[, `:=` (ind = fifelse(tolower(variable) %in% "total",
                               0, ind))] %>%
  .[, .(BrthYear, ind, value)] %>%
  unique() %>%
  dcast(ind ~ BrthYear, value.var = "value")

# coln <- names(tbl3)
# coln[1] <- ""

tbl3 <- as.data.frame(sapply(tbl3, sprintf, fmt = c("%.0f", "%1.1f%%")))
#names(tbl3) <- coln

tbl3[,1] <- c("Total deliveries", "Induction attempted (%)")

# add comma separator to total count
tbl3[1,-1] <- as.character(sapply(as.numeric(tbl3[1,-1]), scales::comma_format(accuracy = 1)))

temp <- as.data.frame(t(c("\u2265 2", rep(NA, ncol(tbl3)-1))))
names(temp) <- names(tbl3)

tbl3 <- rbind(temp,tbl3)

tbl <- setDT(rbind(tbl1, tbl2, tbl3))
#names(tbl)[1] <- "Parity" 

reactable(tbl,
          rownames = FALSE,
          resizable = TRUE, 
          onClick = "expand",
          highlight = TRUE, 
          compact = TRUE,
          defaultColDef = colDef(align = "right",
                                 vAlign = "center",
                                 format = colFormat(locales = "en-CA", separators = TRUE),
                                 style = list(fontSize = 14),
                                 headerStyle = list(fontSize = 14),
                                 minWidth = 50,
                                 maxWidth = 180),
          rowStyle = function(index){
            if (index %in% (seq(1,nrow(tbl), by = 3))) list(color = "black",
                                         backgroundColor = "#e1e1e1",
                                         fontWeight = "bold",
                                         align = "center",
                                         borderBottom = "#dddddd")
          },
          columns = list(
            ind = colDef(name = "Parity",
                             align = "left",
                             width = 180)
          )
) %>%
  reactablefmtr::add_source("Note: The initiation of contractions in a pregnant woman who is not in labour to help her achieve a vaginal birth within 24 to 48 hours.",
                            font_size = 12,
                            #font_weight = "bold"
  )

# knitr::kable(tbl,
#              format = "html",
#              booktabs = TRUE) %>%
#   # add an hline before the total
# #kableExtra::row_spec(nrow(tbl) - 1, hline_after = TRUE) %>%  #kableExtra::row_spec(nrow(tbl), bold = TRUE) %>%
# kableExtra::row_spec(0, bold = TRUE) %>%
# #kableExtra::column_spec(1, border_right = TRUE, bold  = TRUE) %>%
# kableExtra::group_rows("0", start_row = 1, end_row = 2) %>%
# kableExtra::group_rows("1", start_row = 3, end_row = 4) %>%
# kableExtra::group_rows("\u2265 2", start_row = 5, end_row = 6) %>%
# kableExtra::kable_styling(font_size = 12, full_width = FALSE,
#                           bootstrap_options = c("hover","condensed", "responsive")) %>%
# kableExtra::add_footnote(label = c("Note: The initiation of contractions in a pregnant woman who is not in labour to help her achieve a vaginal birth within 24 to 48 hours."))
```

## Indication for labour induction by year, Nova Scotia, `r paste0(params$year1,"-",params$year2)` {#section-42}

```{r plot42, fig.cap= paste('Indication for labour induction by year, Nova Scotia,', paste0(params$year1,"-",params$year2))}

dta1 <- dta[, .(BrthYear,
                LABOUR,
                FAILED_INDUCTION,
                DMINDUCT,
                DLPARA)] %>%
  .[#!is.na(DLPARA) |
    !is.na(LABOUR) |
    !is.na(FAILED_INDUCTION)] %>%
  .[, `:=` (ind = fcase(
    tolower(LABOUR) %in% "i" | tolower(FAILED_INDUCTION) %in% "y", 1,
    default = 0
    ),
    ind_reason = fcase(
    DMINDUCT %in% 4, 1,
    DMINDUCT %in% 12, 2,
    DMINDUCT %in% 5, 3,
    DMINDUCT %in% c(2, 3, 6, 7, 9, 14, 16, 17, 18, 19, 20, 22, 27, 28), 4,
    default = 5
    ))] %>%
  .[order(BrthYear, ind_reason)] %>%
  .[ind %in% 1] %>%
  #unique() %>%
  .[, `:=` (count = .N), by = list(BrthYear, ind_reason)] %>%
  .[, `:=` (total = .N), by = list(BrthYear)] %>%
  .[, `:=` (rate = count/total)] %>%
  unique()

# age category using UNICODE codes
dta1 <- dta1[, `:=` (ind_reason = factor(ind_reason,
                                   levels = c(1,2,3,4,5),
                                   labels = c("Post dates",
                                              "Hypertensive disease",
                                              "PRoM without chorioamnionitis",
                                              "Other medical reason",
                                              "Other or unspecified")))]

## getting color for plotting

pal <- c("Post dates" = "#1b9e77",
         "Hypertensive disease" = "#d95f02",
         "PRoM without chorioamnionitis" = "#7570b3",
         "Other medical reason" = "#a6761d",
         "Other or unspecified" = "#666666")

plot <- ggplot(data = dta1,
               aes(x = BrthYear,
                   y = rate,
                   group = ind_reason,
                   colour = ind_reason,
                   fill = ind_reason,
                   shape = ind_reason,
                   text =paste(ind_reason,
                               "<br>Year:", BrthYear,
                               "<br>Total deliveries:", scales::comma(total, accuracy = 1),
                               paste0("<br>",ind_reason," (%):") , scales::percent(rate, accuracy = .1)
                               )
                   )
               ) +
  geom_point(aes(color = ind_reason, fill = ind_reason, shape = ind_reason)) +
  geom_line(aes(color = ind_reason, fill = ind_reason)) +
# Specifying colors by age group
  scale_color_manual(name = "",
                     values = pal,
                     limits = names(pal)) +
  scale_fill_manual(name = "",
                    values = pal,
                    limits = names(pal)) +
  scale_shape_manual(name = "",
                     values = c(21,22,23,24,25)) +
  # scale_x_date(date_breaks = "2 years",
  #              date_labels = "%Y") +
  scale_x_continuous(#trans = 'log2',
    breaks = pretty) +
  scale_y_continuous(#trans = 'log2',
    breaks = pretty,
    labels = scales::percent_format(accuracy = 1),
    limits = c(0, NA)) +
  labs(y = 'Proportion of induction attempts', x = '') +
  theme_classic()

plotly::ggplotly(plot, tooltip = c("text")) %>%
  plotly::style(hoverlabel = list(
    bgcolor  = "black",
    bordercolor = "transparent",
    font = list(
      color = "white",
      size = 14,
      face = "bold"
    )
  )) %>%
  plotly::layout(legend = list(orientation = "h", y = -0.25, x = 0.5,
                               xanchor = "center"),
                 font = list(face = "bold", size = 10),
                 showlegend = TRUE)
```

```{r tbl42}
tbl <- dta[, .(BrthYear,
                LABOUR,
                FAILED_INDUCTION,
                DMINDUCT,
                DLPARA)] %>%
  .[#!is.na(DLPARA) |
    !is.na(LABOUR) |
    !is.na(FAILED_INDUCTION)] %>%
  .[, `:=` (ind = fcase(
    tolower(LABOUR) %in% "i" | tolower(FAILED_INDUCTION) %in% "y", 1,
    default = 0
    ),
    ind_reason = fcase(
    DMINDUCT %in% 4, 1,
    DMINDUCT %in% 12, 2,
    DMINDUCT %in% 5, 3,
    DMINDUCT %in% c(2, 3, 6, 7, 9, 14, 16, 17, 18, 19, 20, 22, 27, 28), 4,
    default = 5
    ))] %>%
  .[order(BrthYear, ind_reason)] %>%
  .[ind %in% 1] %>%
  #unique() %>%
  .[, `:=` (count = .N), by = list(BrthYear, ind_reason)] %>%
  .[, `:=` (total = .N), by = list(BrthYear)] %>%
  .[, `:=` (rate = 100*count/total)] %>%
  melt(measure.vars = c("rate", "total", "count")) %>%
  .[!tolower(variable) %in% "count"] %>%
  .[, `:=` (ind_reason = fifelse(tolower(variable) %in% "total",
                               0, ind_reason))] %>%
  .[, .(BrthYear, ind_reason, value)] %>%
  unique() %>%
  dcast(ind_reason ~ BrthYear, value.var = "value")

# coln <- names(tbl)
# coln[1] <- ""

tbl <- as.data.frame(sapply(tbl, sprintf, fmt = c("%.0f", "%1.1f%%","%1.1f%%","%1.1f%%","%1.1f%%","%1.1f%%")))
#names(tbl) <- coln

tbl[,1] <- c("Total induction attempts", "Post dates", "Hypertensive disease", "PRoM without chorioamnionitis", "Other medical reason\u1d47", "Other or unspecified")

# add comma separator to total count
tbl[1,-1] <- as.character(sapply(as.numeric(tbl[1,-1]), scales::comma_format(accuracy = 1)))

reactable(tbl,
          rownames = FALSE,
          resizable = TRUE, 
          onClick = "expand",
          highlight = TRUE, 
          compact = TRUE,
          defaultColDef = colDef(align = "right",
                                 vAlign = "center",
                                 format = colFormat(locales = "en-CA", separators = TRUE),
                                 style = list(fontSize = 14),
                                 headerStyle = list(fontSize = 14),
                                 minWidth = 50,
                                 maxWidth = 180),
          # rowStyle = function(index){
          #   if (index == nrow(tbl)) list(color = "black",
          #                                backgroundColor = "#e1e1e1",
          #                                fontWeight = "bold",
          #                                align = "center",
          #                                borderBottom = "#dddddd")
          # },
          columns = list(
            ind_reason = colDef(name = "",
                             align = "left",
                             width = 180)
          )
) %>%
  reactablefmtr::add_source("\u1d43 PRoM: Prelabour rupture of membranes.",
                            font_size = 12,
                            #font_weight = "bold"
  ) %>%
  reactablefmtr::add_source("\u1d47 Please see glossary under 'Indication for labour induction' for complete list.",
                            font_size = 12,
                            #font_weight = "bold"
  )


# knitr::kable(tbl,
#              format = "html",
#              booktabs = TRUE) %>%
#   # add an hline before the total
# #kableExtra::row_spec(nrow(tbl) - 1, hline_after = TRUE) %>%  #kableExtra::row_spec(nrow(tbl), bold = TRUE) %>%
# kableExtra::row_spec(0, bold = TRUE) %>%
# #kableExtra::column_spec(1, border_right = TRUE, bold  = TRUE) %>%
# kableExtra::group_rows("0", start_row = 1, end_row = 2) %>%
# kableExtra::group_rows("1", start_row = 3, end_row = 4) %>%
# kableExtra::group_rows("\u2265 2", start_row = 5, end_row = 6) %>%
# kableExtra::kable_styling(font_size = 12, full_width = FALSE,
#                           bootstrap_options = c("hover","condensed", "responsive")) %>%
# kableExtra::add_footnote(label = c("PRoM: Prelabour rupture of membranes.",
#                                    "Please see Glossary under 'Indication for labour induction' for complete list."))
```

## Medical augmentation of labour among women with spontaneous onset of labour by year, Nova Scotia, `r paste0(params$year1,"-",params$year2)` {#section-43}

```{r plot43, fig.cap= paste('Medical augmentation of labour among women with spontaneous onset of labour by year, Nova Scotia,', paste0(params$year1,"-",params$year2))}

dta1 <- dta[, .(BrthYear,
                AUGMENT)] %>%
  .[!is.na(AUGMENT)] %>%
  .[, `:=` (aug = fcase(
    tolower(AUGMENT) %in% "y", 1,
    default = 0
    ))] %>%
  #unique() %>%
  .[, `:=` (count = .N), by = list(BrthYear, aug)] %>%
  .[, `:=` (total = .N), by = list(BrthYear)] %>%
  .[, `:=` (rate = count/total)] %>%
  unique()

# age category using UNICODE codes
dta1 <- dta1[, `:=` (aug = factor(aug,
                                   levels = c(0,1),
                                   labels = c("No",
                                              "Augmented")))]

## getting color for plotting

pal <- c("No" = "#1b9e77",
         "Augmented" = "#d95f02")

plot <- ggplot(data = dta1[!aug %in% "No"],
               aes(x = BrthYear,
                   y = rate,
                   group = aug,
                   colour = aug,
                   fill = aug,
                   #shape = aug,
                   text =paste(aug,
                               "<br>Year:", BrthYear,
                               "<br>Total spontaneous labour", scales::comma(total, accuracy = 1),
                               paste0("<br>",aug," (%):") , scales::percent(rate, accuracy = .1)
                               )
                   )
               ) +
  geom_point(aes(color = aug, fill = aug)) +
  geom_line(aes(color = aug, fill = aug)) +
# Specifying colors by age group
  scale_color_manual(name = "",
                     values = pal,
                     limits = names(pal)) +
  scale_fill_manual(name = "",
                    values = pal,
                    limits = names(pal)) +
  # scale_shape_manual(name = "",
  #                    values = c(15,16,17,18,19)) +
  # scale_x_date(date_breaks = "2 years",
  #              date_labels = "%Y") +
  scale_x_continuous(#trans = 'log2',
    breaks = pretty) +
  scale_y_continuous(#trans = 'log2',
    breaks = pretty,
    labels = scales::percent_format(accuracy = 1),
    limits = c(0, NA)) +
  labs(y = 'Proportion of spontaneous labour', x = '') +
  theme_classic()

plotly::ggplotly(plot, tooltip = c("text")) %>%
  plotly::style(hoverlabel = list(
    bgcolor  = "black",
    bordercolor = "transparent",
    font = list(
      color = "white",
      size = 14,
      face = "bold"
    )
  )) %>%
  plotly::layout(legend = list(orientation = "h", y = -0.25, x = 0.5,
                               xanchor = "center"),
                 font = list(face = "bold", size = 10),
                 showlegend = FALSE)
```

```{r tbl43}
tbl <- dta[, .(BrthYear,
                AUGMENT)] %>%
  .[!is.na(AUGMENT)] %>%
  .[, `:=` (aug = fcase(
    tolower(AUGMENT) %in% "y", 1,
    default = 0
    ))] %>%
  #unique() %>%
  .[, `:=` (count = .N), by = list(BrthYear, aug)] %>%
  .[, `:=` (total = .N), by = list(BrthYear)] %>%
  .[, `:=` (rate = 100*count/total)] %>%
  melt(measure.vars = c("rate", "total", "count")) %>%
  .[!tolower(variable) %in% "count" &
      !aug %in% 0] %>%
  .[, `:=` (aug = fifelse(tolower(variable) %in% "total",
                               0, aug))] %>%
  .[, .(BrthYear, aug, value)] %>%
  unique() %>%
  dcast(aug ~ BrthYear, value.var = "value")

# coln <- names(tbl)
# coln[1] <- ""

tbl <- as.data.frame(sapply(tbl, sprintf, fmt = c("%.0f", "%1.1f%%")))
#names(tbl) <- coln

tbl[,1] <- c("Total spontaneous labour", "Augmented")

# add comma separator to total count
tbl[1,-1] <- as.character(sapply(as.numeric(tbl[1,-1]), scales::comma_format(accuracy = 1)))

reactable(tbl,
          rownames = FALSE,
          resizable = TRUE, 
          onClick = "expand",
          highlight = TRUE, 
          compact = TRUE,
          defaultColDef = colDef(align = "right",
                                 vAlign = "center",
                                 format = colFormat(locales = "en-CA", separators = TRUE),
                                 style = list(fontSize = 14),
                                 headerStyle = list(fontSize = 14),
                                 minWidth = 50,
                                 maxWidth = 180),
          # rowStyle = function(index){
          #   if (index == nrow(tbl)) list(color = "black",
          #                                backgroundColor = "#e1e1e1",
          #                                fontWeight = "bold",
          #                                align = "center",
          #                                borderBottom = "#dddddd")
          # },
          columns = list(
            aug = colDef(name = "",
                             align = "left",
                             width = 180)
          )
) %>%
  reactablefmtr::add_source("Note: Use of oxytocin to improve contractions after labour has started spontaneously.",
                            font_size = 12,
                            #font_weight = "bold"
  )

# knitr::kable(tbl,
#              format = "html",
#              booktabs = TRUE) %>%
#   # add an hline before the total
# #kableExtra::row_spec(nrow(tbl) - 1, hline_after = TRUE) %>%  #kableExtra::row_spec(nrow(tbl), bold = TRUE) %>%
# kableExtra::row_spec(0, bold = TRUE) %>%
# #kableExtra::column_spec(1, border_right = TRUE, bold  = TRUE) %>%
# kableExtra::kable_styling(font_size = 12, full_width = FALSE,
#                           bootstrap_options = c("hover","condensed", "responsive")) %>%
# kableExtra::add_footnote(label = c("Note: Use of oxytocin to improve contractions after labour has started spontaneously."))
```

## Use of regional anesthesia with vaginal delivery by year, Nova Scotia, `r paste0(params$year1,"-",params$year2)` {#section-44}

```{r plot44, fig.cap= paste('Use of regional anesthesia with vaginal delivery by year, Nova Scotia,', paste0(params$year1,"-",params$year2))}

dta1 <- dta[, .(BrthYear,
                R010_00200, R010_00300, R010_00400, R010_00500, R010_00600, R010_00800, R010_00900, R010_01000,
                R011_00200, R011_00300, R011_00400, R011_00500, R011_00600, R011_00800, R011_00900, R011_01000,
                R012_00200, R012_00300, R012_00400, R012_00500, R012_00600, R012_00800, R012_00900, R012_01000,
                MEPIS, MEPIC, MIFUS, MPCEA, MSEDN, MSPIN, MPUDL, MGENA,
                DMMETHOD)] %>%
  .[, epi := rowSums(.SD, na.rm = TRUE), .SDcols = c("R010_00200","R010_00300", "R010_00400","R010_00500","R010_01000","R011_00200","R011_00300","R011_00400","R011_00500","R011_01000","R012_00200","R012_00300","R012_00400","R012_00500","R012_01000","MEPIS","MEPIC","MIFUS","MPCEA","MSEDN")] %>%
  .[, spi := rowSums(.SD, na.rm = TRUE), .SDcols = c("R010_00900","R010_01000","R011_00900","R011_01000","R012_00900","R012_01000","MSPIN","MSEDN")] %>%
  .[, dou := rowSums(.SD, na.rm = TRUE), .SDcols = c("R010_01000","R011_01000","R012_01000","MSEDN")] %>%
  .[, pud := rowSums(.SD, na.rm = TRUE), .SDcols = c("R010_00800","R011_00800","R012_00800","MPUDL")] %>%
  .[, gen := rowSums(.SD, na.rm = TRUE), .SDcols = c("R010_00600","R011_00600","R012_00600","MGENA")] %>%
  .[, `:=` (anes = fcase(
    epi >= 1 | spi >= 1 | dou >= 1 | pud >= 1 | gen >= 1, 1,
    default = 0
    ),
    deliv = fcase(
      # spontaneous vaginal
      tolower(DMMETHOD) %in% c("biv", "spt"), 1,
      # assisted vaginal
      tolower(DMMETHOD) %in% c("vac", "low", "mid"), 2,
      # c-section
      tolower(DMMETHOD) %in% c("cst"), 3,
      default = 4
    ))] %>%
  #unique() %>%
  # remove missing
  .[!deliv %in% 4] %>%
  .[, `:=` (count = .N), by = list(BrthYear, anes, deliv)] %>%
  .[, `:=` (total = .N), by = list(BrthYear, deliv)] %>%
  .[, `:=` (rate = count/total)] %>%
  unique()

# age category using UNICODE codes
dta1 <- dta1[, `:=` (anes = factor(anes,
                                   levels = c(0,1),
                                   labels = c("No",
                                              "Yes")),
                     deliv = factor(deliv,
                                    levels = c(1,2,3),
                                    labels = c("Spontaneous vaginal",
                                               "Assisted vaginal",
                                               "Caesarean section")))]

## getting color for plotting

pal <- c("Spontaneous vaginal" = "#1b9e77",
         "Assisted vaginal" = "#d95f02",
         "Caesarean section" = "#7570b3" 
         )

plot <- ggplot(data = dta1[!deliv %in% "Caesarean section" & 
                             !anes %in% "No"],
               aes(x = BrthYear,
                   y = rate,
                   group = deliv,
                   colour = deliv,
                   fill = deliv,
                   shape = deliv,
                   text =paste(deliv,
                               "<br>Year:", BrthYear,
                               "<br>Total deliveries", scales::comma(total, accuracy = 1),
                               paste0("<br>",deliv," (%):") , scales::percent(rate, accuracy = .1)
                               )
                   )
               ) +
  geom_point(aes(color = deliv, fill = deliv, shape = deliv)) +
  geom_line(aes(color = deliv, fill = deliv)) +
# Specifying colors by age group
  scale_color_manual(name = "",
                     values = pal,
                     limits = names(pal)) +
  scale_fill_manual(name = "",
                    values = pal,
                    limits = names(pal)) +
  scale_shape_manual(name = "",
                     values = c(21,22)) +
  # scale_x_date(date_breaks = "2 years",
  #              date_labels = "%Y") +
  scale_x_continuous(#trans = 'log2',
    breaks = pretty) +
  scale_y_continuous(#trans = 'log2',
    breaks = pretty,
    labels = scales::percent_format(accuracy = 1),
    limits = c(0, NA)) +
  labs(y = 'Proportion of vaginal deliveries', x = '') +
  theme_classic()

plotly::ggplotly(plot, tooltip = c("text")) %>%
  plotly::style(hoverlabel = list(
    bgcolor  = "black",
    bordercolor = "transparent",
    font = list(
      color = "white",
      size = 14,
      face = "bold"
    )
  )) %>%
  plotly::layout(legend = list(orientation = "h", y = -0.25, x = 0.5,
                               xanchor = "center"),
                 font = list(face = "bold", size = 10),
                 showlegend = TRUE)
```

```{r tbl44}
tbl <- dta[, .(BrthYear,
                R010_00200, R010_00300, R010_00400, R010_00500, R010_00600, R010_00800, R010_00900, R010_01000,
                R011_00200, R011_00300, R011_00400, R011_00500, R011_00600, R011_00800, R011_00900, R011_01000,
                R012_00200, R012_00300, R012_00400, R012_00500, R012_00600, R012_00800, R012_00900, R012_01000,
                MEPIS, MEPIC, MIFUS, MPCEA, MSEDN, MSPIN, MPUDL, MGENA,
                DMMETHOD)] %>%
  .[, epi := rowSums(.SD, na.rm = TRUE), .SDcols = c("R010_00200","R010_00300", "R010_00400","R010_00500","R010_01000","R011_00200","R011_00300","R011_00400","R011_00500","R011_01000","R012_00200","R012_00300","R012_00400","R012_00500","R012_01000","MEPIS","MEPIC","MIFUS","MPCEA","MSEDN")] %>%
  .[, spi := rowSums(.SD, na.rm = TRUE), .SDcols = c("R010_00900","R010_01000","R011_00900","R011_01000","R012_00900","R012_01000","MSPIN","MSEDN")] %>%
  .[, dou := rowSums(.SD, na.rm = TRUE), .SDcols = c("R010_01000","R011_01000","R012_01000","MSEDN")] %>%
  .[, pud := rowSums(.SD, na.rm = TRUE), .SDcols = c("R010_00800","R011_00800","R012_00800","MPUDL")] %>%
  .[, gen := rowSums(.SD, na.rm = TRUE), .SDcols = c("R010_00600","R011_00600","R012_00600","MGENA")] %>%
  .[, `:=` (anes = fcase(
    epi >= 1 | spi >= 1 | dou >= 1 | pud >= 1 | gen >= 1, 1,
    default = 0
    ),
    deliv = fcase(
      # spontaneous vaginal
      tolower(DMMETHOD) %in% c("biv", "spt"), 1,
      # assisted vaginal
      tolower(DMMETHOD) %in% c("vac", "low", "mid"), 2,
      # c-section
      tolower(DMMETHOD) %in% c("cst"), 3,
      default = 4
    ))] %>%
  #unique() %>%
  # remove missing
  .[!deliv %in% 4] %>%
  .[, `:=` (count = .N), by = list(BrthYear, anes, deliv)] %>%
  .[, `:=` (total = .N), by = list(BrthYear, deliv)] %>%
  .[, `:=` (rate = 100*count/total)]

tbl1 <- tbl %>%
  .[!anes %in% 0 &
    !deliv %in% c(2,3)] %>%
  melt(measure.vars = c("rate", "total", "count")) %>%
  .[!tolower(variable) %in% "count"] %>%
  .[, `:=` (deliv = fifelse(tolower(variable) %in% "total",
                               0, deliv))] %>%
  .[, .(BrthYear, deliv, value)] %>%
  unique() %>%
  dcast(deliv ~ BrthYear, value.var = "value")

# coln <- names(tbl1)
# coln[1] <- ""

tbl1 <- as.data.frame(sapply(tbl1, sprintf, fmt = c("%.0f", "%1.1f%%")))
#names(tbl1) <- coln

tbl1[,1] <- c("Total deliveries", "Anesthesia (%)")

# add comma separator to total count
tbl1[1,-1] <- as.character(sapply(as.numeric(tbl1[1,-1]), scales::comma_format(accuracy = 1)))

temp <- as.data.frame(t(c("Spontaneous", rep(NA, ncol(tbl1)-1))))
names(temp) <- names(tbl1)

tbl1 <- rbind(temp,tbl1)

tbl2 <- tbl %>%
  .[!anes %in% 0 &
    !deliv %in% c(1,3)] %>%
  melt(measure.vars = c("rate", "total", "count")) %>%
  .[!tolower(variable) %in% "count"] %>%
  .[, `:=` (deliv = fifelse(tolower(variable) %in% "total",
                               0, deliv))] %>%
  .[, .(BrthYear, deliv, value)] %>%
  unique() %>%
  dcast(deliv ~ BrthYear, value.var = "value")

# coln <- names(tbl2)
# coln[1] <- ""

tbl2 <- as.data.frame(sapply(tbl2, sprintf, fmt = c("%.0f", "%1.1f%%")))
# names(tbl2) <- coln

tbl2[,1] <- c("Total deliveries", "Anesthesia (%)")
# add comma separator to total count
tbl2[1,-1] <- as.character(sapply(as.numeric(tbl2[1,-1]), scales::comma_format(accuracy = 1)))

temp <- as.data.frame(t(c("Assisted", rep(NA, ncol(tbl1)-1))))
names(temp) <- names(tbl2)

tbl2 <- rbind(temp,tbl2)

tbl <- setDT(rbind(tbl1, tbl2))
#names(tbl)[1] <- "Type of delivery" 

reactable(tbl,
          rownames = FALSE,
          resizable = TRUE, 
          onClick = "expand",
          highlight = TRUE, 
          compact = TRUE,
          defaultColDef = colDef(align = "right",
                                 vAlign = "center",
                                 format = colFormat(locales = "en-CA", separators = TRUE),
                                 style = list(fontSize = 14),
                                 headerStyle = list(fontSize = 14),
                                 minWidth = 50,
                                 maxWidth = 180),
          rowStyle = function(index){
            if (index %in% (seq(1, nrow(tbl), by = 3))) list(color = "black",
                                         backgroundColor = "#e1e1e1",
                                         fontWeight = "bold",
                                         align = "center",
                                         borderBottom = "#dddddd")
          },
          columns = list(
            deliv = colDef(name = "Type of delivery",
                             align = "left",
                             width = 180)
          )
) %>%
  reactablefmtr::add_source("\u1d43 Regional anesthesia including epidural, spinal, and/or pudendal anesthesia during labour and/or delivery.",
                            font_size = 12,
                            #font_weight = "bold"
  )

# knitr::kable(tbl,
#              format = "html",
#              booktabs = TRUE) %>%
#   # add an hline before the total
# #kableExtra::row_spec(nrow(tbl) - 1, hline_after = TRUE) %>%  #kableExtra::row_spec(nrow(tbl), bold = TRUE) %>%
# kableExtra::row_spec(0, bold = TRUE) %>%
# #kableExtra::column_spec(1, border_right = TRUE, bold  = TRUE) %>%
# kableExtra::group_rows("Spontaneous vaginal", start_row = 1, end_row = 2) %>%
# kableExtra::group_rows("Assisted vaginal", start_row = 3, end_row = 4) %>%
# kableExtra::kable_styling(font_size = 12, full_width = FALSE,
#                           bootstrap_options = c("hover","condensed", "responsive")) %>%
# kableExtra::add_footnote(label = c("Regional anesthesia including epidural, spinal, and/or pudendal anesthesia during labour and/or delivery."))
```

## Type of delivery by year, Nova Scotia, `r paste0(params$year1,"-",params$year2)` {#section-45}

```{r plot45, fig.cap= paste('Type of delivery by year, Nova Scotia,', paste0(params$year1,"-",params$year2))}

dta1 <- dta[, .(BrthYear,
                DMMETHOD)] %>%
  .[, `:=` (deliv = fcase(
      # spontaneous vaginal
      tolower(DMMETHOD) %in% c("biv", "spt"), 1,
      # assisted vaginal
      tolower(DMMETHOD) %in% c("vac", "low", "mid"), 2,
      # c-section
      tolower(DMMETHOD) %in% c("cst"), 3,
      default = 4
    ))] %>%
  #unique() %>%
  # remove missing
  .[!deliv %in% 4] %>%
  .[, `:=` (count = .N), by = list(BrthYear, deliv)] %>%
  .[, `:=` (total = .N), by = list(BrthYear)] %>%
  .[, `:=` (rate = count/total)] %>%
  unique()

# age category using UNICODE codes
dta1 <- dta1[, `:=` (deliv = factor(deliv,
                                    levels = c(1,2,3),
                                    labels = c("Spontaneous vaginal",
                                               "Assisted vaginal",
                                               "Caesarean section")))]

## getting color for plotting

pal <- c("Spontaneous vaginal" = "#1b9e77",
         "Assisted vaginal" = "#d95f02",
         "Caesarean section" = "#7570b3" )

plot <- ggplot(data = dta1,
               aes(x = BrthYear,
                   y = rate,
                   group = deliv,
                   colour = deliv,
                   fill = deliv,
                   shape = deliv,
                   text =paste(deliv,
                               "<br>Year:", BrthYear,
                               "<br>Total deliveries:", scales::comma(total, accuracy = 1),
                               paste0("<br>",deliv," (%):"), scales::percent(rate, accuracy = .1)
                               )
                   )
               ) +
  geom_point(aes(color = deliv, fill = deliv, shape = deliv)) +
  geom_line(aes(color = deliv, fill = deliv)) +
# Specifying colors by age group
  scale_color_manual(name = "",
                     values = pal,
                     limits = names(pal)) +
  scale_fill_manual(name = "",
                    values = pal,
                    limits = names(pal)) +
  scale_shape_manual(name = "",
                     values = c(21,22,23)) +
  # scale_x_date(date_breaks = "2 years",
  #              date_labels = "%Y") +
  scale_x_continuous(#trans = 'log2',
    breaks = pretty) +
  scale_y_continuous(#trans = 'log2',
    breaks = pretty,
    labels = scales::percent_format(accuracy = 1),
    limits = c(0, NA)) +
  labs(y = 'Proportion of deliveries', x = '') +
  theme_classic()

plotly::ggplotly(plot, tooltip = c("text")) %>%
  plotly::style(hoverlabel = list(
    bgcolor  = "black",
    bordercolor = "transparent",
    font = list(
      color = "white",
      size = 14,
      face = "bold"
    )
  )) %>%
  plotly::layout(legend = list(orientation = "h", y = -0.25, x = 0.5,
                               xanchor = "center"),
                 font = list(face = "bold", size = 10),
                 showlegend = TRUE)
```

```{r tbl45}
tbl <- dta[, .(BrthYear,
                DMMETHOD)] %>%
  .[, `:=` (deliv = fcase(
      # spontaneous vaginal
      tolower(DMMETHOD) %in% c("biv", "spt"), 1,
      # assisted vaginal
      tolower(DMMETHOD) %in% c("vac", "low", "mid"), 2,
      # c-section
      tolower(DMMETHOD) %in% c("cst"), 3,
      default = 4
    ))] %>%
  #unique() %>%
  # remove missing
  .[!deliv %in% 4] %>%
  .[, `:=` (count = .N), by = list(BrthYear, deliv)] %>%
  .[, `:=` (total = .N), by = list(BrthYear)] %>%
  .[, `:=` (rate = 100*count/total)] %>%
  melt(measure.vars = c("rate", "total", "count")) %>%
  .[!tolower(variable) %in% "count"] %>%
  .[, `:=` (deliv = fifelse(tolower(variable) %in% "total",
                               0, deliv))] %>%
  .[, .(BrthYear, deliv, value)] %>%
  unique() %>%
  dcast(deliv ~ BrthYear, value.var = "value")

# coln <- names(tbl)
# coln[1] <- ""

tbl <- as.data.frame(sapply(tbl, sprintf, fmt = c("%.0f", "%1.1f%%", "%1.1f%%", "%1.1f%%")))
#names(tbl) <- coln

tbl[,1] <- c("Total deliveries", "Spontaneous vaginal", "Assisted vaginal", "Caesarean section")

# add comma separator to total count
tbl[1,-1] <- as.character(sapply(as.numeric(tbl[1,-1]), scales::comma_format(accuracy = 1)))

reactable(tbl,
          rownames = FALSE,
          resizable = TRUE, 
          onClick = "expand",
          highlight = TRUE, 
          compact = TRUE,
          defaultColDef = colDef(align = "right",
                                 vAlign = "center",
                                 format = colFormat(locales = "en-CA", separators = TRUE),
                                 style = list(fontSize = 14),
                                 headerStyle = list(fontSize = 14),
                                 minWidth = 50,
                                 maxWidth = 180),
          # rowStyle = function(index){
          #   if (index == nrow(tbl)) list(color = "black",
          #                                backgroundColor = "#e1e1e1",
          #                                fontWeight = "bold",
          #                                align = "center",
          #                                borderBottom = "#dddddd")
          # },
          columns = list(
            deliv = colDef(name = "",
                             align = "left",
                             width = 180)
          )
) %>%
  reactablefmtr::add_source("\u1d43 With known type of delivery.",
                            font_size = 12,
                            #font_weight = "bold"
  )

# knitr::kable(tbl,
#              format = "html",
#              booktabs = TRUE) %>%
#   # add an hline before the total
# #kableExtra::row_spec(nrow(tbl) - 1, hline_after = TRUE) %>%  #kableExtra::row_spec(nrow(tbl), bold = TRUE) %>%
# kableExtra::row_spec(0, bold = TRUE) %>%
# #kableExtra::column_spec(1, border_right = TRUE, bold  = TRUE) %>%
# kableExtra::group_rows("Spontaneous vaginal", start_row = 1, end_row = 2) %>%
# kableExtra::group_rows("Assisted vaginal", start_row = 3, end_row = 4) %>%
# kableExtra::kable_styling(font_size = 12, full_width = FALSE,
#                           bootstrap_options = c("hover","condensed", "responsive")) %>%
# kableExtra::add_footnote(label = c("With known type of delivery."))
```

## Stage of labour before Caesarean delivery by year, Nova Scotia, `r paste0(params$year1,"-",params$year2)` {#section-46}

```{r plot46, fig.cap= paste('Stage of labour before Caesarean delivery by year, Nova Scotia,', paste0(params$year1,"-",params$year2))}

dta1 <- dta[, .(BrthYear,
                DMCSSTAG)] %>%
  .[, `:=` (cs = fcase(
      # spontaneous vaginal
      tolower(DMCSSTAG) %in% c("cnol"), 1,
      # assisted vaginal
      tolower(DMCSSTAG) %in% c("c1st"), 2,
      # c-section
      tolower(DMCSSTAG) %in% c("c2nd"), 3,
      default = 4
    ))] %>%
  #unique() %>%
  # remove missing
  .[!cs %in% 4] %>%
  .[, `:=` (count = .N), by = list(BrthYear, cs)] %>%
  .[, `:=` (total = .N), by = list(BrthYear)] %>%
  .[, `:=` (rate = count/total)] %>%
  unique()

# age category using UNICODE codes
dta1 <- dta1[, `:=` (cs = factor(cs,
                                    levels = c(1,2,3),
                                    labels = c("No labour",
                                               "1\u02e2\u1d57 stage",
                                               "2\u207f\u1d48 stage")))]

## getting color for plotting

pal <- c("No labour" = "#1b9e77",
         "1\u02e2\u1d57 stage" = "#d95f02",
         "2\u207f\u1d48 stage" = "#7570b3" )

plot <- ggplot(data = dta1,
               aes(x = BrthYear,
                   y = rate,
                   group = cs,
                   colour = cs,
                   fill = cs,
                   shape = cs,
                   text =paste(cs,
                               "<br>Year:", BrthYear,
                               "<br>Total caesarean sections:", scales::comma(total, accuracy = 1),
                               paste0("<br>", cs, " (%):") , scales::percent(rate, accuracy = .1)
                               )
                   )
               ) +
  geom_point(aes(color = cs, fill = cs, shape = cs)) +
  geom_line(aes(color = cs, fill = cs)) +
# Specifying colors by age group
  scale_color_manual(name = "",
                     values = pal,
                     limits = names(pal)) +
  scale_fill_manual(name = "",
                    values = pal,
                    limits = names(pal)) +
  scale_shape_manual(name = "",
                     values = c(21,22,23)) +
  # scale_x_date(date_breaks = "2 years",
  #              date_labels = "%Y") +
  scale_x_continuous(#trans = 'log2',
    breaks = pretty) +
  scale_y_continuous(#trans = 'log2',
    breaks = pretty,
    labels = scales::percent_format(accuracy = 1),
    limits = c(0, NA)) +
  labs(y = 'Proportion of caesarean sections', x = '') +
  theme_classic()

plotly::ggplotly(plot, tooltip = c("text")) %>%
  plotly::style(hoverlabel = list(
    bgcolor  = "black",
    bordercolor = "transparent",
    font = list(
      color = "white",
      size = 14,
      face = "bold"
    )
  )) %>%
  plotly::layout(legend = list(orientation = "h", y = -0.25, x = 0.5,
                               xanchor = "center"),
                 font = list(face = "bold", size = 10),
                 showlegend = TRUE)
```

```{r tbl46}
tbl <- dta[, .(BrthYear,
                DMCSSTAG)] %>%
  .[, `:=` (cs = fcase(
      # spontaneous vaginal
      tolower(DMCSSTAG) %in% c("cnol"), 1,
      # assisted vaginal
      tolower(DMCSSTAG) %in% c("c1st"), 2,
      # c-section
      tolower(DMCSSTAG) %in% c("c2nd"), 3,
      default = 4
    ))] %>%
  #unique() %>%
  # remove missing
  .[!cs %in% 4] %>%
  .[, `:=` (count = .N), by = list(BrthYear, cs)] %>%
  .[, `:=` (total = .N), by = list(BrthYear)] %>%
  .[, `:=` (rate = 100*count/total)] %>%
  melt(measure.vars = c("rate", "total", "count")) %>%
  .[!tolower(variable) %in% "count"] %>%
  .[, `:=` (cs = fifelse(tolower(variable) %in% "total",
                               0, cs))] %>%
  .[, .(BrthYear, cs, value)] %>%
  unique() %>%
  dcast(cs ~ BrthYear, value.var = "value")

# coln <- names(tbl)
# coln[1] <- ""

tbl <- as.data.frame(sapply(tbl, sprintf, fmt = c("%.0f", "%1.1f%%", "%1.1f%%", "%1.1f%%")))
#names(tbl) <- coln

tbl[,1] <- c("Total caesarean sections", "No labour", "1\u02e2\u1d57 stage", "2\u207f\u1d48 stage")

# add comma separator to total count
tbl[1,-1] <- as.character(sapply(as.numeric(tbl[1,-1]), scales::comma_format(accuracy = 1)))

reactable(tbl,
          rownames = FALSE,
          resizable = TRUE, 
          onClick = "expand",
          highlight = TRUE, 
          compact = TRUE,
          defaultColDef = colDef(align = "right",
                                 vAlign = "center",
                                 format = colFormat(locales = "en-CA", separators = TRUE),
                                 style = list(fontSize = 14),
                                 headerStyle = list(fontSize = 14),
                                 minWidth = 50,
                                 maxWidth = 180),
          # rowStyle = function(index){
          #   if (index == nrow(tbl)) list(color = "black",
          #                                backgroundColor = "#e1e1e1",
          #                                fontWeight = "bold",
          #                                align = "center",
          #                                borderBottom = "#dddddd")
          # },
          columns = list(
            cs = colDef(name = "",
                             align = "left",
                             width = 180)
          )
) %>%
  reactablefmtr::add_source("The 1\u02e2\u1d57 stage is the period from the onset of labour until the cervix is fully dilated (10 cm). The 2\u207f\u1d48 stage is the period from 10 cm dilation of the cervix until the baby is delivered.",
                            font_size = 12,
                            #font_weight = "bold"
  )

# knitr::kable(tbl,
#              format = "html",
#              booktabs = TRUE) %>%
#   # add an hline before the total
# #kableExtra::row_spec(nrow(tbl) - 1, hline_after = TRUE) %>%  #kableExtra::row_spec(nrow(tbl), bold = TRUE) %>%
# kableExtra::row_spec(0, bold = TRUE) %>%
# #kableExtra::column_spec(1, border_right = TRUE, bold  = TRUE) %>%
# kableExtra::kable_styling(font_size = 12, full_width = FALSE,
#                           bootstrap_options = c("hover","condensed", "responsive")) %>%
# kableExtra::add_footnote(label = c("The 1$^{st}$ stage is the period from the onset of labour until the cervix is fully dilated (10 cm). The 2$^{nd}$ stage is the period from 10 cm dilation of the cervix until the baby is delivered."),
# escape = FALSE)
```

## Primary indication for Caesarean delivery by year, Nova Scotia, `r paste0(params$year1,"-",params$year2)` {#section-47}

```{r plot47, fig.cap= paste('Primary indication for Caesarean delivery by year, Nova Scotia,', paste0(params$year1,"-",params$year2))}

dta1 <- dta[, .(BrthYear,
                INDICCS1)] %>%
  .[, `:=` (cs = fcase(
      toupper(INDICCS1) %in% c("PCS"), 1,
      toupper(INDICCS1) %in% c("BCH","MLP","TLI"), 2,
      toupper(INDICCS1) %in% c("DYS"), 3,
      toupper(INDICCS1) %in% c("FDS"), 4,
      toupper(INDICCS1) %in% c("CXD","DBT","HTD","HIV","HSV","ISO","PTD","UTS"), 5,
      toupper(INDICCS1) %in% c("FGT","MAC","SFA","MTP","OFC"), 6,
      toupper(INDICCS1) %in% c("APL","PLP","PLC"), 7,
      toupper(INDICCS1) %in% c("AMA","MAT","PRM","FID","OOC","SUR","OTR","PMC",""), 8,
      default = 9
    ))] %>%
  #unique() %>%
  # remove missing
  .[!cs %in% 9] %>%
  .[, `:=` (count = .N), by = list(BrthYear, cs)] %>%
  .[, `:=` (total = .N), by = list(BrthYear)] %>%
  .[, `:=` (rate = count/total)] %>%
  unique()

# age category using UNICODE codes
dta1 <- dta1[, `:=` (cs = factor(cs,
                                    levels = c(1,2,3,4,5,6,7,8),
                                    labels = c("Previous caesarean",
                                               "Breech or other malpresentation",
                                               "Labour dystocia",
                                               "Fetal distress",
                                               "Maternal conditions",
                                               "Fetal conditions",
                                               "Placenta or cord related",
                                               "Other")))]

## getting color for plotting

pal <- c("Previous caesarean" = "#1b9e77",
        "Breech or other malpresentation" = "#d95f02",
        "Labour dystocia" = "#7570b3",
        "Fetal distress" = "#e7298a",
        "Maternal conditions" = "#66a61e",
        "Fetal conditions" = "#e6ab02",
        "Placenta or cord related" = "#a6761d",
        "Other" = "#666666")

plot <- ggplot(data = dta1,
               aes(x = BrthYear,
                   y = rate,
                   group = cs,
                   colour = cs,
                   fill = cs,
                   shape = cs,
                   text = paste(cs,
                               "<br>Year:", BrthYear,
                               "<br>Total caesarean sections:", scales::comma(total, accuracy = 1),
                               paste0("<br>%",cs, " (%):") , scales::percent(rate, accuracy = .1)
                               )
                   )
               ) +
  geom_point(aes(color = cs, fill = cs, shape = cs)) +
  geom_line(aes(color = cs, fill = cs)) +
# Specifying colors by age group
  scale_color_manual(name = "",
                     values = pal,
                     limits = names(pal)) +
  scale_fill_manual(name = "",
                    values = pal,
                    limits = names(pal)) +
  scale_shape_manual(name = "",
                     values = c(21,22,23,24,25,9,10,11)) +
  # scale_x_date(date_breaks = "2 years",
  #              date_labels = "%Y") +
  scale_x_continuous(#trans = 'log2',
    breaks = pretty) +
  scale_y_continuous(#trans = 'log2',
    breaks = pretty,
    labels = scales::percent_format(accuracy = 1),
    limits = c(0, NA)) +
  labs(y = 'Proportion of caesarean sections', x = '') +
  theme_classic()

plotly::ggplotly(plot, tooltip = c("text")) %>%
  plotly::style(hoverlabel = list(
    bgcolor  = "black",
    bordercolor = "transparent",
    font = list(
      color = "white",
      size = 14,
      face = "bold"
    )
  )) %>%
  plotly::layout(legend = list(orientation = "h", y = -0.25, x = 0.5,
                               xanchor = "center"),
                 font = list(face = "bold", size = 10),
                 showlegend = TRUE)
```

```{r tbl47}
tbl <- dta[, .(BrthYear,
                INDICCS1)] %>%
  .[, `:=` (cs = fcase(
      toupper(INDICCS1) %in% c("PCS"), 1,
      toupper(INDICCS1) %in% c("BCH","MLP","TLI"), 2,
      toupper(INDICCS1) %in% c("DYS"), 3,
      toupper(INDICCS1) %in% c("FDS"), 4,
      toupper(INDICCS1) %in% c("CXD","DBT","HTD","HIV","HSV","ISO","PTD","UTS"), 5,
      toupper(INDICCS1) %in% c("FGT","MAC","SFA","MTP","OFC"), 6,
      toupper(INDICCS1) %in% c("APL","PLP","PLC"), 7,
      toupper(INDICCS1) %in% c("AMA","MAT","PRM","FID","OOC","SUR","OTR","PMC",""), 8,
      default = 9
    ))] %>%
  #unique() %>%
  # remove missing
  .[!cs %in% 9] %>%
  .[, `:=` (count = .N), by = list(BrthYear, cs)] %>%
  .[, `:=` (total = .N), by = list(BrthYear)] %>%
  .[, `:=` (rate = 100*count/total)] %>%
  melt(measure.vars = c("rate", "total", "count")) %>%
  .[!tolower(variable) %in% "count"] %>%
  .[, `:=` (cs = fifelse(tolower(variable) %in% "total",
                               0, cs))] %>%
  .[, .(BrthYear, cs, value)] %>%
  unique() %>%
  dcast(cs ~ BrthYear, value.var = "value")

# coln <- names(tbl)
# coln[1] <- ""

tbl <- as.data.frame(sapply(tbl, sprintf, fmt = c("%.0f", "%1.1f%%", "%1.1f%%", "%1.1f%%","%1.1f%%","%1.1f%%","%1.1f%%","%1.1f%%","%1.1f%%")))
#names(tbl) <- coln

tbl[,1] <- c("Total caesarean sections", "Previous caesarean", "Breech or other malpresentation", "Labour dystocia", "Fetal distress", "Maternal conditions", "Fetal conditions", "Placenta or cord related", "Other")

# add comma separator to total count
tbl[1,-1] <- as.character(sapply(as.numeric(tbl[1,-1]), scales::comma_format(accuracy = 1)))

reactable(tbl,
          rownames = FALSE,
          resizable = TRUE, 
          onClick = "expand",
          highlight = TRUE, 
          compact = TRUE,
          defaultColDef = colDef(align = "right",
                                 vAlign = "center",
                                 format = colFormat(locales = "en-CA", separators = TRUE),
                                 style = list(fontSize = 14),
                                 headerStyle = list(fontSize = 14),
                                 minWidth = 50,
                                 maxWidth = 180),
          # rowStyle = function(index){
          #   if (index == nrow(tbl)) list(color = "black",
          #                                backgroundColor = "#e1e1e1",
          #                                fontWeight = "bold",
          #                                align = "center",
          #                                borderBottom = "#dddddd")
          # },
          columns = list(
            cs = colDef(name = "",
                             align = "left",
                             width = 180)
          )
)

# knitr::kable(tbl,
#              format = "html",
#              booktabs = TRUE) %>%
#   # add an hline before the total
# #kableExtra::row_spec(nrow(tbl) - 1, hline_after = TRUE) %>%  #kableExtra::row_spec(nrow(tbl), bold = TRUE) %>%
# kableExtra::row_spec(0, bold = TRUE) %>%
# #kableExtra::column_spec(1, border_right = TRUE, bold  = TRUE) %>%
# kableExtra::kable_styling(font_size = 12, full_width = FALSE,
#                           bootstrap_options = c("hover","condensed", "responsive"))
```

## Caesarean delivery by Robson group and year, Nova Scotia, `r paste0(params$year1,"-",params$year2)` {#section-48}

```{r plot48, fig.cap= paste('Caesarean delivery by Robson group and year, Nova Scotia,', paste0(params$year1,"-",params$year2))}

dta1 <- dta[, .(BrthYear,
                DLNUMFET,
                GA_BEST,
                POSATDEL,
                DLPARA,
                LABOUR,
                DLPRVCS,
                DMMETHOD)] %>%
  .[, `:=` (robsongrp = fcase(
  DLNUMFET %in% 1 & GA_BEST >= 37 & toupper(POSATDEL) %in% c('VTX','POP','BOW','CPD','FAC') & DLPARA %in% 0 & tolower(LABOUR) %in% "s", 1,
  DLNUMFET %in% 1 & GA_BEST >= 37 & toupper(POSATDEL) %in% c('VTX','POP','BOW','CPD','FAC') & DLPARA %in% 0 & tolower(LABOUR) %in% "i", 2.1,
  DLNUMFET %in% 1 & GA_BEST >= 37 & toupper(POSATDEL) %in% c('VTX','POP','BOW','CPD','FAC') & DLPARA %in% 0 & tolower(LABOUR) %in% "n", 2.2,
  DLNUMFET %in% 1 & DLPARA > 0 & DLPRVCS %in% 0 & tolower(LABOUR) %in% "s", 3,
  DLNUMFET %in% 1 & DLPARA > 0 & DLPRVCS %in% 0 & tolower(LABOUR) %in% "i", 4.1,
  DLNUMFET %in% 1 & DLPARA > 0 & DLPRVCS %in% 0 & tolower(LABOUR) %in% "n", 4.2,
  DLNUMFET %in% 1 & DLPARA > 0 & DLPRVCS > 0 & tolower(LABOUR) %in% "s", 5.1,
  DLNUMFET %in% 1 & DLPARA > 0 & DLPRVCS > 0 & tolower(LABOUR) %in% "i", 5.2,
  DLNUMFET %in% 1 & DLPARA > 0 & DLPRVCS > 0 & tolower(LABOUR) %in% "n", 5.3,
  DLNUMFET %in% 1 & tolower(POSATDEL) %in% c("bch","frb","ftb") & DLPARA %in% 0, 6,
  DLNUMFET %in% 1 & tolower(POSATDEL) %in% c("bch","frb","ftb") & DLPARA > 0, 7,
  DLNUMFET > 1, 8,
  DLNUMFET %in% 1 & tolower(POSATDEL) %in% c("shl","tli"), 9,
  DLNUMFET %in% 1 & GA_BEST < 37 & toupper(POSATDEL) %in% c('VTX','POP','BOW','CPD','FAC',''), 10,
  default = 0
  ))] %>%
  .[, `:=` (robson = fcase(
    robsongrp %in% 1, 1,
    robsongrp >= 2 & robsongrp < 3, 2,
    robsongrp %in% 3, 3,
    robsongrp >= 4 & robsongrp < 5, 4,
    robsongrp >= 5 & robsongrp < 6, 5,
    robsongrp %in% 6 | robsongrp %in% 7, 6,
    robsongrp %in% 8, 8,
    robsongrp %in% 10, 10,
    default = 0
  ),
  cs = fcase(
    tolower(DMMETHOD) %in% "cst", 1,
    default = 0
  ))] %>%
  #unique() %>%
  # remove missing
  .[!robson %in% 0] %>%
  .[, `:=` (count = .N), by = list(BrthYear, robson, cs)] %>%
  .[, `:=` (total = .N), by = list(BrthYear, robson)] %>%
  .[, `:=` (rate = count/total)] %>%
  unique()

# age category using UNICODE codes
dta1 <- dta1[, `:=` (robsonnam = fcase(
  robson %in% 1, "(1) Nulliparous, singleton, cephalic, term, spontaneous labour",
  robson %in% 2, "(2) Nulliparous, singleton, cephalic, term, induced or no labour",
  robson %in% 3, "(3) Multiparous, singleton, cephalic, term, no previous CS, spontaneous labour",
  robson %in% 4, "(4) Multiparous, singleton, cephalic, term, no previous CS, induced or no labour",
  robson %in% 5, "(5) Multiparous, singleton, cephalic, term, previous CS",
  robson %in% c(6,7), "(6-7) Nulliparous or multiparous, singleton, breech",
  robson %in% 8,"(8) Multiple pregnancy",
  robson %in% 10, "(10) Singleton, cephalic, preterm",
  default = ""
),
                     robson = factor(robson,
                                    levels = c(1,2,3,4,5,6,8,10),
                                    labels = c("1",
                                               "2",
                                               "3",
                                               "4",
                                               "5",
                                               "6-7",
                                               "8",
                                               "10")))]

## getting color for plotting

pal <- c("1" = "#1b9e77",
        "2" = "#d95f02",
        "3" = "#7570b3",
        "4" = "#e7298a",
        "5" = "#66a61e",
        "6-7" = "#e6ab02",
        "8" = "#a6761d",
        "10" = "#666666")

plot <- ggplot(data = dta1[!cs %in% 0],
               aes(x = BrthYear,
                   y = rate,
                   group = robson,
                   colour = robson,
                   fill = robson,
                   shape = robson,
                   text = paste(robsonnam,
                               "<br>Year:", BrthYear,
                               "<br>Total deliveries:", scales::comma(total, accuracy = 1),
                               paste0("<br>", robson, " (%):") , scales::percent(rate, accuracy = .1)
                               )
                   )
               ) +
  geom_point(aes(color = robson, fill = robson, shape = robson)) +
  geom_line(aes(color = robson, fill = robson)) +
# Specifying colors by age group
  scale_color_manual(name = "",
                     values = pal,
                     limits = names(pal)) +
  scale_fill_manual(name = "",
                    values = pal,
                    limits = names(pal)) +
  scale_shape_manual(name = "",
                     values = c(21,22,23,24,25,9,10,11)) +
  # scale_x_date(date_breaks = "2 years",
  #              date_labels = "%Y") +
  scale_x_continuous(#trans = 'log2',
    breaks = pretty) +
  scale_y_continuous(#trans = 'log2',
    breaks = pretty,
    labels = scales::percent_format(accuracy = 1),
    limits = c(0, NA)) +
  labs(y = 'Proportion of deliveries', x = '') +
  theme_classic()

plotly::ggplotly(plot, tooltip = c("text")) %>%
  plotly::style(hoverlabel = list(
    bgcolor  = "black",
    bordercolor = "transparent",
    font = list(
      color = "white",
      size = 14,
      face = "bold"
    )
  )) %>%
  plotly::layout(legend = list(orientation = "h", y = -0.25, x = 0.5,
                               xanchor = "center"),
                 font = list(face = "bold", size = 10),
                 showlegend = TRUE)
```

```{r tbl48}
tbl <- dta[, .(BrthYear,
                DLNUMFET,
                GA_BEST,
                POSATDEL,
                DLPARA,
                LABOUR,
                DLPRVCS,
                DMMETHOD)] %>%
  .[, `:=` (robsongrp = fcase(
  DLNUMFET %in% 1 & GA_BEST >= 37 & toupper(POSATDEL) %in% c('VTX','POP','BOW','CPD','FAC') & DLPARA %in% 0 & tolower(LABOUR) %in% "s", 1,
  DLNUMFET %in% 1 & GA_BEST >= 37 & toupper(POSATDEL) %in% c('VTX','POP','BOW','CPD','FAC') & DLPARA %in% 0 & tolower(LABOUR) %in% "i", 2.1,
  DLNUMFET %in% 1 & GA_BEST >= 37 & toupper(POSATDEL) %in% c('VTX','POP','BOW','CPD','FAC') & DLPARA %in% 0 & tolower(LABOUR) %in% "n", 2.2,
  DLNUMFET %in% 1 & DLPARA > 0 & DLPRVCS %in% 0 & tolower(LABOUR) %in% "s", 3,
  DLNUMFET %in% 1 & DLPARA > 0 & DLPRVCS %in% 0 & tolower(LABOUR) %in% "i", 4.1,
  DLNUMFET %in% 1 & DLPARA > 0 & DLPRVCS %in% 0 & tolower(LABOUR) %in% "n", 4.2,
  DLNUMFET %in% 1 & DLPARA > 0 & DLPRVCS > 0 & tolower(LABOUR) %in% "s", 5.1,
  DLNUMFET %in% 1 & DLPARA > 0 & DLPRVCS > 0 & tolower(LABOUR) %in% "i", 5.2,
  DLNUMFET %in% 1 & DLPARA > 0 & DLPRVCS > 0 & tolower(LABOUR) %in% "n", 5.3,
  DLNUMFET %in% 1 & tolower(POSATDEL) %in% c("bch","frb","ftb") & DLPARA %in% 0, 6,
  DLNUMFET %in% 1 & tolower(POSATDEL) %in% c("bch","frb","ftb") & DLPARA > 0, 7,
  DLNUMFET > 1, 8,
  DLNUMFET %in% 1 & tolower(POSATDEL) %in% c("shl","tli"), 9,
  DLNUMFET %in% 1 & GA_BEST < 37 & toupper(POSATDEL) %in% c('VTX','POP','BOW','CPD','FAC',''), 10,
  default = 0
  ))] %>%
  .[, `:=` (robson = fcase(
    robsongrp %in% 1, 1,
    robsongrp >= 2 & robsongrp < 3, 2,
    robsongrp %in% 3, 3,
    robsongrp >= 4 & robsongrp < 5, 4,
    robsongrp >= 5 & robsongrp < 6, 5,
    robsongrp %in% 6 | robsongrp %in% 7, 6,
    robsongrp %in% 8, 8,
    robsongrp %in% 10, 10,
    default = 0
  ),
  cs = fcase(
    tolower(DMMETHOD) %in% "cst", 1,
    default = 0
  ))] %>%
  #unique() %>%
  # remove missing
  .[!robson %in% 0] %>%
  .[, `:=` (count = .N), by = list(BrthYear, robson, cs)] %>%
  .[, `:=` (total = .N), by = list(BrthYear, robson)] %>%
  .[, `:=` (rate = 100*count/total)]

tbl1 <- tbl %>%
  .[robson %in% 1 &
      cs %in% 1] %>%
  melt(measure.vars = c("rate", "total", "count")) %>%
  .[!tolower(variable) %in% "count"] %>%
  .[, `:=` (robson = fifelse(tolower(variable) %in% "total",
                               0, robson))] %>%
  .[, .(BrthYear, robson, value)] %>%
  unique() %>%
  dcast(robson ~ BrthYear, value.var = "value")

# coln <- names(tbl1)
# coln[1] <- ""

tbl1 <- as.data.frame(sapply(tbl1, sprintf, fmt = c("%.0f", "%1.1f%%")))
#names(tbl1) <- coln

tbl1[,1] <- c("Total deliveries", "Proportion of caesarean")

# add comma separator to total count
tbl1[1,-1] <- as.character(sapply(as.numeric(tbl1[1,-1]), scales::comma_format(accuracy = 1)))

temp <- as.data.frame(t(c("(1) Nulliparous, singleton, cephalic, term, spontaneous labour", rep(NA, ncol(tbl1)-1))))
names(temp) <- names(tbl1)

tbl1 <- rbind(temp,tbl1)

tbl2 <- tbl %>%
  .[robson %in% 2 &
      cs %in% 1] %>%
  melt(measure.vars = c("rate", "total", "count")) %>%
  .[!tolower(variable) %in% "count"] %>%
  .[, `:=` (robson = fifelse(tolower(variable) %in% "total",
                               0, robson))] %>%
  .[, .(BrthYear, robson, value)] %>%
  unique() %>%
  dcast(robson ~ BrthYear, value.var = "value")

# coln <- names(tbl2)
# coln[1] <- ""

tbl2 <- as.data.frame(sapply(tbl2, sprintf, fmt = c("%.0f", "%1.1f%%")))
#names(tbl2) <- coln

tbl2[,1] <- c("Total deliveries", "Proportion of caesarean")

# add comma separator to total count
tbl2[1,-1] <- as.character(sapply(as.numeric(tbl2[1,-1]), scales::comma_format(accuracy = 1)))

temp <- as.data.frame(t(c("(2) Nulliparous, singleton, cephalic, term, induced or no labour", rep(NA, ncol(tbl2)-1))))
names(temp) <- names(tbl2)

tbl2 <- rbind(temp,tbl2)

tbl3 <- tbl %>%
  .[robson %in% 3 &
      cs %in% 1] %>%
  melt(measure.vars = c("rate", "total", "count")) %>%
  .[!tolower(variable) %in% "count"] %>%
  .[, `:=` (robson = fifelse(tolower(variable) %in% "total",
                               0, robson))] %>%
  .[, .(BrthYear, robson, value)] %>%
  unique() %>%
  dcast(robson ~ BrthYear, value.var = "value")

# coln <- names(tbl3)
# coln[1] <- ""

tbl3 <- as.data.frame(sapply(tbl3, sprintf, fmt = c("%.0f", "%1.1f%%")))
#names(tbl3) <- coln

tbl3[,1] <- c("Total deliveries", "Proportion of caesarean")

# add comma separator to total count
tbl3[1,-1] <- as.character(sapply(as.numeric(tbl3[1,-1]), scales::comma_format(accuracy = 1)))

temp <- as.data.frame(t(c("(3) Multiparous, singleton, cephalic, term, no previous CS, spontaneous labour", rep(NA, ncol(tbl3)-1))))
names(temp) <- names(tbl3)

tbl3 <- rbind(temp,tbl3)

tbl4 <- tbl %>%
  .[robson %in% 4 &
      cs %in% 1] %>%
  melt(measure.vars = c("rate", "total", "count")) %>%
  .[!tolower(variable) %in% "count"] %>%
  .[, `:=` (robson = fifelse(tolower(variable) %in% "total",
                               0, robson))] %>%
  .[, .(BrthYear, robson, value)] %>%
  unique() %>%
  dcast(robson ~ BrthYear, value.var = "value")

# coln <- names(tbl4)
# coln[1] <- ""

tbl4 <- as.data.frame(sapply(tbl4, sprintf, fmt = c("%.0f", "%1.1f%%")))
#names(tbl4) <- coln

tbl4[,1] <- c("Total deliveries", "Proportion of caesarean")

# add comma separator to total count
tbl4[1,-1] <- as.character(sapply(as.numeric(tbl4[1,-1]), scales::comma_format(accuracy = 1)))

temp <- as.data.frame(t(c("(4) Multiparous, singleton, cephalic, term, no previous CS, induced or no labour", rep(NA, ncol(tbl4)-1))))
names(temp) <- names(tbl4)

tbl4 <- rbind(temp,tbl4)

tbl5 <- tbl %>%
  .[robson %in% 5 &
      cs %in% 1] %>%
  melt(measure.vars = c("rate", "total", "count")) %>%
  .[!tolower(variable) %in% "count"] %>%
  .[, `:=` (robson = fifelse(tolower(variable) %in% "total",
                               0, robson))] %>%
  .[, .(BrthYear, robson, value)] %>%
  unique() %>%
  dcast(robson ~ BrthYear, value.var = "value")

# coln <- names(tbl5)
# coln[1] <- ""

tbl5 <- as.data.frame(sapply(tbl5, sprintf, fmt = c("%.0f", "%1.1f%%")))
#names(tbl5) <- coln

tbl5[,1] <- c("Total deliveries", "Proportion of caesarean")

# add comma separator to total count
tbl5[1,-1] <- as.character(sapply(as.numeric(tbl5[1,-1]), scales::comma_format(accuracy = 1)))

temp <- as.data.frame(t(c("(5) Multiparous, singleton, cephalic, term, previous CS", rep(NA, ncol(tbl5)-1))))
names(temp) <- names(tbl5)

tbl5 <- rbind(temp,tbl5)

tbl6 <- tbl %>%
  .[robson %in% 6 &
      cs %in% 1] %>%
  melt(measure.vars = c("rate", "total", "count")) %>%
  .[!tolower(variable) %in% "count"] %>%
  .[, `:=` (robson = fifelse(tolower(variable) %in% "total",
                               0, robson))] %>%
  .[, .(BrthYear, robson, value)] %>%
  unique() %>%
  dcast(robson ~ BrthYear, value.var = "value")

# coln <- names(tbl6)
# coln[1] <- ""

tbl6 <- as.data.frame(sapply(tbl6, sprintf, fmt = c("%.0f", "%1.1f%%")))
#names(tbl6) <- coln

tbl6[,1] <- c("Total deliveries", "Proportion of caesarean")

# add comma separator to total count
tbl6[1,-1] <- as.character(sapply(as.numeric(tbl6[1,-1]), scales::comma_format(accuracy = 1)))

temp <- as.data.frame(t(c("(6-7) Nulliparous or multiparous, singleton, breech", rep(NA, ncol(tbl6)-1))))
names(temp) <- names(tbl6)

tbl6 <- rbind(temp,tbl6)

tbl7 <- tbl %>%
  .[robson %in% 8 &
      cs %in% 1] %>%
  melt(measure.vars = c("rate", "total", "count")) %>%
  .[!tolower(variable) %in% "count"] %>%
  .[, `:=` (robson = fifelse(tolower(variable) %in% "total",
                               0, robson))] %>%
  .[, .(BrthYear, robson, value)] %>%
  unique() %>%
  dcast(robson ~ BrthYear, value.var = "value")

# coln <- names(tbl7)
# coln[1] <- ""

tbl7 <- as.data.frame(sapply(tbl7, sprintf, fmt = c("%.0f", "%1.1f%%")))
#names(tbl7) <- coln

tbl7[,1] <- c("Total deliveries", "Proportion of caesarean")

# add comma separator to total count
tbl7[1,-1] <- as.character(sapply(as.numeric(tbl7[1,-1]), scales::comma_format(accuracy = 1)))

temp <- as.data.frame(t(c("(8) Multiple pregnancy", rep(NA, ncol(tbl7)-1))))
names(temp) <- names(tbl7)

tbl7 <- rbind(temp,tbl7)

tbl8 <- tbl %>%
  .[robson %in% 10 &
      cs %in% 1] %>%
  melt(measure.vars = c("rate", "total", "count")) %>%
  .[!tolower(variable) %in% "count"] %>%
  .[, `:=` (robson = fifelse(tolower(variable) %in% "total",
                               0, robson))] %>%
  .[, .(BrthYear, robson, value)] %>%
  unique() %>%
  dcast(robson ~ BrthYear, value.var = "value")

# coln <- names(tbl8)
# coln[1] <- ""

tbl8 <- as.data.frame(sapply(tbl8, sprintf, fmt = c("%.0f", "%1.1f%%")))
#names(tbl8) <- coln

tbl8[,1] <- c("Total deliveries", "Proportion of caesarean")

# add comma separator to total count
tbl8[1,-1] <- as.character(sapply(as.numeric(tbl8[1,-1]), scales::comma_format(accuracy = 1)))

temp <- as.data.frame(t(c("(10) Singleton, cephalic, preterm", rep(NA, ncol(tbl8)-1))))
names(temp) <- names(tbl8)

tbl8 <- rbind(temp,tbl8)

tbl <- setDT(rbind(tbl1, tbl2, tbl3, tbl4,
                   tbl5, tbl6, tbl7, tbl8))

#names(tbl)[1] <- "Robson group"

reactable(tbl,
          rownames = FALSE,
          resizable = TRUE, 
          onClick = "expand",
          highlight = TRUE, 
          compact = TRUE,
          defaultPageSize = nrow(tbl),
          defaultColDef = colDef(align = "right",
                                 vAlign = "center",
                                 format = colFormat(locales = "en-CA", separators = TRUE),
                                 style = list(fontSize = 14),
                                 headerStyle = list(fontSize = 14),
                                 minWidth = 50,
                                 maxWidth = 180),
          rowStyle = function(index){
            if (index %in% (seq(1,nrow(tbl), by = 3))) list(color = "black",
                                         backgroundColor = "#e1e1e1",
                                         fontWeight = "bold",
                                         align = "center",
                                         borderBottom = "#dddddd")
          },
          columns = list(
            robson = colDef(name = "",
                             align = "left",
                             width = 180)
          )
) %>%
  reactablefmtr::add_source("The Robson criteria for the classification of deliveries into ten mutually exclusive groups by maternal characteristics allows comparison of Caesarean section rates at regional and national levels. Please note that for the purposes of this report: 
                            (1) group 6 (nulliparous breeches) and group 7 (multiparous breeches) are combined; and 
                            (2) group 9 (abnormal lies excluding breeches) is omitted due to small numbers. [Ref: Robson MS. Classification of caesarean sections. Fetal and Maternal Medicine Review 2001;12(1):23-39].",
                            font_size = 12,
                            #font_weight = "bold"
  )

# knitr::kable(tbl,
#              format = "html",
#              booktabs = TRUE) %>%
#   # add an hline before the total
# #kableExtra::row_spec(nrow(tbl) - 1, hline_after = TRUE) %>%  #kableExtra::row_spec(nrow(tbl), bold = TRUE) %>%
# kableExtra::row_spec(0, bold = TRUE) %>%
# #kableExtra::column_spec(1, border_right = TRUE, bold  = TRUE) %>%
# kableExtra::group_rows("(1) Nulliparous, singleton, cephalic, term, spontaneous labour ", start_row = 1, end_row = 2) %>%
#   kableExtra::group_rows("(2) Nulliparous, singleton, cephalic, term, induced or no labour", start_row = 3, end_row = 4) %>%
#   kableExtra::group_rows("(3) Multiparous, singleton, cephalic, term, no previous CS, spontaneous labour", start_row = 5, end_row = 6) %>%
#   kableExtra::group_rows("(4) Multiparous, singleton, cephalic, term, no previous CS, induced or no labour", start_row = 7, end_row = 8) %>%
#   kableExtra::group_rows("(5) Multiparous, singleton, cephalic, term, previous CS", start_row = 9, end_row = 10) %>%
#   kableExtra::group_rows("(6-7) Nulliparous or multiparous, singleton, breech", start_row = 11, end_row = 12) %>%
#   kableExtra::group_rows("(8) Multiple pregnancy", start_row = 13, end_row = 14) %>%
#   kableExtra::group_rows("(10) Singleton, cephalic, preterm", start_row = 15, end_row = 16) %>%
# kableExtra::kable_styling(font_size = 12, full_width = FALSE,
#                           bootstrap_options = c("hover","condensed", "responsive")) %>%
# kableExtra::add_footnote(label = c("The Robson criteria for the classification of deliveries into ten mutually exclusive groups by maternal characteristics allows comparison of Caesarean section rates at regional and national levels. Please note that for the purposes of this report: (1) group 6 (nulliparous breeches) and group 7 (multiparous breeches) are combined; and (2) group 9 (abnormal lies excluding breeches) is omitted due to small numbers. [Ref: Robson MS. Classification of caesarean sections. Feta and Maternal Medicine Review 2001;12(1):23-39]."))
```

## Any labour among candidates for vaginal birth after Caesarean by year, `r paste0(params$year1,"-",params$year2)` {#section-49}

```{r plot49, fig.cap= paste('Any labour among candidates for vaginal birth after Caesarean by year, Nova Scotia,', paste0(params$year1,"-",params$year2))}

dta1 <- dta[, .(BrthYear,
                LABOUR,
                DMMETHOD,
                POSATDEL,
                DLNUMFET,
                DLPRVCS,
                INDICCS1)] %>%
  .[, `:=` (lab = fcase(
    !tolower(LABOUR) %in% "n", 1, 
    default = 0
  ),
  vbac = fcase(
    tolower(POSATDEL) %in% c("vtx", "pop", "") & DLNUMFET %in% 1 & DLPRVCS %in% 1 & !toupper(INDICCS1) %in% c("PLP","BCH","MLP","HSV","SFA","UTS","CXD","HIV"), 1,
    default = 0
  ))] %>%
  #unique() %>%
  # remove missing
  .[!vbac %in% 0] %>%
  .[, `:=` (count = .N), by = list(BrthYear, lab)] %>%
  .[, `:=` (total = .N), by = list(BrthYear)] %>%
  .[, `:=` (rate = count/total)] %>%
  unique()

# age category using UNICODE codes
dta1 <- dta1[, `:=` (lab = factor(lab,
                                    levels = c(0,1),
                                    labels = c("No labour and no induction attempt",
                                               "Labour or induction attempt")))]

## getting color for plotting

pal <- c("No labour and no induction attempt" = "#1b9e77",
        "Labour or induction attempt" = "#d95f02")

plot <- ggplot(data = dta1,
               aes(x = BrthYear,
                   y = rate,
                   group = lab,
                   colour = lab,
                   fill = lab,
                   shape = lab,
                   text = paste(lab,
                               "<br>Year:", BrthYear,
                               "<br>Total VBAC cadidates:", scales::comma(total, accuracy = 1),
                               paste0("<br>",lab," (%):") , scales::percent(rate, accuracy = .1)
                               )
                   )
               ) +
  geom_point(aes(color = lab, fill = lab, shape = lab)) +
  geom_line(aes(color = lab, fill = lab)) +
# Specifying colors by age group
  scale_color_manual(name = "",
                     values = pal,
                     limits = names(pal)) +
  scale_fill_manual(name = "",
                    values = pal,
                    limits = names(pal)) +
  scale_shape_manual(name = "",
                     values = c(21,22)) +
  # scale_x_date(date_breaks = "2 years",
  #              date_labels = "%Y") +
  scale_x_continuous(#trans = 'log2',
    breaks = pretty) +
  scale_y_continuous(#trans = 'log2',
    breaks = pretty,
    labels = scales::percent_format(accuracy = 1),
    limits = c(0, NA)) +
  labs(y = 'Proportion of VBAC candidates', x = '') +
  theme_classic()

plotly::ggplotly(plot, tooltip = c("text")) %>%
  plotly::style(hoverlabel = list(
    bgcolor  = "black",
    bordercolor = "transparent",
    font = list(
      color = "white",
      size = 14,
      face = "bold"
    )
  )) %>%
  plotly::layout(legend = list(orientation = "h", y = -0.25, x = 0.5,
                               xanchor = "center"),
                 font = list(face = "bold", size = 10),
                 showlegend = TRUE)
```

```{r tbl49}
tbl <- dta[, .(BrthYear,
                LABOUR,
                DMMETHOD,
                POSATDEL,
                DLNUMFET,
                DLPRVCS,
                INDICCS1)] %>%
  .[, `:=` (lab = fcase(
    !tolower(LABOUR) %in% "n", 1, 
    default = 0
  ),
  vbac = fcase(
    tolower(POSATDEL) %in% c("vtx", "pop", "") & DLNUMFET %in% 1 & DLPRVCS %in% 1 & !toupper(INDICCS1) %in% c("PLP","BCH","MLP","HSV","SFA","UTS","CXD","HIV"), 1,
    default = 0
  ))] %>%
  #unique() %>%
  # remove missing
  .[!vbac %in% 0] %>%
  .[, `:=` (count = .N), by = list(BrthYear, lab)] %>%
  .[, `:=` (total = .N), by = list(BrthYear)] %>%
  .[, `:=` (rate = 100*count/total)] %>%
  melt(measure.vars = c("rate", "total", "count")) %>%
  .[!tolower(variable) %in% "count"] %>%
  .[, `:=` (lab = fifelse(tolower(variable) %in% "total",
                               -1, lab))] %>%
  .[, .(BrthYear, lab, value)] %>%
  unique() %>%
  dcast(lab ~ BrthYear, value.var = "value")

# coln <- names(tbl)
# coln[1] <- ""

tbl <- as.data.frame(sapply(tbl, sprintf, fmt = c("%.0f", "%1.1f%%", "%1.1f%%")))
#names(tbl) <- coln

tbl[,1] <- c("Total VBAC candidates", "Labour or induction attempt", "No labour and no induction attempt")

# add comma separator to total count
tbl[1,-1] <- as.character(sapply(as.numeric(tbl[1,-1]), scales::comma_format(accuracy = 1)))

reactable(tbl,
          rownames = FALSE,
          resizable = TRUE, 
          onClick = "expand",
          highlight = TRUE, 
          compact = TRUE,
          defaultPageSize = nrow(tbl),
          defaultColDef = colDef(align = "right",
                                 vAlign = "center",
                                 format = colFormat(locales = "en-CA", separators = TRUE),
                                 style = list(fontSize = 14),
                                 headerStyle = list(fontSize = 14),
                                 minWidth = 50,
                                 maxWidth = 180),
          # rowStyle = function(index){
          #   if (index == nrow(tbl)) list(color = "black",
          #                                backgroundColor = "#e1e1e1",
          #                                fontWeight = "bold",
          #                                align = "center",
          #                                borderBottom = "#dddddd")
          # },
          columns = list(
            lab = colDef(name = "",
                             align = "left",
                             width = 180)
          )
) %>%
  reactablefmtr::add_source("Note: For the purposes of this report, a candidate for vaginal birth after Caesarean (VBAC) is woman who has had no more than one previous Caesarean section delivery; whose current pregnancy is a singleton in vertex presentation; and who has no contraindications for labour. On an individual basis when more information is available, such as type of previous Caesarean delivery, other factors are taken into account and women with two previous Caesarean deliveries may be considered for VBAC. [Ref: Society of Obstetricians and Gynaecologists of Canada. Guidelines for vaginal birth after previous caesarean birth. SOGC clinical practice guidelines. Number 155, February 2005. International Journal of Gynaecology and Obstetrics 2005;89(3):319-31].",
                            font_size = 12,
                            #font_weight = "bold"
  )

# knitr::kable(tbl,
#              format = "html",
#              booktabs = TRUE) %>%
#   # add an hline before the total
# #kableExtra::row_spec(nrow(tbl) - 1, hline_after = TRUE) %>%  #kableExtra::row_spec(nrow(tbl), bold = TRUE) %>%
# kableExtra::row_spec(0, bold = TRUE) %>%
# #kableExtra::column_spec(1, border_right = TRUE, bold  = TRUE) %>%
# kableExtra::kable_styling(font_size = 12, full_width = FALSE,
#                           bootstrap_options = c("hover","condensed", "responsive")) %>%
# kableExtra::add_footnote(label = c("Note: For the purposes of this report, a candidate for vaginal birth after Caesarean (VBAC) is woman who has had no more than one previous Caesarean section delivery; whose current pregnancy is a singleton in vertex presentation; and who has no contraindications for labour. On an individual basis when more information is available, such as type of previous Caesarean delivery, other factors are taken into account and women with two previous Caesarean deliveries may be considered for VBAC. [Ref: Society of Obstetricians and Gynaecologists of Canada. Guidelines for vaginal birth after previous caesarean birth. SOGC clinical practice guidelines. Number 155, February 2005. International Journal of Gynaecology and Obstetrics 2005;89(3):319-31]."))
```

## Type of delivery among candidates for vaginal birth after Caesarean who had any labour by year, `r paste0(params$year1,"-",params$year2)` {#section-410}

```{r plot410, fig.cap= paste('Type of delivery among candidates for vaginal birth after Caesarean who had any labour by year, Nova Scotia,', paste0(params$year1,"-",params$year2))}

dta1 <- dta[, .(BrthYear,
                LABOUR,
                DMMETHOD,
                POSATDEL,
                DLNUMFET,
                DLPRVCS,
                INDICCS1)] %>%
  .[, `:=` (lab = fcase(
    !tolower(LABOUR) %in% "n", 1, 
    default = 0
  ),
  vbac = fcase(
    tolower(POSATDEL) %in% c("vtx", "pop", "") & DLNUMFET %in% 1 & DLPRVCS %in% 1 & !toupper(INDICCS1) %in% c("PLP","BCH","MLP","HSV","SFA","UTS","CXD","HIV"), 1,
    default = 0
  ),
  deliv = fcase(
      # vaginal
      tolower(DMMETHOD) %in% c("biv", "spt", "vac", "low", "mid"), 1,
      # c-section
      tolower(DMMETHOD) %in% c("cst"), 2,
      default = 3
    ))] %>%
  #unique() %>%
  # remove missing
  .[!lab %in% 0 &
    !vbac %in% 0] %>%
  .[, `:=` (count = .N), by = list(BrthYear, deliv)] %>%
  .[, `:=` (total = .N), by = list(BrthYear)] %>%
  .[, `:=` (rate = count/total)] %>%
  unique()

# age category using UNICODE codes
dta1 <- dta1[, `:=` (deliv = factor(deliv,
                                    levels = c(1,2),
                                    labels = c("Vaginal delivery",
                                               "Caesarean section")))]

## getting color for plotting

pal <- c("Vaginal delivery" = "#1b9e77",
         "Caesarean section" = "#d95f02")

plot <- ggplot(data = dta1,
               aes(x = BrthYear,
                   y = rate,
                   group = deliv,
                   colour = deliv,
                   fill = deliv,
                   shape = deliv,
                   text = paste(deliv,
                               "<br>Year:", BrthYear,
                               "<br>Total women with labour or induction:", scales::comma(total, accuracy = 1),
                               paste0("<br>",deliv, " (%):") , scales::percent(rate, accuracy = .1)
                               )
                   )
               ) +
  geom_point(aes(color = deliv, fill = deliv, shape = deliv)) +
  geom_line(aes(color = deliv, fill = deliv)) +
# Specifying colors by age group
  scale_color_manual(name = "",
                     values = pal,
                     limits = names(pal)) +
  scale_fill_manual(name = "",
                    values = pal,
                    limits = names(pal)) +
  scale_shape_manual(name = "",
                     values = c(21,22)) +
  # scale_x_date(date_breaks = "2 years",
  #              date_labels = "%Y") +
  scale_x_continuous(#trans = 'log2',
    breaks = pretty) +
  scale_y_continuous(#trans = 'log2',
    breaks = pretty,
    labels = scales::percent_format(accuracy = 1),
    limits = c(0, NA)) +
  labs(y = 'Proportion of women with labour or induction', x = '') +
  theme_classic()

plotly::ggplotly(plot, tooltip = c("text")) %>%
  plotly::style(hoverlabel = list(
    bgcolor  = "black",
    bordercolor = "transparent",
    font = list(
      color = "white",
      size = 14,
      face = "bold"
    )
  )) %>%
  plotly::layout(legend = list(orientation = "h", y = -0.25, x = 0.5,
                               xanchor = "center"),
                 font = list(face = "bold", size = 10),
                 showlegend = TRUE)
```

```{r tbl410}
tbl <- dta[, .(BrthYear,
                LABOUR,
                DMMETHOD,
                POSATDEL,
                DLNUMFET,
                DLPRVCS,
                INDICCS1)] %>%
  .[, `:=` (lab = fcase(
    !tolower(LABOUR) %in% "n", 1, 
    default = 0
  ),
  vbac = fcase(
    tolower(POSATDEL) %in% c("vtx", "pop", "") & DLNUMFET %in% 1 & DLPRVCS %in% 1 & !toupper(INDICCS1) %in% c("PLP","BCH","MLP","HSV","SFA","UTS","CXD","HIV"), 1,
    default = 0
  ),
  deliv = fcase(
      # vaginal
      tolower(DMMETHOD) %in% c("biv", "spt", "vac", "low", "mid"), 1,
      # c-section
      tolower(DMMETHOD) %in% c("cst"), 2,
      default = 3
    ))] %>%
  #unique() %>%
  # remove missing
  .[!lab %in% 0 &
    !vbac %in% 0] %>%
  .[, `:=` (count = .N), by = list(BrthYear, deliv)] %>%
  .[, `:=` (total = .N), by = list(BrthYear)] %>%
  .[, `:=` (rate = 100*count/total)] %>%
  melt(measure.vars = c("rate", "total", "count")) %>%
  .[!tolower(variable) %in% "count"] %>%
  .[, `:=` (deliv = fifelse(tolower(variable) %in% "total",
                               0, deliv))] %>%
  .[, .(BrthYear, deliv, value)] %>%
  unique() %>%
  dcast(deliv ~ BrthYear, value.var = "value")

# coln <- names(tbl)
# coln[1] <- ""

tbl <- as.data.frame(sapply(tbl, sprintf, fmt = c("%.0f", "%1.1f%%", "%1.1f%%")))
#names(tbl) <- coln

tbl[,1] <- c("Total women with labour or induction", "Vaginal delivery", "Caesarean section")

# add comma separator to total count
tbl[1,-1] <- as.character(sapply(as.numeric(tbl[1,-1]), scales::comma_format(accuracy = 1)))

reactable(tbl,
          rownames = FALSE,
          resizable = TRUE, 
          onClick = "expand",
          highlight = TRUE, 
          compact = TRUE,
          defaultPageSize = nrow(tbl),
          defaultColDef = colDef(align = "right",
                                 vAlign = "center",
                                 format = colFormat(locales = "en-CA", separators = TRUE),
                                 style = list(fontSize = 14),
                                 headerStyle = list(fontSize = 14),
                                 minWidth = 50,
                                 maxWidth = 180),
          # rowStyle = function(index){
          #   if (index == nrow(tbl)) list(color = "black",
          #                                backgroundColor = "#e1e1e1",
          #                                fontWeight = "bold",
          #                                align = "center",
          #                                borderBottom = "#dddddd")
          # },
          columns = list(
            deliv = colDef(name = "",
                             align = "left",
                             width = 180)
          )
)

# knitr::kable(tbl,
#              format = "html",
#              booktabs = TRUE) %>%
#   # add an hline before the total
# #kableExtra::row_spec(nrow(tbl) - 1, hline_after = TRUE) %>%  #kableExtra::row_spec(nrow(tbl), bold = TRUE) %>%
# kableExtra::row_spec(0, bold = TRUE) %>%
# #kableExtra::column_spec(1, border_right = TRUE, bold  = TRUE) %>%
# kableExtra::kable_styling(font_size = 12, full_width = FALSE,
#                           bootstrap_options = c("hover","condensed", "responsive"))
```

## Labour to 4 cm dilation among candidates for vaginal birth after Caesarean by year, `r paste0(params$year1,"-",params$year2)` {#section-411}

```{r plot411, fig.cap= paste('Labour to 4 cm dilation among candidates for vaginal birth after Caesarean by year, Nova Scotia,', paste0(params$year1,"-",params$year2))}

dta1 <- dta[, .(BrthYear,
                LABOUR,
                DMMETHOD,
                POSATDEL,
                DLNUMFET,
                DLPRVCS,
                INDICCS1,
                CDILCS)] %>%
  .[, `:=` (vbac = fcase(
    tolower(POSATDEL) %in% c("vtx", "pop", "") & DLNUMFET %in% 1 & DLPRVCS %in% 1 & !toupper(INDICCS1) %in% c("PLP","BCH","MLP","HSV","SFA","UTS","CXD","HIV"), 1,
    default = 0
  ),
  vbac4 = fcase(
    !tolower(DMMETHOD) %in% c("cst") | CDILCS >= 4, 1,
    tolower(DMMETHOD) %in% c("cst") & CDILCS < 4, 2,
    default = 0
  ))] %>%
  #unique() %>%
  # remove missing
  .[!vbac4 %in% 0 &
    !vbac %in% 0] %>%
  .[, `:=` (count = .N), by = list(BrthYear, vbac4, vbac)] %>%
  .[, `:=` (total = .N), by = list(BrthYear, vbac)] %>%
  .[, `:=` (rate = count/total)] %>%
  unique()

# age category using UNICODE codes
dta1 <- dta1[, `:=` (vbac4 = factor(vbac4,
                                    levels = c(1,2),
                                    labels = c("Dilation to 4 cm or induction attempt",
                                               "Less than 4 cm and no induction attempt")))]

## getting color for plotting

pal <- c("Dilation to 4 cm or induction attempt" = "#1b9e77",
         "Less than 4 cm and no induction attempt" = "#d95f02")

plot <- ggplot(data = dta1,
               aes(x = BrthYear,
                   y = rate,
                   group = vbac4,
                   colour = vbac4,
                   fill = vbac4,
                   shape = vbac4,
                   text = paste(vbac4,
                               "<br>Year:", BrthYear,
                               "<br>Total VBAC candidates:", scales::comma(total, accuracy = 1),
                               paste0("<br>",vbac4) , scales::percent(rate, accuracy = .1)
                               )
                   )
               ) +
  geom_point(aes(color = vbac4, fill = vbac4, shape = vbac4)) +
  geom_line(aes(color = vbac4, fill = vbac4)) +
# Specifying colors by age group
  scale_color_manual(name = "",
                     values = pal,
                     limits = names(pal)) +
  scale_fill_manual(name = "",
                    values = pal,
                    limits = names(pal)) +
  scale_shape_manual(name = "",
                     values = c(21,22)) +
  # scale_x_date(date_breaks = "2 years",
  #              date_labels = "%Y") +
  scale_x_continuous(#trans = 'log2',
    breaks = pretty) +
  scale_y_continuous(#trans = 'log2',
    breaks = pretty,
    labels = scales::percent_format(accuracy = 1),
    limits = c(0, NA)) +
  labs(y = 'Proportion of VBAC candidates', x = '') +
  theme_classic()

plotly::ggplotly(plot, tooltip = c("text")) %>%
  plotly::style(hoverlabel = list(
    bgcolor  = "black",
    bordercolor = "transparent",
    font = list(
      color = "white",
      size = 14,
      face = "bold"
    )
  )) %>%
  plotly::layout(legend = list(orientation = "h", y = -0.25, x = 0.5,
                               xanchor = "center"),
                 font = list(face = "bold", size = 10),
                 showlegend = TRUE)
```

```{r tbl411}
tbl <- dta[, .(BrthYear,
                LABOUR,
                DMMETHOD,
                POSATDEL,
                DLNUMFET,
                DLPRVCS,
                INDICCS1,
                CDILCS)] %>%
  .[, `:=` (vbac = fcase(
    tolower(POSATDEL) %in% c("vtx", "pop", "") & DLNUMFET %in% 1 & DLPRVCS %in% 1 & !toupper(INDICCS1) %in% c("PLP","BCH","MLP","HSV","SFA","UTS","CXD","HIV"), 1,
    default = 0
  ),
  vbac4 = fcase(
    !tolower(DMMETHOD) %in% c("cst") | CDILCS >= 4, 1,
    tolower(DMMETHOD) %in% c("cst") & CDILCS < 4, 2,
    default = 0
  ))] %>%
  #unique() %>%
  # remove missing
  .[!vbac4 %in% 0 &
    !vbac %in% 0] %>%
  .[, `:=` (count = .N), by = list(BrthYear, vbac4, vbac)] %>%
  .[, `:=` (total = .N), by = list(BrthYear, vbac)] %>%
  .[, `:=` (rate = 100*count/total)] %>%
  melt(measure.vars = c("rate", "total", "count")) %>%
  .[!tolower(variable) %in% "count"] %>%
  .[, `:=` (vbac4 = fifelse(tolower(variable) %in% "total",
                               0, vbac4))] %>%
  .[, .(BrthYear, vbac4, value)] %>%
  unique() %>%
  dcast(vbac4 ~ BrthYear, value.var = "value")

# coln <- names(tbl)
# coln[1] <- ""

tbl <- as.data.frame(sapply(tbl, sprintf, fmt = c("%.0f", "%1.1f%%", "%1.1f%%")))
#names(tbl) <- coln

tbl[,1] <- c("Total VBAC candidates", "Dilation to 4 cm or induction attempt", "Less than 4 cm and no induction attempt")

# add comma separator to total count
tbl[1,-1] <- as.character(sapply(as.numeric(tbl[1,-1]), scales::comma_format(accuracy = 1)))

reactable(tbl,
          rownames = FALSE,
          resizable = TRUE, 
          onClick = "expand",
          highlight = TRUE, 
          compact = TRUE,
          defaultPageSize = nrow(tbl),
          defaultColDef = colDef(align = "right",
                                 vAlign = "center",
                                 format = colFormat(locales = "en-CA", separators = TRUE),
                                 style = list(fontSize = 14),
                                 headerStyle = list(fontSize = 14),
                                 minWidth = 50,
                                 maxWidth = 180),
          # rowStyle = function(index){
          #   if (index == nrow(tbl)) list(color = "black",
          #                                backgroundColor = "#e1e1e1",
          #                                fontWeight = "bold",
          #                                align = "center",
          #                                borderBottom = "#dddddd")
          # },
          columns = list(
            vbac4 = colDef(name = "",
                             align = "left",
                             width = 180)
          )
) %>%
  reactablefmtr::add_source("\u1d43 VBAC candidates with known cervical dilation who reached 4 cm.",
                            font_size = 12,
                            #font_weight = "bold"
  ) %>%
  reactablefmtr::add_source("Note: Women who are VBAC candidates and reach 4 cm cervical dilation may better represent those who have chosen to attempt a vaginal delivery. Intention to attempt a vaginal delivery is not recorded in the Atlee Database.",
                            font_size = 12,
                            #font_weight = "bold"
  )

# knitr::kable(tbl,
#              format = "html",
#              booktabs = TRUE) %>%
#   # add an hline before the total
# #kableExtra::row_spec(nrow(tbl) - 1, hline_after = TRUE) %>%  #kableExtra::row_spec(nrow(tbl), bold = TRUE) %>%
# kableExtra::row_spec(0, bold = TRUE) %>%
# #kableExtra::column_spec(1, border_right = TRUE, bold  = TRUE) %>%
# kableExtra::kable_styling(font_size = 12, full_width = FALSE,
#                           bootstrap_options = c("hover","condensed", "responsive")) %>%
# kableExtra::add_footnote(label = c(" VBAC candidates with known cervical dilation who reached 4 cm.",
#                                    " Note: Women who are VBAC candidates and reach 4 cm cervical dilation may better represent those who have chosen to attempt a vaginal delivery. Intention to attempt a vaginal delivery is not recorded in the Atlee Database."))
```

## Type of delivery among candidates for vaginal birth after Caesarean who had labour to 4 cm dilation by year, `r paste0(params$year1,"-",params$year2)` {#section-412}

```{r plot412, fig.cap= paste('Type of delivery among candidates for vaginal birth after Caesarean who had labour to 4 cm dilation by year, Nova Scotia,', paste0(params$year1,"-",params$year2))}

dta1 <- dta[, .(BrthYear,
                LABOUR,
                DMMETHOD,
                POSATDEL,
                DLNUMFET,
                DLPRVCS,
                INDICCS1,
                CDILCS)] %>%
  .[, `:=` (vbac = fcase(
    tolower(POSATDEL) %in% c("vtx", "pop", "") & DLNUMFET %in% 1 & DLPRVCS %in% 1 & !toupper(INDICCS1) %in% c("PLP","BCH","MLP","HSV","SFA","UTS","CXD","HIV"), 1,
    default = 0
  ),
  vbac4 = fcase(
    !tolower(DMMETHOD) %in% c("cst") | CDILCS >= 4, 1,
    tolower(DMMETHOD) %in% c("cst") & CDILCS < 4, 2,
    default = 0
  ),
  deliv = fcase(
      # vaginal
      tolower(DMMETHOD) %in% c("biv", "spt", "vac", "low", "mid"), 1,
      # c-section
      tolower(DMMETHOD) %in% c("cst"), 2,
      default = 3
    ))] %>%
  #unique() %>%
  # remove missing
  .[!vbac4 %in% c(0,2) &
    !vbac %in% 0] %>%
  .[, `:=` (count = .N), by = list(BrthYear, vbac4, deliv)] %>%
  .[, `:=` (total = .N), by = list(BrthYear, vbac4)] %>%
  .[, `:=` (rate = count/total)] %>%
  unique()

# age category using UNICODE codes
dta1 <- dta1[, `:=` (deliv = factor(deliv,
                                    levels = c(1,2),
                                    labels = c("Vaginal delivery",
                                               "Caesarean section")))]

## getting color for plotting

pal <- c("Vaginal delivery" = "#1b9e77",
         "Caesarean section" = "#d95f02")

plot <- ggplot(data = dta1,
               aes(x = BrthYear,
                   y = rate,
                   group = deliv,
                   colour = deliv,
                   fill = deliv,
                   shape = deliv,
                   text = paste(deliv,
                               "<br>Year:", BrthYear,
                               "<br>Total with 4 cm labour or induction", scales::comma(total, accuracy = 1),
                               paste0("<br>%",deliv," (%):") , scales::percent(rate, accuracy = .1)
                               )
                   )
               ) +
  geom_point(aes(color = deliv, fill = deliv, shape = deliv)) +
  geom_line(aes(color = deliv, fill = deliv)) +
# Specifying colors by age group
  scale_color_manual(name = "",
                     values = pal,
                     limits = names(pal)) +
  scale_fill_manual(name = "",
                    values = pal,
                    limits = names(pal)) +
  scale_shape_manual(name = "",
                     values = c(21,22)) +
  # scale_x_date(date_breaks = "2 years",
  #              date_labels = "%Y") +
  scale_x_continuous(#trans = 'log2',
    breaks = pretty) +
  scale_y_continuous(#trans = 'log2',
    breaks = pretty,
    labels = scales::percent_format(accuracy = 1),
    limits = c(0, NA)) +
  labs(y = 'Proportion with 4 cm labour or induction', x = '') +
  theme_classic()

plotly::ggplotly(plot, tooltip = c("text")) %>%
  plotly::style(hoverlabel = list(
    bgcolor  = "black",
    bordercolor = "transparent",
    font = list(
      color = "white",
      size = 14,
      face = "bold"
    )
  )) %>%
  plotly::layout(legend = list(orientation = "h", y = -0.25, x = 0.5,
                               xanchor = "center"),
                 font = list(face = "bold", size = 10),
                 showlegend = TRUE)
```

```{r tbl412}
tbl <- dta[, .(BrthYear,
                LABOUR,
                DMMETHOD,
                POSATDEL,
                DLNUMFET,
                DLPRVCS,
                INDICCS1,
                CDILCS)] %>%
  .[, `:=` (vbac = fcase(
    tolower(POSATDEL) %in% c("vtx", "pop", "") & DLNUMFET %in% 1 & DLPRVCS %in% 1 & !toupper(INDICCS1) %in% c("PLP","BCH","MLP","HSV","SFA","UTS","CXD","HIV"), 1,
    default = 0
  ),
  vbac4 = fcase(
    !tolower(DMMETHOD) %in% c("cst") | CDILCS >= 4, 1,
    tolower(DMMETHOD) %in% c("cst") & CDILCS < 4, 2,
    default = 0
  ),
  deliv = fcase(
      # vaginal
      tolower(DMMETHOD) %in% c("biv", "spt", "vac", "low", "mid"), 1,
      # c-section
      tolower(DMMETHOD) %in% c("cst"), 2,
      default = 3
    ))] %>%
  #unique() %>%
  # remove missing
  .[!vbac4 %in% c(0,2) &
    !vbac %in% 0] %>%
  .[, `:=` (count = .N), by = list(BrthYear, vbac4, deliv)] %>%
  .[, `:=` (total = .N), by = list(BrthYear, vbac4)] %>%
  .[, `:=` (rate = 100*count/total)] %>%
  melt(measure.vars = c("rate", "total", "count")) %>%
  .[!tolower(variable) %in% "count"] %>%
  .[, `:=` (deliv = fifelse(tolower(variable) %in% "total",
                               0, deliv))] %>%
  .[, .(BrthYear, deliv, value)] %>%
  unique() %>%
  dcast(deliv ~ BrthYear, value.var = "value")

# coln <- names(tbl)
# coln[1] <- ""

tbl <- as.data.frame(sapply(tbl, sprintf, fmt = c("%.0f", "%1.1f%%", "%1.1f%%")))
#names(tbl) <- coln

tbl[,1] <- c("Total with 4 cm labour or induction", "Vaginal delivery", "Caesarean section")

# add comma separator to total count
tbl[1,-1] <- as.character(sapply(as.numeric(tbl[1,-1]), scales::comma_format(accuracy = 1)))

reactable(tbl,
          rownames = FALSE,
          resizable = TRUE, 
          onClick = "expand",
          highlight = TRUE, 
          compact = TRUE,
          defaultPageSize = nrow(tbl),
          defaultColDef = colDef(align = "right",
                                 vAlign = "center",
                                 format = colFormat(locales = "en-CA", separators = TRUE),
                                 style = list(fontSize = 14),
                                 headerStyle = list(fontSize = 14),
                                 minWidth = 50,
                                 maxWidth = 180),
          # rowStyle = function(index){
          #   if (index == nrow(tbl)) list(color = "black",
          #                                backgroundColor = "#e1e1e1",
          #                                fontWeight = "bold",
          #                                align = "center",
          #                                borderBottom = "#dddddd")
          # },
          columns = list(
            deliv = colDef(name = "",
                             align = "left",
                             width = 180)
          )
)

# knitr::kable(tbl,
#              format = "html",
#              booktabs = TRUE) %>%
#   # add an hline before the total
# #kableExtra::row_spec(nrow(tbl) - 1, hline_after = TRUE) %>%  #kableExtra::row_spec(nrow(tbl), bold = TRUE) %>%
# kableExtra::row_spec(0, bold = TRUE) %>%
# #kableExtra::column_spec(1, border_right = TRUE, bold  = TRUE) %>%
# kableExtra::kable_styling(font_size = 12, full_width = FALSE,
#                           bootstrap_options = c("hover","condensed", "responsive"))
```

## Episiotomy by parity and year, `r paste0(params$year1,"-",params$year2)` {#section-413}

```{r plot413, fig.cap= paste('Episiotomy by parity and year, Nova Scotia,', paste0(params$year1,"-",params$year2))}

dta1 <- dta[, .(BrthYear,
                DMMETHOD,
                DMEPISIO,
                DLPARA)] %>%
  .[, `:=` (deliv = fcase(
      # vaginal
      tolower(DMMETHOD) %in% c("biv", "spt", "vac", "low", "mid"), 1,
      # c-section
      tolower(DMMETHOD) %in% c("cst"), 2,
      default = 3
    ),
    epi = fcase(
      tolower(DMEPISIO) %in% c("md","ml"), 1,
      default = 0
    ),
    parity = fcase(
    DLPARA %in% 0 , 1,
    DLPARA >= 1, 2,
    default = 3
    ))] %>%
  #unique() %>%
  # remove missing
  .[deliv %in% 1 &
    !is.na(DLPARA)] %>%
  .[, `:=` (count = .N), by = list(BrthYear, epi, parity)] %>%
  .[, `:=` (total = .N), by = list(BrthYear, parity)] %>%
  .[, `:=` (rate = count/total)] %>%
  unique()

# age category using UNICODE codes
dta1 <- dta1[, `:=` (parity = factor(parity,
                                    levels = c(1,2),
                                    labels = c("Primiparous",
                                               "Multiparous")))]

## getting color for plotting

pal <- c("Primiparous" = "#1b9e77",
         "Multiparous" = "#d95f02")

plot <- ggplot(data = dta1[epi %in% 1],
               aes(x = BrthYear,
                   y = rate,
                   group = parity,
                   colour = parity,
                   fill = parity,
                   shape = parity,
                   text = paste(parity,
                               "<br>Year:", BrthYear,
                               "<br>Total vagina deliveries", scales::comma(total, accuracy = 1),
                               paste0("<br>%",parity, " (%):") , scales::percent(rate, accuracy = .1)
                               )
                   )
               ) +
  geom_point(aes(color = parity, fill = parity, shape = parity)) +
  geom_line(aes(color = parity, fill = parity)) +
# Specifying colors by age group
  scale_color_manual(name = "",
                     values = pal,
                     limits = names(pal)) +
  scale_fill_manual(name = "",
                    values = pal,
                    limits = names(pal)) +
  scale_shape_manual(name = "",
                     values = c(21,22)) +
  # scale_x_date(date_breaks = "2 years",
  #              date_labels = "%Y") +
  scale_x_continuous(#trans = 'log2',
    breaks = pretty) +
  scale_y_continuous(#trans = 'log2',
    breaks = pretty,
    labels = scales::percent_format(accuracy = 1),
    limits = c(0, NA)) +
  labs(y = 'Proportion of vaginal deliveries', x = '') +
  theme_classic()

plotly::ggplotly(plot, tooltip = c("text")) %>%
  plotly::style(hoverlabel = list(
    bgcolor  = "black",
    bordercolor = "transparent",
    font = list(
      color = "white",
      size = 14,
      face = "bold"
    )
  )) %>%
  plotly::layout(legend = list(orientation = "h", y = -0.25, x = 0.5,
                               xanchor = "center"),
                 font = list(face = "bold", size = 10),
                 showlegend = TRUE)
```

```{r tbl413}
tbl <-  dta[, .(BrthYear,
                DMMETHOD,
                DMEPISIO,
                DLPARA)] %>%
  .[, `:=` (deliv = fcase(
      # vaginal
      tolower(DMMETHOD) %in% c("biv", "spt", "vac", "low", "mid"), 1,
      # c-section
      tolower(DMMETHOD) %in% c("cst"), 2,
      default = 3
    ),
    epi = fcase(
      tolower(DMEPISIO) %in% c("md","ml"), 1,
      default = 0
    ),
    parity = fcase(
    DLPARA %in% 0 , 1,
    DLPARA >= 1, 2,
    default = 3
    ))] %>%
  #unique() %>%
  # remove missing
  .[deliv %in% 1 &
    !is.na(DLPARA)] %>%
  .[, `:=` (count = .N), by = list(BrthYear, epi, parity)] %>%
  .[, `:=` (total = .N), by = list(BrthYear, parity)] %>%
  .[, `:=` (rate = 100*count/total)]

tbl1 <- tbl %>%
  .[epi %in% 1 &
    parity %in% 1] %>%
  melt(measure.vars = c("rate", "total", "count")) %>%
  .[!tolower(variable) %in% "count"] %>%
  .[, `:=` (parity = fifelse(tolower(variable) %in% "total",
                               0, parity))] %>%
  .[, .(BrthYear, parity, value)] %>%
  unique() %>%
  dcast(parity ~ BrthYear, value.var = "value")

# coln <- names(tbl1)
# coln[1] <- ""

tbl1 <- as.data.frame(sapply(tbl1, sprintf, fmt = c("%.0f", "%1.1f%%")))
#names(tbl1) <- coln

tbl1[,1] <- c("Total vaginal deliveries", "Episiotomy (%)")

# add comma separator to total count
tbl1[1,-1] <- as.character(sapply(as.numeric(tbl1[1,-1]), scales::comma_format(accuracy = 1)))

temp <- as.data.frame(t(c("Primiparous", rep(NA, ncol(tbl1)-1))))
names(temp) <- names(tbl1)

tbl1 <- rbind(temp,tbl1)

tbl2 <- tbl %>%
  .[epi %in% 1 &
    parity %in% 2] %>%
  melt(measure.vars = c("rate", "total", "count")) %>%
  .[!tolower(variable) %in% "count"] %>%
  .[, `:=` (parity = fifelse(tolower(variable) %in% "total",
                               0, parity))] %>%
  .[, .(BrthYear, parity, value)] %>%
  unique() %>%
  dcast(parity ~ BrthYear, value.var = "value")

# coln <- names(tbl2)
# coln[1] <- ""

tbl2 <- as.data.frame(sapply(tbl2, sprintf, fmt = c("%.0f", "%1.1f%%")))
#names(tbl2) <- coln

tbl2[,1] <- c("Total vaginal deliveries", "Episiotomy (%)")

# add comma separator to total count
tbl2[1,-1] <- as.character(sapply(as.numeric(tbl2[1,-1]), scales::comma_format(accuracy = 1)))

temp <- as.data.frame(t(c("Multiparous", rep(NA, ncol(tbl2)-1))))
names(temp) <- names(tbl2)

tbl2 <- rbind(temp,tbl2)

tbl <- setDT(rbind(tbl1, tbl2))

#names(tbl)[1] <- "Parity"

reactable(tbl,
          rownames = FALSE,
          resizable = TRUE, 
          onClick = "expand",
          highlight = TRUE, 
          compact = TRUE,
          defaultPageSize = nrow(tbl),
          defaultColDef = colDef(align = "right",
                                 vAlign = "center",
                                 format = colFormat(locales = "en-CA", separators = TRUE),
                                 style = list(fontSize = 14),
                                 headerStyle = list(fontSize = 14),
                                 minWidth = 50,
                                 maxWidth = 180),
          rowStyle = function(index){
            if (index %in% (seq(1,nrow(tbl), by = 3))) list(color = "black",
                                         backgroundColor = "#e1e1e1",
                                         fontWeight = "bold",
                                         align = "center",
                                         borderBottom = "#dddddd")
          },
          columns = list(
            parity = colDef(name = "Parity",
                             align = "left",
                             width = 180)
          )
) %>%
  reactablefmtr::add_source("Note: An episiotomy is a mediolateral or midline incision made in the perineum during childbirth.",
                            font_size = 12,
                            #font_weight = "bold"
  )

# knitr::kable(tbl,
#              format = "html",
#              booktabs = TRUE) %>%
#   # add an hline before the total
# #kableExtra::row_spec(nrow(tbl) - 1, hline_after = TRUE) %>%  #kableExtra::row_spec(nrow(tbl), bold = TRUE) %>%
# kableExtra::row_spec(0, bold = TRUE) %>%
# #kableExtra::column_spec(1, border_right = TRUE, bold  = TRUE) %>%
# kableExtra::group_rows("Primiparous", start_row = 1, end_row = 2) %>%
# kableExtra::group_rows("Multiparous", start_row = 3, end_row = 4) %>%
# kableExtra::kable_styling(font_size = 12, full_width = FALSE,
#                           bootstrap_options = c("hover","condensed", "responsive")) %>%
# kableExtra::add_footnote(label = c("Note: An episiotomy is a mediolateral or midline incision made in the perineum during childbirth."))
```

## Obstetrical intervention by year, `r paste0(params$year1,"-",params$year2)` {#section-414}

```{r plot414, fig.cap= paste('Obstetrical intervention by year, Nova Scotia,', paste0(params$year1,"-",params$year2))}

dta1 <- dta[, .(BrthYear,
                LABOUR,
                FAILED_INDUCTION,
                AUGMENT,
                DMEPISIO,
                DMMETHOD,
                R010_00200, R010_00300, R010_00400, R010_00500, R010_00600, R010_00800, R010_00900, R010_01000,
                R011_00200, R011_00300, R011_00400, R011_00500, R011_00600, R011_00800, R011_00900, R011_01000,
                R012_00200, R012_00300, R012_00400, R012_00500, R012_00600, R012_00800, R012_00900, R012_01000,
                MEPIS, MEPIC, MIFUS, MPCEA, MSEDN, MSPIN, MPUDL, MGENA)] %>%
  .[, epi := rowSums(.SD, na.rm = TRUE), .SDcols = c("R010_00200","R010_00300", "R010_00400","R010_00500","R010_01000","R011_00200","R011_00300","R011_00400","R011_00500","R011_01000","R012_00200","R012_00300","R012_00400","R012_00500","R012_01000","MEPIS","MEPIC","MIFUS","MPCEA","MSEDN")] %>%
  .[, spi := rowSums(.SD, na.rm = TRUE), .SDcols = c("R010_00900","R010_01000","R011_00900","R011_01000","R012_00900","R012_01000","MSPIN","MSEDN")] %>%
  .[, dou := rowSums(.SD, na.rm = TRUE), .SDcols = c("R010_01000","R011_01000","R012_01000","MSEDN")] %>%
  .[, pud := rowSums(.SD, na.rm = TRUE), .SDcols = c("R010_00800","R011_00800","R012_00800","MPUDL")] %>%
  .[, gen := rowSums(.SD, na.rm = TRUE), .SDcols = c("R010_00600","R011_00600","R012_00600","MGENA")] %>%
  .[, `:=` (ind = fcase(
    tolower(LABOUR) %in% "i" | tolower(FAILED_INDUCTION) %in% "y", 1,
    default = 0
  ),
  augm = fcase(
    tolower(AUGMENT) %in% "y", 1,
    default = 0
  ),
  epis = fcase(
      tolower(DMEPISIO) %in% c("md","ml"), 1,
      default = 0
    ),
  op_del = fcase(
    toupper(DMMETHOD) %in% c("VAC","LOW","MID","CST"), 1,
    default = 0
  ),
  anes = fcase(
    epi >= 1 | spi >= 1 | dou >= 1 | pud >= 1 | gen >= 1, 1,
    default = 0
    )
)] %>%
  .[, interv := rowSums(.SD, na.rm = TRUE), .SDcols = c("ind","augm","epis","op_del","anes")] %>%
  .[, `:=` (obs_int = fcase(
    interv >= 1, 1,
    default = 0
  ))] %>%
  #unique() %>%
  # remove missing
  .[!is.na(LABOUR) |
    !is.na(FAILED_INDUCTION) |
    !is.na(AUGMENT) |
    !is.na(DMEPISIO) |
    !is.na(DMMETHOD)] %>%
  .[, `:=` (count = .N), by = list(BrthYear, obs_int)] %>%
  .[, `:=` (total = .N), by = list(BrthYear)] %>%
  .[, `:=` (rate = count/total)] %>%
  unique()

# age category using UNICODE codes
dta1 <- dta1[, `:=` (obs_int = factor(obs_int,
                                    levels = c(0,1),
                                    labels = c("No",
                                               "Yes")))]

## getting color for plotting

pal <- c("No" = "#1b9e77",
         "Yes" = "#d95f02")

plot <- ggplot(data = dta1[obs_int %in% "Yes"],
               aes(x = BrthYear,
                   y = rate,
                   group = obs_int,
                   colour = obs_int,
                   fill = obs_int,
                   #shape = obs_int,
                   text = paste(obs_int,
                               "<br>Year:", BrthYear,
                               "<br>Total deliveries", scales::comma(total, accuracy = 1),
                               "Obstetrical intervention (%):" , scales::percent(rate, accuracy = .1)
                               )
                   )
               ) +
  geom_point(aes(color = obs_int, fill = obs_int)) +
  geom_line(aes(color = obs_int, fill = obs_int)) +
# Specifying colors by age group
  scale_color_manual(name = "",
                     values = pal,
                     limits = names(pal)) +
  scale_fill_manual(name = "",
                    values = pal,
                    limits = names(pal)) +
  # scale_shape_manual(name = "",
  #                    values = c(15,16)) +
  # scale_x_date(date_breaks = "2 years",
  #              date_labels = "%Y") +
  scale_x_continuous(#trans = 'log2',
    breaks = pretty) +
  scale_y_continuous(#trans = 'log2',
    breaks = pretty,
    labels = scales::percent_format(accuracy = 1),
    limits = c(0, NA)) +
  labs(y = 'Proportion of deliveries', x = '') +
  theme_classic()

plotly::ggplotly(plot, tooltip = c("text")) %>%
  plotly::style(hoverlabel = list(
    bgcolor  = "black",
    bordercolor = "transparent",
    font = list(
      color = "white",
      size = 14,
      face = "bold"
    )
  )) %>%
  plotly::layout(legend = list(orientation = "h", y = -0.25, x = 0.5,
                               xanchor = "center"),
                 font = list(face = "bold", size = 10),
                 showlegend = FALSE)
```

```{r tbl414}
tbl <-  dta[, .(BrthYear,
                LABOUR,
                FAILED_INDUCTION,
                AUGMENT,
                DMEPISIO,
                DMMETHOD,
                R010_00200, R010_00300, R010_00400, R010_00500, R010_00600, R010_00800, R010_00900, R010_01000,
                R011_00200, R011_00300, R011_00400, R011_00500, R011_00600, R011_00800, R011_00900, R011_01000,
                R012_00200, R012_00300, R012_00400, R012_00500, R012_00600, R012_00800, R012_00900, R012_01000,
                MEPIS, MEPIC, MIFUS, MPCEA, MSEDN, MSPIN, MPUDL, MGENA)] %>%
  .[, epi := rowSums(.SD, na.rm = TRUE), .SDcols = c("R010_00200","R010_00300", "R010_00400","R010_00500","R010_01000","R011_00200","R011_00300","R011_00400","R011_00500","R011_01000","R012_00200","R012_00300","R012_00400","R012_00500","R012_01000","MEPIS","MEPIC","MIFUS","MPCEA","MSEDN")] %>%
  .[, spi := rowSums(.SD, na.rm = TRUE), .SDcols = c("R010_00900","R010_01000","R011_00900","R011_01000","R012_00900","R012_01000","MSPIN","MSEDN")] %>%
  .[, dou := rowSums(.SD, na.rm = TRUE), .SDcols = c("R010_01000","R011_01000","R012_01000","MSEDN")] %>%
  .[, pud := rowSums(.SD, na.rm = TRUE), .SDcols = c("R010_00800","R011_00800","R012_00800","MPUDL")] %>%
  .[, gen := rowSums(.SD, na.rm = TRUE), .SDcols = c("R010_00600","R011_00600","R012_00600","MGENA")] %>%
  .[, `:=` (ind = fcase(
    tolower(LABOUR) %in% "i" | tolower(FAILED_INDUCTION) %in% "y", 1,
    default = 0
  ),
  augm = fcase(
    tolower(AUGMENT) %in% "y", 1,
    default = 0
  ),
  epis = fcase(
      tolower(DMEPISIO) %in% c("md","ml"), 1,
      default = 0
    ),
  op_del = fcase(
    toupper(DMMETHOD) %in% c("VAC","LOW","MID","CST"), 1,
    default = 0
  ),
  anes = fcase(
    epi >= 1 | spi >= 1 | dou >= 1 | pud >= 1 | gen >= 1, 1,
    default = 0
    )
)] %>%
  .[, interv := rowSums(.SD, na.rm = TRUE), .SDcols = c("ind","augm","epis","op_del","anes")] %>%
  .[, `:=` (obs_int = fcase(
    interv >= 1, 1,
    default = 0
  ))] %>%
  #unique() %>%
  # remove missing
  .[!is.na(LABOUR) |
    !is.na(FAILED_INDUCTION) |
    !is.na(AUGMENT) |
    !is.na(DMEPISIO) |
    !is.na(DMMETHOD)] %>%
  .[, `:=` (count = .N), by = list(BrthYear, obs_int)] %>%
  .[, `:=` (total = .N), by = list(BrthYear)] %>%
  .[, `:=` (rate = 100*count/total)] %>%
  melt(measure.vars = c("rate", "total", "count")) %>%
  .[!tolower(variable) %in% "count" &
    !obs_int %in% 0] %>%
  .[, `:=` (obs_int = fifelse(tolower(variable) %in% "total",
                               0, obs_int))] %>%
  .[, .(BrthYear, obs_int, value)] %>%
  unique() %>%
  dcast(obs_int ~ BrthYear, value.var = "value")

# coln <- names(tbl)
# coln[1] <- ""

tbl <- as.data.frame(sapply(tbl, sprintf, fmt = c("%.0f", "%1.1f%%")))
#names(tbl) <- coln

tbl[,1] <- c("Total deliveries", "Obstetrical intervention (%)")

# add comma separator to total count
tbl[1,-1] <- as.character(sapply(as.numeric(tbl[1,-1]), scales::comma_format(accuracy = 1)))

reactable(tbl,
          rownames = FALSE,
          resizable = TRUE, 
          onClick = "expand",
          highlight = TRUE, 
          compact = TRUE,
          defaultPageSize = nrow(tbl),
          defaultColDef = colDef(align = "right",
                                 vAlign = "center",
                                 format = colFormat(locales = "en-CA", separators = TRUE),
                                 style = list(fontSize = 14),
                                 headerStyle = list(fontSize = 14),
                                 minWidth = 50,
                                 maxWidth = 180),
          # rowStyle = function(index){
          #   if (index == nrow(tbl)) list(color = "black",
          #                                backgroundColor = "#e1e1e1",
          #                                fontWeight = "bold",
          #                                align = "center",
          #                                borderBottom = "#dddddd")
          # },
          columns = list(
            obs_int = colDef(name = "",
                             align = "left",
                             width = 180)
          )
) %>%
  reactablefmtr::add_source("Note: Obstetrical intervention includes the use of any of: induction, medical augmentation, anesthesia, Caesarean delivery, vaginal delivery involving the use of forceps and/or vacuum, or episiotomy.",
                            font_size = 12,
                            #font_weight = "bold"
  )

# knitr::kable(tbl,
#              format = "html",
#              booktabs = TRUE) %>%
#   # add an hline before the total
# #kableExtra::row_spec(nrow(tbl) - 1, hline_after = TRUE) %>%  #kableExtra::row_spec(nrow(tbl), bold = TRUE) %>%
# kableExtra::row_spec(0, bold = TRUE) %>%
# #kableExtra::column_spec(1, border_right = TRUE, bold  = TRUE) %>%
# kableExtra::kable_styling(font_size = 12, full_width = FALSE,
#                           bootstrap_options = c("hover","condensed", "responsive")) %>%
# kableExtra::add_footnote(label = c("Note: Obstetrical intervention includes the use of any of: induction, medical augmentation, anesthesia, Caesarean delivery, vaginal delivery involving the use of forceps and/or vacuum, or episiotomy."))
```