---
params:
  year1: 2010
  year2: 2019
fig.caption: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE, 
  message = FALSE, 
  warning = FALSE,
  fig.pos = "h", 
  out.extra = "",
  out.width = '90%',
  fig.align = 'center',
  eval.after = "fig.cap"
  #cache = TRUE
  )

pckgs <- c('tidyverse', 'knitr', 'scales','janitor',
           'kableExtra', 'data.table','reactablefmtr',
           'forcats','lubridate', 'anytime','reactable',
           'htmltools', 'plotly','purrr')
# for (pp in `pckgs`) { if (!require(pp)) install.packages(pp); library(pp, character.only = T)  }

if (!require("pacman")) install.packages("pacman")
pacman::p_load(pckgs, character.only = TRUE)
## data input

# add data
dta <- readr::read_csv("data/NSAtleePD.csv") %>%
  setDT() %>%
  .[data.table::between(BrthYear, params$year1, params$year2) &
    ProvNo %in% 12 &
    (BIRTHWT >= 500 | GA_BEST >= 20), ]
```

<!-- # (PART) Deliveries and Births {-} -->
  
# Maternal Health Outcomes {#section-5}
  
## Gestational diabetes by year, Nova Scotia, `r paste0(params$year1,"-",params$year2)` {#section-51}
  
```{r plot51, fig.cap= paste('Gestational diabetes by year, Nova Scotia,', paste0(params$year1,"-",params$year2))}

dta1 <- dta[, .(BrthYear,
                # pre-pregnancy diabetes
                 R014_00900,
                 R014_01000,
                 R014_01100,
                 R014_01200,
                 MO240,
                 MO241,
                 MO242,
                 MO243,
                 MO245, MO24501, MO24502, MO24504,
                 MO246, MO24601, MO24602, MO24604,
                 MO247, MO24701, MO24704,
                 ME10, ME11, ME13, ME14,
                 MH360,
                 MG632,
                 IP701,
                # gestational diabetes
                 R014_01300,
                 R014_01400,
                 MO244, MO24401, MO24404, MO24409,
                 MO248, MO24801, MO24802, MO24803, MO24804,
                 MO249,
                 IP700,
               CONTCTID)] %>%
  unique() %>%
  .[, idx_244 := rowSums(.SD, na.rm = TRUE), .SDcols = names(dta)[which(grepl("MO2440.", names(dta)))]] %>%
  .[, idx_245 := rowSums(.SD, na.rm = TRUE), .SDcols = names(dta)[which(grepl("MO2450.", names(dta)))]] %>%
  .[, idx_246 := rowSums(.SD, na.rm = TRUE), .SDcols = names(dta)[which(grepl("MO2460.", names(dta)))]] %>%
  .[, idx_247 := rowSums(.SD, na.rm = TRUE), .SDcols = names(dta)[which(grepl("MO2470.", names(dta)))]] %>%
  .[, idx_248 := rowSums(.SD, na.rm = TRUE), .SDcols = names(dta)[which(grepl("MO2480.", names(dta)))]] %>%
  .[, `:=` (predm_nsapd = fcase(
    R014_00900 >=1 | R014_01000 >=1 | 
    R014_01100 >=1 | R014_01200 >=1, 1,
   default = 0
    ),
   predm_icd = fcase(
    MO240 >=1 | MO241 >=1 | MO242 >=1 | MO243 >=1 |
    MO245 >=1 | idx_245 >=1 |
    MO246 >=1 | idx_246 >=1 |
    MO247 >=1 | idx_247 >=1 |
    ME10 >=1 | ME11 >=1 | ME13 >=1 | ME14 >=1 |
    MH360 >=1 | MG632 >=1 | IP701  >=1, 1,
   default = 0
    ),
   gdm_nsapd = fcase(
    R014_01300 >=1 | R014_01400 >=1, 1,
   default = 0
    ),
   gdm_icd = fcase(
    MO244 >=1 | MO249 >=1 | MO248 >=1 |
    idx_244 >=1 | idx_248 >=1 |
    IP700  >=1, 1,
   default = 0
    ))] %>%
  .[, `:=` (prexdiab = fcase(
    predm_nsapd %in% 1 | predm_icd %in% 1, 1,
   default = 0
    ),
   gestdiab = fcase(
     gdm_nsapd %in% 1 | gdm_icd %in% 1, 1,
     default = 0
   ))] %>%
  #unique() %>%
  # among women without pre-existing diabetes
  .[prexdiab %in% 0] %>%
  .[, `:=` (count = .N), by = list(BrthYear, gestdiab)] %>%
  .[, `:=` (total = .N), by = list(BrthYear)] %>%
  .[, `:=` (rate = count/total)] %>%
  .[,.(BrthYear, rate, total, gestdiab)] %>%
  unique()

# age category using UNICODE codes
dta1 <- dta1[, `:=` (gestdiab = factor(gestdiab,
                                  levels = c(0,1),
                                  labels = c("No",
                                             "Gestational diabetes")))]

## getting color for plotting

pal <- c("No" = "#1b9e77",
         "Gestational diabetes" = "#d95f02")

plot <- ggplot(data = dta1[!gestdiab %in% c("No")],
               aes(x = BrthYear,
                   y = rate,
                   group = gestdiab,
                   colour = gestdiab,
                   fill = gestdiab,
                   #shape = gestdiab,
                   text =paste(paste0(gestdiab," - ", BrthYear),
                               "<br>Total deliveries:", scales::comma(total, accuracy = 1),
                               paste0("<br>",gestdiab, " (%):") , scales::percent(rate, accuracy = .1)
                   )
               )
) +
  geom_point(aes(color = gestdiab, fill = gestdiab)) +
  geom_line(aes(color = gestdiab, fill = gestdiab)) +
  # Specifying colors by age group
  scale_color_manual(name = "",
                     values = pal,
                     limits = names(pal)) +
  scale_fill_manual(name = "",
                    values = pal,
                    limits = names(pal)) +
  # scale_shape_manual(name = "",
  #                    values = c(21, 22, 23)) +
  # scale_x_date(date_breaks = "2 years",
  #              date_labels = "%Y") +
  scale_x_continuous(#trans = 'log2',
    breaks = pretty) +
  scale_y_continuous(#trans = 'log2',
    breaks = pretty,
    labels = scales::percent_format(accuracy = 1),
    limits = c(0, NA)) +
  labs(y = 'Proportion of deliveries', x = '') +
  theme_classic()

plotly::ggplotly(plot, tooltip = c("text")) %>%
  plotly::style(hoverlabel = list(
    bgcolor  = "black",
    bordercolor = "transparent",
    align = "left",
    font = list(
      color = "white",
      size = 14,
      face = "bold"
    )
  )) %>%
  plotly::layout(legend = list(orientation = "h", y = -0.25, x = 0.5,
                               xanchor = "center"),
                 font = list(face = "bold", size = 10),
                 showlegend = FALSE)
```

```{r tbl51}
tbl <- dta[, .(BrthYear,
                # pre-pregnancy diabetes
                 R014_00900,
                 R014_01000,
                 R014_01100,
                 R014_01200,
                 MO240,
                 MO241,
                 MO242,
                 MO243,
                 MO245, MO24501, MO24502, MO24504,
                 MO246, MO24601, MO24602, MO24604,
                 MO247, MO24701, MO24704,
                 ME10, ME11, ME13, ME14,
                 MH360,
                 MG632,
                 IP701,
                # gestational diabetes
                 R014_01300,
                 R014_01400,
                 MO244, MO24401, MO24404, MO24409,
                 MO248, MO24801, MO24802, MO24803, MO24804,
                 MO249,
                 IP700,
               CONTCTID)] %>%
  unique() %>%
  .[, idx_244 := rowSums(.SD, na.rm = TRUE), .SDcols = names(dta)[which(grepl("MO2440.", names(dta)))]] %>%
  .[, idx_245 := rowSums(.SD, na.rm = TRUE), .SDcols = names(dta)[which(grepl("MO2450.", names(dta)))]] %>%
  .[, idx_246 := rowSums(.SD, na.rm = TRUE), .SDcols = names(dta)[which(grepl("MO2460.", names(dta)))]] %>%
  .[, idx_247 := rowSums(.SD, na.rm = TRUE), .SDcols = names(dta)[which(grepl("MO2470.", names(dta)))]] %>%
  .[, idx_248 := rowSums(.SD, na.rm = TRUE), .SDcols = names(dta)[which(grepl("MO2480.", names(dta)))]] %>%
  .[, `:=` (predm_nsapd = fcase(
    R014_00900 >=1 | R014_01000 >=1 | 
    R014_01100 >=1 | R014_01200 >=1, 1,
   default = 0
    ),
   predm_icd = fcase(
    MO240 >=1 | MO241 >=1 | MO242 >=1 | MO243 >=1 |
    MO245 >=1 | idx_245 >=1 |
    MO246 >=1 | idx_246 >=1 |
    MO247 >=1 | idx_247 >=1 |
    ME10 >=1 | ME11 >=1 | ME13 >=1 | ME14 >=1 |
    MH360 >=1 | MG632 >=1 | IP701  >=1, 1,
   default = 0
    ),
   gdm_nsapd = fcase(
    R014_01300 >=1 | R014_01400 >=1, 1,
   default = 0
    ),
   gdm_icd = fcase(
    MO244 >=1 | MO249 >=1 | MO248 >=1 |
    idx_244 >=1 | idx_248 >=1 |
    IP700  >=1, 1,
   default = 0
    ))] %>%
  .[, `:=` (prexdiab = fcase(
    predm_nsapd %in% 1 | predm_icd %in% 1, 1,
   default = 0
    ),
   gestdiab = fcase(
     gdm_nsapd %in% 1 | gdm_icd %in% 1, 1,
     default = 0
   ))] %>%
  #unique() %>%
  # among women without pre-existing diabetes
  .[prexdiab %in% 0] %>%
  .[, `:=` (count = .N), by = list(BrthYear, gestdiab)] %>%
  .[, `:=` (total = .N), by = list(BrthYear)] %>%
  .[, `:=` (rate = 100*count/total)] %>%
  melt(measure.vars = c("rate", "total", "count")) %>%
  .[!tolower(variable) %in% "count" &
    !gestdiab %in% 0] %>%
  .[, `:=` (gestdiab = fifelse(tolower(variable) %in% "total",
                          -1, gestdiab))] %>%
  .[, .(BrthYear, gestdiab, value)] %>%
  unique() %>%
  dcast(gestdiab ~ BrthYear, value.var = "value")

# coln <- names(tbl)
# coln[1] <- ""

tbl <- as.data.frame(sapply(tbl, sprintf, fmt = c("%.0f", "%1.1f%%")))
#names(tbl) <- coln

tbl[,1] <- c("Total deliveries\u1d43", "Gestational diabetes (%)")

# add comma separator to total count
tbl[1,-1] <- as.character(sapply(as.numeric(tbl[1,-1]), scales::comma_format(accuracy = 1)))

reactable(tbl,
          rownames = FALSE,
          resizable = TRUE, 
          onClick = "expand",
          highlight = TRUE, 
          compact = TRUE,
          defaultPageSize = nrow(tbl),
          defaultColDef = colDef(align = "right",
                                 vAlign = "center",
                                 format = colFormat(locales = "en-CA", separators = TRUE),
                                 style = list(fontSize = 14),
                                 headerStyle = list(fontSize = 14),
                                 minWidth = 50,
                                 maxWidth = 180),
          # rowStyle = function(index){
          #   if (index == nrow(tbl)) list(color = "black",
          #                                backgroundColor = "#e1e1e1",
          #                                fontWeight = "bold",
          #                                align = "center",
          #                                borderBottom = "#dddddd")
          # },
          columns = list(
            gestdiab = colDef(name = "",
                             align = "left",
                             width = 180)
          )
) %>%
  reactablefmtr::add_source("\u1d43 Among women without pre-existing diabetes.",
                            font_size = 12,
                            #font_weight = "bold"
  ) %>%
  reactablefmtr::add_source("Note: Diabetes mellitus first detected in pregnancy as recorded in the medical record. Please note that the criteria for the diagnosis of gestational diabetes were revised by Diabetes Canada (formerly the Canadian Diabetes Association) in 2013. Therefore, the rates of gestational diabetes were expected to increase as the new criteria are adopted across Nova Scotia, starting approximately in late 2014.",
                            font_size = 12,
                            #font_weight = "bold"
  )

# knitr::kable(tbl,
#              format = "html",
#              booktabs = TRUE) %>%
#   # add an hline before the total
#   #kableExtra::row_spec(nrow(tbl) - 1, hline_after = TRUE) %>%  #kableExtra::row_spec(nrow(tbl), bold = TRUE) %>%
# kableExtra::row_spec(0, bold = TRUE) %>%
# #kableExtra::column_spec(1, border_right = TRUE, bold  = TRUE) %>%
# kableExtra::kable_styling(font_size = 12, full_width = FALSE,
#                             bootstrap_options = c("hover","condensed", "responsive")) %>%
#   kableExtra::add_footnote(label = c("Among women without pre-existing diabetes.",
#                                      "Note: Diabetes mellitus first detected in pregnancy as recorded in the medical record. Please note that the criteria for the diagnosis of gestational diabetes were revised by Diabetes Canada (formerly the Canadian Diabetes Association) in 2013. Therefore, the rates of gestational diabetes were expected to increase as the new criteria are adopted across Nova Scotia, starting approximately in late 2014."))
```

## Gestational hypertension by year, Nova Scotia, `r paste0(params$year1,"-",params$year2)` {#section-52}
  
```{r plot52, fig.cap= paste('Gestational hypertension by year, Nova Scotia,', paste0(params$year1,"-",params$year2))}

dta1 <- dta[, .(BrthYear,
                # pre-eclampsia
                R014_00600, R014_00800,
                MO14, MO11, MO15,
                # gh any
                MO13,
                R014_00500, R014_01900,
                # pre htn
                R014_00700, R014_00800, 
                MO10,
                MI10, MI11, MI12, MI13, MI15, MCHTD,
               CONTCTID)] %>%
  unique() %>%
  .[, `:=` (preeclamp = fcase(
    R014_00600 >=1 | R014_00800 >=1 |
    MO14 >=1 | MO11 >=1 | MO15 >=1, 1,
   default = 0
    ),
   ghany = fcase(
    MO13 >=1 | MO14 >=1 | MO15 >=1 |
    R014_00500 >=1 | R014_01900 >=1, 1,
   default = 0
   ),
   prehtn = fcase(
    R014_00700 >=1 | R014_00800 >=1 | 
    MO10 >=1 | MO11 >=1 |
    MI10 >=1 | MI11 >=1 | MI12 >=1 | MI13 >=1 | MI15 >=1 | MCHTD >=1, 1,
   default = 0
    ))] %>%
  .[, `:=` (ghinsig = fcase(
    ghany >=1 & !preeclamp >=1 & !prehtn >=1 , 1,
   default = 0
    ))] %>%
  .[, `:=` (ghcat = fcase(
    !preeclamp >=1 & !ghinsig >=1, 1,
    !preeclamp >=1 & ghinsig >=1, 2,
    preeclamp >=1 , 3,
   default = 0
    ))] %>%
  # remove prehtn
  .[!prehtn >= 1] %>%
  #unique() %>%
  .[, `:=` (count = .N), by = list(BrthYear, ghcat)] %>%
  .[, `:=` (total = .N), by = list(BrthYear)] %>%
  .[, `:=` (rate = count/total)] %>%
  .[!ghcat %in% 1, .(BrthYear, ghcat, count, total, rate)] %>% 
  unique()

temp <- unique(dta1) %>%
  .[,`:=` (ghcat = 1,
           count = sum(count),
           total = total,
           rate = sum(rate)), by = list(BrthYear)]


dta1 <- unique(rbind(dta1, temp))[,
            `:=` (ghcat = factor(ghcat,
                                   levels = c(1,2,3),
                                   labels = c("GH-total",
                                              "GH-no sig proteinuria",
                                              "GH-sig proteinuria")))][,
              .(BrthYear, rate, total, ghcat)]

## getting color for plotting

pal <- c("GH-total" = "#1b9e77",
         "GH-sig proteinuria" = "#d95f02",
         "GH-no sig proteinuria" = "#7570b3")

plot <- ggplot(data = dta1,
               aes(x = BrthYear,
                   y = rate,
                   group = ghcat,
                   colour = ghcat,
                   fill = ghcat,
                   shape = ghcat,
                   text =paste(paste0(ghcat," - ", BrthYear),
                               "<br>Total deliveries:", scales::comma(total, accuracy = 1),
                               paste0("<br>",ghcat, " (%):") , scales::percent(rate, accuracy = .1)
                   )
               )
) +
  geom_point(aes(color = ghcat, fill = ghcat, shape = ghcat)) +
  geom_line(aes(color = ghcat, fill = ghcat)) +
  # Specifying colors by age group
  scale_color_manual(name = "",
                     values = pal,
                     limits = names(pal)) +
  scale_fill_manual(name = "",
                    values = pal,
                    limits = names(pal)) +
  scale_shape_manual(name = "",
                     values = c(21, 22, 23)) +
  # scale_x_date(date_breaks = "2 years",
  #              date_labels = "%Y") +
  scale_x_continuous(#trans = 'log2',
    breaks = pretty) +
  scale_y_continuous(#trans = 'log2',
    breaks = pretty,
    labels = scales::percent_format(accuracy = 1),
    limits = c(0, NA)) +
  labs(y = 'Proportion of deliveries', x = '') +
  theme_classic()

plotly::ggplotly(plot, tooltip = c("text")) %>%
  plotly::style(hoverlabel = list(
    bgcolor  = "black",
    bordercolor = "transparent",
    align = "left",
    font = list(
      color = "white",
      size = 14,
      face = "bold"
    )
  )) %>%
  plotly::layout(legend = list(orientation = "h", y = -0.25, x = 0.5,
                               xanchor = "center"),
                 font = list(face = "bold", size = 10),
                 showlegend = TRUE)
  # plotly::config(
  #   toImageButtonOptions = list(
  #   format = "jpeg"
  # ))
```

```{r tbl52}
tbl <- dta[, .(BrthYear,
                # pre-eclampsia
                R014_00600, R014_00800, R014_01800,
                MO14, MO11, MO15,
                # gh any
                MO13,
                R014_00500, R014_01900,
                # pre htn
                R014_00700, R014_00800, 
                MO10,
                MI10, MI11, MI12, MI13, MI15, MCHTD,
               CONTCTID)] %>%
  unique() %>%
  .[, `:=` (preeclamp = fcase(
    R014_00600 >=1 | R014_00800 >=1 |
    MO14 >=1 | MO11 >=1 | MO15 >=1, 1,
   default = 0
    ),
   ghany = fcase(
    MO13 >=1 | MO14 >=1 | MO15 >=1 |
    R014_00500 >=1 | R014_01900 >=1, 1,
   default = 0
   ),
   prehtn = fcase(
    R014_00700 >=1 | R014_00800 >=1 | 
    MO10 >=1 | MO11 >=1 |
    MI10 >=1 | MI11 >=1 | MI12 >=1 | MI13 >=1 | MI15 >=1 | MCHTD >=1, 1,
   default = 0
    ))] %>%
  .[, `:=` (ghinsig = fcase(
    ghany >=1 & !preeclamp >=1 & !prehtn >=1 , 1,
   default = 0
    ))] %>%
  .[, `:=` (ghcat = fcase(
    !preeclamp >=1 & !ghinsig >=1, 1,
    !preeclamp >=1 & ghinsig >=1, 2,
    preeclamp >=1 , 3,
   default = 0
    ))] %>%
  # remove prehtn
  .[!prehtn >= 1] %>%
  #unique() %>%
  .[, `:=` (count = .N), by = list(BrthYear, ghcat)] %>%
  .[, `:=` (total = .N), by = list(BrthYear)] %>%
  .[, `:=` (rate = 100*count/total)] %>%
  .[!ghcat %in% 1, .(BrthYear, ghcat, count, total, rate)] %>% 
  unique()

temp <- unique(tbl) %>%
  .[,`:=` (ghcat = 4,
           count = sum(count),
           total = total,
           rate = sum(rate)), by = list(BrthYear)]


tbl <- unique(rbind(tbl, temp)) %>%
  melt(measure.vars = c("count", "total", "rate")) %>%
  .[!tolower(variable) %in% "count"] %>%
  .[, `:=` (ghcat = fifelse(tolower(variable) %in% "total",
                          -1, ghcat))] %>%
  .[, .(BrthYear, ghcat, value)] %>%
  unique() %>%
  dcast(ghcat ~ BrthYear, value.var = "value")

# coln <- names(tbl)
# coln[1] <- ""

tbl <- as.data.frame(sapply(tbl, sprintf, fmt = c("%.0f", "%1.1f%%", "%1.1f%%", "%1.1f%%")))

#names(tbl) <- coln

tbl[,1] <- c("Total deliveries\u1d43", "GH-no sig proteinuria", "GH-sig proteinuria", "GH-total")

# add comma separator to total count
tbl[1,-1] <- as.character(sapply(as.numeric(tbl[1,-1]), scales::comma_format(accuracy = 1)))

reactable(tbl,
          rownames = FALSE,
          resizable = TRUE, 
          onClick = "expand",
          highlight = TRUE, 
          compact = TRUE,
          defaultPageSize = nrow(tbl),
          defaultColDef = colDef(align = "right",
                                 vAlign = "center",
                                 format = colFormat(locales = "en-CA", separators = TRUE),
                                 style = list(fontSize = 14),
                                 headerStyle = list(fontSize = 14),
                                 minWidth = 50,
                                 maxWidth = 180),
          rowStyle = function(index){
            if (index == nrow(tbl)) list(color = "black",
                                         backgroundColor = "#e1e1e1",
                                         fontWeight = "bold",
                                         align = "center",
                                         borderBottom = "#dddddd")
          },
          columns = list(
            ghcat = colDef(name = "",
                             align = "left",
                             width = 180)
          )
) %>%
  reactablefmtr::add_source("\u1d43 Among women without pre-existing hypertension.",
                            font_size = 12,
                            #font_weight = "bold"
  ) %>%
  reactablefmtr::add_source("Note: Gestational hypertension is hypertension that is first detected after the 20th week of gestation. Gestational hypertension with significant proteinuria includes those cases denoted as such; severe pre-eclampsia; HELLP syndrome (Hemolysis, Elevated Liver Enzymes, Low Platelets); and eclampsia.",
                            font_size = 12,
                            #font_weight = "bold"
  )

# knitr::kable(tbl,
#              format = "html",
#              booktabs = TRUE) %>%
#   # add an hline before the total
#   #kableExtra::row_spec(nrow(tbl) - 1, hline_after = TRUE) %>%  #kableExtra::row_spec(nrow(tbl), bold = TRUE) %>%
# kableExtra::row_spec(0, bold = TRUE) %>%
# #kableExtra::column_spec(1, border_right = TRUE, bold  = TRUE) %>%
# kableExtra::kable_styling(font_size = 12, full_width = FALSE,
#                             bootstrap_options = c("hover","condensed", "responsive")) %>%
#   kableExtra::add_footnote(label = c("Among women without pre-existing hypertension.",
#                                      "Note: Gestational hypertension is hypertension that is first detected after the 20th week of gestation. Gestational hypertension with significant proteinuria includes those cases denoted as such; severe pre-eclampsia; HELLP syndrome (Hemolysis, Elevated Liver Enzymes, Low Platelets); and eclampsia."))
```

## Pre-eclampsia by year, Nova Scotia, `r paste0(params$year1,"-",params$year2)` {#section-53}
  
```{r plot53, fig.cap= paste('Pre-eclampsia by year, Nova Scotia,', paste0(params$year1,"-",params$year2))}

dta1 <- tbl <- dta[, .(BrthYear,
                # pre-eclampsia
                R014_00600, R014_00800, R014_01800,
                MO14, MO11, MO15,
                # gh any
                MO13,
                R014_00500, R014_01900,
                # pre htn
                R014_00700, R014_00800, 
                MO10,
                MI10, MI11, MI12, MI13, MI15, MCHTD,
               CONTCTID)] %>%
  unique() %>%
  .[, `:=` (preeclamp = fcase(
    R014_00600 >=1 | R014_00800 >=1 |
      R014_01800 >=1 | R014_01900 >=1 |
    MO14 >=1 | MO11 >=1 | MO15 >=1, 1,
   default = 0
    ))] %>%
  #unique() %>%
  .[, `:=` (count = .N), by = list(BrthYear, preeclamp)] %>%
  .[, `:=` (total = .N), by = list(BrthYear)] %>%
  .[, `:=` (rate = count/total)] %>%
  .[,.(BrthYear, rate, total, preeclamp)] %>%
  unique()


dta1 <- dta1[, `:=` (preeclamp = factor(preeclamp,
                                   levels = c(0,1),
                                   labels = c("No",
                                              "Pre-eclampsia")))]

## getting color for plotting

pal <- c("No" = "#1b9e77",
         "Pre-eclampsia" = "#d95f02")

plot <- ggplot(data = dta1[!tolower(preeclamp) %in% "no"],
               aes(x = BrthYear,
                   y = rate,
                   group = preeclamp,
                   colour = preeclamp,
                   fill = preeclamp,
                   #shape = preeclamp,
                   text =paste(paste0(preeclamp," - ", BrthYear),
                               "<br>Total deliveries:", scales::comma(total, accuracy = 1),
                               paste0("<br>",preeclamp," (%):") , scales::percent(rate, accuracy = .1)
                   )
               )
) +
  geom_point(aes(color = preeclamp, fill = preeclamp)) +
  geom_line(aes(color = preeclamp, fill = preeclamp)) +
  # Specifying colors by age group
  scale_color_manual(name = "",
                     values = pal,
                     limits = names(pal)) +
  scale_fill_manual(name = "",
                    values = pal,
                    limits = names(pal)) +
  # scale_shape_manual(name = "",
  #                    values = c(15, 16, 17)) +
  # scale_x_date(date_breaks = "2 years",
  #              date_labels = "%Y") +
  scale_x_continuous(#trans = 'log2',
    breaks = pretty) +
  scale_y_continuous(#trans = 'log2',
    breaks = pretty,
    labels = scales::percent_format(accuracy = 1),
    limits = c(0, NA)) +
  labs(y = 'Proportion of deliveries', x = '') +
  theme_classic()

plotly::ggplotly(plot, tooltip = c("text")) %>%
  plotly::style(hoverlabel = list(
    bgcolor  = "black",
    bordercolor = "transparent",
    align = "left",
    font = list(
      color = "white",
      size = 14,
      face = "bold"
    )
  )) %>%
  plotly::layout(legend = list(orientation = "h", y = -0.25, x = 0.5,
                               xanchor = "center"),
                 font = list(face = "bold", size = 10),
                 showlegend = FALSE)
```

```{r tbl53}
tbl <- dta[, .(BrthYear,
                # pre-eclampsia
                R014_00600, R014_00800, R014_01800,
                MO14, MO11, MO15,
                # gh any
                MO13,
                R014_00500, R014_01900,
                # pre htn
                R014_00700, R014_00800, 
                MO10,
                MI10, MI11, MI12, MI13, MI15, MCHTD,
               CONTCTID)] %>%
  unique() %>%
  .[, `:=` (preeclamp = fcase(
    R014_00600 >=1 | R014_00800 >=1 |
      R014_01800 >=1 | R014_01900 >=1 |
    MO14 >=1 | MO11 >=1 | MO15 >=1, 1,
   default = 0
    ))] %>%
  #unique() %>%
  .[, `:=` (count = .N), by = list(BrthYear, preeclamp)] %>%
  .[, `:=` (total = .N), by = list(BrthYear)] %>%
  .[, `:=` (rate = 100*count/total)] %>%
  unique() %>%
  melt(measure.vars = c("count", "total", "rate")) %>%
  .[!tolower(variable) %in% "count" &
    !preeclamp %in% 0] %>%
  .[, `:=` (preeclamp = fifelse(tolower(variable) %in% "total",
                          -1, preeclamp))] %>%
  .[, .(BrthYear, preeclamp, value)] %>%
  unique() %>%
  dcast(preeclamp ~ BrthYear, value.var = "value")

# coln <- names(tbl)
# coln[1] <- ""

tbl <- as.data.frame(sapply(tbl, sprintf, fmt = c("%.0f", "%1.1f%%")))

#names(tbl) <- coln

tbl[,1] <- c("Total deliveries", "Pre-eclampsia (%)")

# add comma separator to total count
tbl[1,-1] <- as.character(sapply(as.numeric(tbl[1,-1]), scales::comma_format(accuracy = 1)))

reactable(tbl,
          rownames = FALSE,
          resizable = TRUE, 
          onClick = "expand",
          highlight = TRUE, 
          compact = TRUE,
          defaultPageSize = nrow(tbl),
          defaultColDef = colDef(align = "right",
                                 vAlign = "center",
                                 format = colFormat(locales = "en-CA", separators = TRUE),
                                 style = list(fontSize = 14),
                                 headerStyle = list(fontSize = 14),
                                 minWidth = 50,
                                 maxWidth = 180),
          # rowStyle = function(index){
          #   if (index == nrow(tbl)) list(color = "black",
          #                                backgroundColor = "#e1e1e1",
          #                                fontWeight = "bold",
          #                                align = "center",
          #                                borderBottom = "#dddddd")
          # },
          columns = list(
            preeclamp = colDef(name = "",
                             align = "left",
                             width = 180)
          )
) %>%
  reactablefmtr::add_source("Note: Pre-eclampsia includes women coded as having gestational hypertension with significant proteinuria, moderate or severe pre-eclampsia, HELLP syndrome (Hemolysis, Elevated Liver Enzymes, Low Platelets), eclampsia, or pre-existing hypertension with superimposed proteinuria.",
                            font_size = 12,
                            #font_weight = "bold"
  )

# knitr::kable(tbl,
#              format = "html",
#              booktabs = TRUE) %>%
#   # add an hline before the total
#   #kableExtra::row_spec(nrow(tbl) - 1, hline_after = TRUE) %>%  #kableExtra::row_spec(nrow(tbl), bold = TRUE) %>%
# kableExtra::row_spec(0, bold = TRUE) %>%
# #kableExtra::column_spec(1, border_right = TRUE, bold  = TRUE) %>%
# kableExtra::kable_styling(font_size = 12, full_width = FALSE,
#                             bootstrap_options = c("hover","condensed", "responsive")) %>%
#   kableExtra::add_footnote(label = c("Note: Pre-eclampsia includes women coded as having gestational hypertension with significant proteinuria, moderate or severe pre-eclampsia, HELLP syndrome (Hemolysis, Elevated Liver Enzymes, Low Platelets), eclampsia, or pre-existing hypertension with superimposed proteinuria."))
```

## Placenta previa by year, Nova Scotia, `r paste0(params$year1,"-",params$year2)` {#section-54}
  
```{r plot54, fig.cap= paste('Placenta previa by year, Nova Scotia,', paste0(params$year1,"-",params$year2))}

dta1 <- dta[, .(BrthYear,
        MO44,
        CONTCTID)] %>%
  unique() %>%
  .[, `:=` (plaprev = fcase(
    MO44 >=1, 1, 
    default = 0
  ))] %>%
  #unique() %>%
  .[, `:=` (count = .N), by = list(BrthYear, plaprev)] %>%
  .[, `:=` (total = .N), by = list(BrthYear)] %>%
  .[, `:=` (rate = count/total)] %>%
  .[,.(BrthYear, rate, total, plaprev)] %>%
  unique()


dta1 <- dta1[, `:=` (plaprev = factor(plaprev,
                                   levels = c(0,1),
                                   labels = c("No",
                                              "Placenta previa")))]

## getting color for plotting

pal <- c("No" = "#1b9e77",
         "Placenta previa" = "#d95f02")

plot <- ggplot(data = dta1[!tolower(plaprev) %in% "no"],
               aes(x = BrthYear,
                   y = rate,
                   group = plaprev,
                   colour = plaprev,
                   fill = plaprev,
                   #shape = plaprev,
                   text =paste(paste0(plaprev," - ", BrthYear),
                               "<br>Total deliveries:", scales::comma(total, accuracy = 1),
                               paste0("<br>",plaprev," (%):") , scales::percent(rate, accuracy = .01)
                   )
               )
) +
  geom_point(aes(color = plaprev, fill = plaprev)) +
  geom_line(aes(color = plaprev, fill = plaprev)) +
  # Specifying colors by age group
  scale_color_manual(name = "",
                     values = pal,
                     limits = names(pal)) +
  scale_fill_manual(name = "",
                    values = pal,
                    limits = names(pal)) +
  # scale_shape_manual(name = "",
  #                    values = c(15, 16, 17)) +
  # scale_x_date(date_breaks = "2 years",
  #              date_labels = "%Y") +
  scale_x_continuous(#trans = 'log2',
    breaks = pretty) +
  scale_y_continuous(#trans = 'log2',
    breaks = pretty,
    labels = scales::percent_format(accuracy = .1),
    limits = c(0, NA)) +
  labs(y = 'Proportion of deliveries', x = '') +
  theme_classic()

plotly::ggplotly(plot, tooltip = c("text")) %>%
  plotly::style(hoverlabel = list(
    bgcolor  = "black",
    bordercolor = "transparent",
    align = "left",
    font = list(
      color = "white",
      size = 14,
      face = "bold"
    )
  )) %>%
  plotly::layout(legend = list(orientation = "h", y = -0.25, x = 0.5,
                               xanchor = "center"),
                 font = list(face = "bold", size = 10),
                 showlegend = FALSE)
```

```{r tbl54}
tbl <- dta[, .(BrthYear,
        MO44,
        CONTCTID)] %>%
  unique() %>%
  .[, `:=` (plaprev = fcase(
    MO44 >=1, 1, 
    default = 0
  ))] %>%
  #unique() %>%
  .[, `:=` (count = .N), by = list(BrthYear, plaprev)] %>%
  .[, `:=` (total = .N), by = list(BrthYear)] %>%
  .[, `:=` (rate = 100*count/total)] %>%
  unique() %>%
  melt(measure.vars = c("count", "total", "rate")) %>%
  .[!tolower(variable) %in% "count" &
    !plaprev %in% 0] %>%
  .[, `:=` (plaprev = fifelse(tolower(variable) %in% "total",
                          0, plaprev))] %>%
  .[, .(BrthYear, plaprev, value)] %>%
  unique() %>%
  dcast(plaprev ~ BrthYear, value.var = "value")

# coln <- names(tbl)
# coln[1] <- ""

tbl <- as.data.frame(sapply(tbl, sprintf, fmt = c("%.0f", "%1.2f%%")))

#names(tbl) <- coln

tbl[,1] <- c("Total deliveries", "Placenta previa (%)")

# add comma separator to total count
tbl[1,-1] <- as.character(sapply(as.numeric(tbl[1,-1]), scales::comma_format(accuracy = 1)))

reactable(tbl,
          rownames = FALSE,
          resizable = TRUE, 
          onClick = "expand",
          highlight = TRUE, 
          compact = TRUE,
          defaultPageSize = nrow(tbl),
          defaultColDef = colDef(align = "right",
                                 vAlign = "center",
                                 format = colFormat(locales = "en-CA", separators = TRUE),
                                 style = list(fontSize = 14),
                                 headerStyle = list(fontSize = 14),
                                 minWidth = 50,
                                 maxWidth = 180),
          # rowStyle = function(index){
          #   if (index == nrow(tbl)) list(color = "black",
          #                                backgroundColor = "#e1e1e1",
          #                                fontWeight = "bold",
          #                                align = "center",
          #                                borderBottom = "#dddddd")
          # },
          columns = list(
            plaprev = colDef(name = "",
                             align = "left",
                             width = 180)
          )
) %>%
  reactablefmtr::add_source("Note: Placenta previa is diagnosed when the placenta entirely or partially covers the opening of the uterus (cervix). The diagnosis is not made on ultrasound alone and must be confirmed clinically.",
                            font_size = 12,
                            #font_weight = "bold"
  )

# knitr::kable(tbl,
#              format = "html",
#              booktabs = TRUE) %>%
#   # add an hline before the total
#   #kableExtra::row_spec(nrow(tbl) - 1, hline_after = TRUE) %>%  #kableExtra::row_spec(nrow(tbl), bold = TRUE) %>%
# kableExtra::row_spec(0, bold = TRUE) %>%
# #kableExtra::column_spec(1, border_right = TRUE, bold  = TRUE) %>%
# kableExtra::kable_styling(font_size = 12, full_width = FALSE,
#                             bootstrap_options = c("hover","condensed", "responsive")) %>%
#   kableExtra::add_footnote(label = c("Note: Placenta previa is diagnosed when the placenta entirely or partially covers the opening of the uterus (cervix). The diagnosis is not made on ultrasound alone and must be confirmed clinically."))
```

## Placental abruption by year, Nova Scotia, `r paste0(params$year1,"-",params$year2)` {#section-55}
  
```{r plot55, fig.cap= paste('Placental abruption by year, Nova Scotia,', paste0(params$year1,"-",params$year2))}

dta1 <- dta[, .(BrthYear,
        MO45,
        CONTCTID)] %>%
  unique() %>%
  .[, `:=` (plaabrp = fcase(
    MO45 >=1, 1, 
    default = 0
  ))] %>%
  #unique() %>%
  .[, `:=` (count = .N), by = list(BrthYear, plaabrp)] %>%
  .[, `:=` (total = .N), by = list(BrthYear)] %>%
  .[, `:=` (rate = count/total)] %>%
  .[,.(BrthYear, rate, total, plaabrp)] %>%
  unique()


dta1 <- dta1[, `:=` (plaabrp = factor(plaabrp,
                                   levels = c(0,1),
                                   labels = c("No",
                                              "Placenta abruption")))]

## getting color for plotting

pal <- c("No" = "#1b9e77",
         "Placenta abruption" = "#d95f02")

plot <- ggplot(data = dta1[!tolower(plaabrp) %in% "no"],
               aes(x = BrthYear,
                   y = rate,
                   group = plaabrp,
                   colour = plaabrp,
                   fill = plaabrp,
                   #shape = plaabrp,
                   text =paste(paste0(plaabrp," - ", BrthYear),
                               "<br>Total deliveries:", scales::comma(total, accuracy = 1),
                               paste0("<br>"," (%):") , scales::percent(rate, accuracy = .01)
                   )
               )
) +
  geom_point(aes(color = plaabrp, fill = plaabrp)) +
  geom_line(aes(color = plaabrp, fill = plaabrp)) +
  # Specifying colors by age group
  scale_color_manual(name = "",
                     values = pal,
                     limits = names(pal)) +
  scale_fill_manual(name = "",
                    values = pal,
                    limits = names(pal)) +
  # scale_shape_manual(name = "",
  #                    values = c(15, 16, 17)) +
  # scale_x_date(date_breaks = "2 years",
  #              date_labels = "%Y") +
  scale_x_continuous(#trans = 'log2',
    breaks = pretty) +
  scale_y_continuous(#trans = 'log2',
    breaks = pretty,
    labels = scales::percent_format(accuracy = .1),
    limits = c(0, NA)) +
  labs(y = 'Proportion of deliveries', x = '') +
  theme_classic()

plotly::ggplotly(plot, tooltip = c("text")) %>%
  plotly::style(hoverlabel = list(
    bgcolor  = "black",
    bordercolor = "transparent",
    align = "left",
    font = list(
      color = "white",
      size = 14,
      face = "bold"
    )
  )) %>%
  plotly::layout(legend = list(orientation = "h", y = -0.25, x = 0.5,
                               xanchor = "center"),
                 font = list(face = "bold", size = 10),
                 showlegend = FALSE)
```

```{r tbl55}
tbl <- dta[, .(BrthYear,
        MO45,
        CONTCTID)] %>%
  unique() %>%
  .[, `:=` (plaabrp = fcase(
    MO45 >=1, 1, 
    default = 0
  ))] %>%
  #unique() %>%
  .[, `:=` (count = .N), by = list(BrthYear, plaabrp)] %>%
  .[, `:=` (total = .N), by = list(BrthYear)] %>%
  .[, `:=` (rate = 100*count/total)] %>%
  unique() %>%
  melt(measure.vars = c("count", "total", "rate")) %>%
  .[!tolower(variable) %in% "count" &
    !plaabrp %in% 0] %>%
  .[, `:=` (plaabrp = fifelse(tolower(variable) %in% "total",
                          0, plaabrp))] %>%
  .[, .(BrthYear, plaabrp, value)] %>%
  unique() %>%
  dcast(plaabrp ~ BrthYear, value.var = "value")

# coln <- names(tbl)
# coln[1] <- ""

tbl <- as.data.frame(sapply(tbl, sprintf, fmt = c("%.0f", "%1.2f%%")))

#names(tbl) <- coln

tbl[,1] <- c("Total deliveries", "Placenta abruption (%)")

# add comma separator to total count
tbl[1,-1] <- as.character(sapply(as.numeric(tbl[1,-1]), scales::comma_format(accuracy = 1)))

reactable(tbl,
          rownames = FALSE,
          resizable = TRUE, 
          onClick = "expand",
          highlight = TRUE, 
          compact = TRUE,
          defaultPageSize = nrow(tbl),
          defaultColDef = colDef(align = "right",
                                 vAlign = "center",
                                 format = colFormat(locales = "en-CA", separators = TRUE),
                                 style = list(fontSize = 14),
                                 headerStyle = list(fontSize = 14),
                                 minWidth = 50,
                                 maxWidth = 180),
          # rowStyle = function(index){
          #   if (index == nrow(tbl)) list(color = "black",
          #                                backgroundColor = "#e1e1e1",
          #                                fontWeight = "bold",
          #                                align = "center",
          #                                borderBottom = "#dddddd")
          # },
          columns = list(
            plaabrp = colDef(name = "",
                             align = "left",
                             width = 180)
          )
) %>%
  reactablefmtr::add_source("Note: Placental abruption is defined as bleeding from the placental site due to the partial or complete separation of the placenta. The diagnosis is not made on ultrasound alone and must be confirmed clinically.",
                            font_size = 12,
                            #font_weight = "bold"
  )

# knitr::kable(tbl,
#              format = "html",
#              booktabs = TRUE) %>%
#   # add an hline before the total
#   #kableExtra::row_spec(nrow(tbl) - 1, hline_after = TRUE) %>%  #kableExtra::row_spec(nrow(tbl), bold = TRUE) %>%
# kableExtra::row_spec(0, bold = TRUE) %>%
# #kableExtra::column_spec(1, border_right = TRUE, bold  = TRUE) %>%
# kableExtra::kable_styling(font_size = 12, full_width = FALSE,
#                             bootstrap_options = c("hover","condensed", "responsive")) %>%
#   kableExtra::add_footnote(label = c("Placental abruption is defined as bleeding from the placental site due to the partial or complete separation of the placenta. The diagnosis is not made on ultrasound alone and must be confirmed clinically."))
```

## Perineal laceration deliveries by parity and year, Nova Scotia, `r paste0(params$year1,"-",params$year2)` {#section-56}
  
```{r plot56, fig.cap= paste('Perineal laceration deliveries by parity and year, Nova Scotia,', paste0(params$year1,"-",params$year2))}

dta1 <- dta[, .(BrthYear,
       MO701,
       MO702, MO70201, MO70204, MO70211, MO70214, MO70221, MO70231, MO70234, MO70281, MO70291,
       MO703, MO70301, MO70304,
       DLPARA,
       DMMETHOD,
       MOTHERID)] %>%
  unique() %>%
  # Primiparous women
  # vaginal delivery
  .[#DLPARA %in% 0 &
    # vaginal
    tolower(DMMETHOD) %in% c("biv", "spt", "vac", "low", "mid")] %>%
  .[, `:=` (lacer = fcase(
    # second degree
    MO701 >=1, 1,
    # 3rd degree
    MO702 >=1 | MO70201 >=1 | MO70204 >=1 | MO70211 >=1 | MO70214 >=1 | MO70221 >=1 | MO70231 >=1 | MO70234 >=1 | MO70281 >=1 | MO70291 >=1 |
    MO703 >=1 | MO70301 >=1 | MO70304 >=1, 2, 
    default = 0
  ),
  parity = fcase(
    DLPARA %in% 0 , 1,
    DLPARA >= 1, 2,
    default = 3
    ))] %>%
  #unique() %>%
  .[, `:=` (count = .N), by = list(BrthYear, lacer, parity)] %>%
  .[, `:=` (total = .N), by = list(BrthYear, parity)] %>%
  .[, `:=` (rate = count/total)] %>%
  .[,.(BrthYear, rate, total, lacer, parity)] %>%
  .[!parity %in% 3] %>%
  unique()


dta1 <- dta1[, `:=` (lacer = factor(lacer,
                                   levels = c(0,1,2),
                                   labels = c("No",
                                              "2\u207f\u1d48 degree laceration",
                                              "3\u02b3\u1d48 or 4\u1d57\u02b0 degree laceration")),
                      parity = factor(parity,
                                   levels = c(1,2),
                                   labels = c("Primiparous",
                                              "Multiparous")))]

## getting color for plotting

pal <- c("No" = "#1b9e77",
         "2\u207f\u1d48 degree laceration" = "#d95f02",
         "3\u02b3\u1d48 or 4\u1d57\u02b0 degree laceration" = "#7570b3")

plot <- ggplot(data = dta1[!tolower(lacer) %in% "no"],
               aes(x = BrthYear,
                   y = rate,
                   group = interaction(lacer, parity),
                   colour = lacer,
                   fill = lacer,
                   shape = lacer,
                   linetype = parity,
                   text =paste(paste0(lacer," - ", parity," - ",BrthYear),
                               "<br>Total deliveries:", scales::comma(total, accuracy = 1),
                               paste0("<br>",lacer," (%):") , scales::percent(rate, accuracy = .1)
                   )
               )
) +
  geom_point(aes(color = lacer, fill = lacer, shape = lacer)) +
  geom_line(aes(color = lacer, fill = lacer)) +
  # Specifying colors by age group
  scale_color_manual(name = "",
                     values = pal,
                     limits = names(pal)) +
  scale_fill_manual(name = "",
                    values = pal,
                    limits = names(pal)) +
  scale_shape_manual(name = "",
                     values = c(21, 22)) +
  scale_linetype_manual(name = "",
                        values = c(1,2)) +
  # scale_x_date(date_breaks = "2 years",
  #              date_labels = "%Y") +
  scale_x_continuous(#trans = 'log2',
    breaks = pretty) +
  scale_y_continuous(#trans = 'log2',
    breaks = pretty,
    labels = scales::percent_format(accuracy = .1),
    limits = c(0, NA)) +
  labs(y = 'Proportion of vaginal deliveries', x = '') +
  theme_classic()

plotly::ggplotly(plot, tooltip = c("text")) %>%
  plotly::style(hoverlabel = list(
    bgcolor  = "black",
    bordercolor = "transparent",
    align = "left",
    align = "left",
    font = list(
      color = "white",
      size = 14,
      face = "bold"
    )
  )) %>%
  plotly::layout(legend = list(orientation = "h", y = -0.25, x = 0.5,
                               xanchor = "center"),
                 font = list(face = "bold", size = 10),
                 showlegend = TRUE)
```

```{r tbl56}
tbl <- dta[, .(BrthYear,
       MO701,
       MO702, MO70201, MO70204, MO70211, MO70214, MO70221, MO70231, MO70234, MO70281, MO70291,
       MO703, MO70301, MO70304,
       DLPARA,
       DMMETHOD,
       MOTHERID)] %>%
  unique() %>%
  # Primiparous women
  # vaginal delivery
  .[#DLPARA %in% 0 &
    # vaginal
    tolower(DMMETHOD) %in% c("biv", "spt", "vac", "low", "mid")] %>%
  .[, `:=` (lacer = fcase(
    # second degree
    MO701 >=1, 1,
    # 3rd degree
    MO702 >=1 | MO70201 >=1 | MO70204 >=1 | MO70211 >=1 | MO70214 >=1 | MO70221 >=1 | MO70231 >=1 | MO70234 >=1 | MO70281 >=1 | MO70291 >=1 |
    MO703 >=1 | MO70301 >=1 | MO70304 >=1, 2, 
    default = 0
  ),
 parity = fcase(
    DLPARA %in% 0 , 1,
    DLPARA >= 1, 2,
    default = 3
    ))] %>%
  #unique() %>%
  .[, `:=` (count = .N), by = list(BrthYear, lacer, parity)] %>%
  .[, `:=` (total = .N), by = list(BrthYear, parity)] %>%
  .[, `:=` (rate = 100*count/total)] %>%
  .[!parity %in% 3] %>%
  unique()

tbl1 <- tbl %>%
  .[parity %in% 1] %>%
  melt(measure.vars = c("count", "total", "rate")) %>%
  .[!tolower(variable) %in% "count" &
    !lacer %in% 0] %>%
  .[, `:=` (lacer = fifelse(tolower(variable) %in% "total",
                          0, lacer))] %>%
  .[, .(BrthYear, lacer, value)] %>%
  unique() %>%
  dcast(lacer ~ BrthYear, value.var = "value")

# coln <- names(tbl)
# coln[1] <- ""

tbl1 <- as.data.frame(sapply(tbl1, sprintf, fmt = c("%.0f", "%1.1f%%", "%1.1f%%")))

#names(tbl) <- coln

tbl1[,1] <- c("Total vaginal deliveries\u1d43", "2\u207f\u1d48 degree laceration", "3\u02b3\u1d48 or 4\u1d57\u02b0 degree laceration")

# add comma separator to total count
tbl1[1,-1] <- as.character(sapply(as.numeric(tbl1[1,-1]), scales::comma_format(accuracy = 1)))

temp <- as.data.frame(t(c("Primiparous", rep(NA, ncol(tbl1)-1))))
names(temp) <- names(tbl1)

tbl1 <- rbind(temp,tbl1)

tbl2 <- tbl %>%
  .[parity %in% 2] %>%
  melt(measure.vars = c("count", "total", "rate")) %>%
  .[!tolower(variable) %in% "count" &
    !lacer %in% 0] %>%
  .[, `:=` (lacer = fifelse(tolower(variable) %in% "total",
                          0, lacer))] %>%
  .[, .(BrthYear, lacer, value)] %>%
  unique() %>%
  dcast(lacer ~ BrthYear, value.var = "value")

# coln <- names(tbl)
# coln[1] <- ""

tbl2 <- as.data.frame(sapply(tbl2, sprintf, fmt = c("%.0f", "%1.1f%%", "%1.1f%%")))

#names(tbl) <- coln

tbl2[,1] <- c("Total vaginal deliveries\u1d43", "2\u207f\u1d48 degree laceration", "3\u02b3\u1d48 or 4\u1d57\u02b0 degree laceration")

# add comma separator to total count
tbl2[1,-1] <- as.character(sapply(as.numeric(tbl2[1,-1]), scales::comma_format(accuracy = 1)))

temp <- as.data.frame(t(c("Multiparous", rep(NA, ncol(tbl2)-1))))
names(temp) <- names(tbl2)

tbl2 <- rbind(temp,tbl2)

tbl <- setDT(rbind(tbl1, tbl2))

reactable(tbl,
          rownames = FALSE,
          resizable = TRUE, 
          onClick = "expand",
          highlight = TRUE, 
          compact = TRUE,
          defaultPageSize = nrow(tbl),
          defaultColDef = colDef(align = "right",
                                 vAlign = "center",
                                 format = colFormat(locales = "en-CA", separators = TRUE),
                                 style = list(fontSize = 14),
                                 headerStyle = list(fontSize = 14),
                                 minWidth = 50,
                                 maxWidth = 180),
          rowStyle = function(index){
            if (index %in% (seq(1,nrow(tbl), by = 4))) list(color = "black",
                                         backgroundColor = "#e1e1e1",
                                         fontWeight = "bold",
                                         align = "center",
                                         borderBottom = "#dddddd")
          },
          columns = list(
            lacer = colDef(name = "",
                             align = "left",
                             width = 180)
          )
) %>%
  reactablefmtr::add_source("\u1d43 With known partner status.",
                            font_size = 12,
                            #font_weight = "bold"
  ) %>%
  reactablefmtr::add_source("Note: Maternal perineal laceration, rupture or tear during delivery involving the pelvic floor, perineal muscles, or vaginal muscles (2\u207f\u1d48 degree), anal sphincter (3\u02b3\u1d48 degree), or rectal mucosa (4\u1d57\u02b0 degree).",
                            font_size = 12,
                            #font_weight = "bold"
  )

# knitr::kable(tbl,
#              format = "html",
#              booktabs = TRUE) %>%
#   # add an hline before the total
#   #kableExtra::row_spec(nrow(tbl) - 1, hline_after = TRUE) %>%  #kableExtra::row_spec(nrow(tbl), bold = TRUE) %>%
# kableExtra::row_spec(0, bold = TRUE) %>%
# #kableExtra::column_spec(1, border_right = TRUE, bold  = TRUE) %>%
# kableExtra::kable_styling(font_size = 12, full_width = FALSE,
#                             bootstrap_options = c("hover","condensed", "responsive")) %>%
#   kableExtra::add_footnote(label = c("Note: Maternal perineal laceration, rupture or tear during delivery involving the pelvic floor, perineal muscles, or vaginal muscles (2$^{nd}$ degree), anal sphincter (3$^{rd}$ degree), or rectal mucosa (4$^{th}$ degree)."), escape = FALSE)
```


## Postpartum hemorrhage by year, Nova Scotia, `r paste0(params$year1,"-",params$year2)` {#section-58}
  
```{r plot58, fig.cap= paste('Postpartum hemorrhage by year, Nova Scotia,', paste0(params$year1,"-",params$year2))}

dta1 <- dta[, .(BrthYear,
     R003_00400, R003_00900, R003_01300, R003_01400, R003_02200,
     R027_00400, 
     R029_00100, R029_00200, R029_00300, R029_00400, 
     MO72,
     M_1KT51, M_1RM13GQC2,
     M_1RM13GQW0, M_5PC91HR, M_5PC91HT,
     M_5PC91HU, M_5PC91HV, M_5PC91LA,
     CONTCTID)] %>%
  unique() %>%
  .[, `:=` (postph = fcase(
    R003_00400 >=1 | R003_00900 >=1 | R003_01300 >=1 | R003_01400 >=1 | R003_02200 >=1 |
     R027_00400 >=1 | 
     R029_00100 >=1 | R029_00200 >=1 | R029_00300 >=1 | R029_00400 >=1 | 
     MO72 >=1 |
     M_1KT51 >=1 | M_1RM13GQC2 >=1 |
     M_1RM13GQW0 >=1 | M_5PC91HR >=1 | M_5PC91HT >=1 |
     M_5PC91HU >=1 | M_5PC91HV >=1 | M_5PC91LA >=1, 1,
    default = 0
  ))] %>%
  #unique() %>%
  .[, `:=` (count = .N), by = list(BrthYear, postph)] %>%
  .[, `:=` (total = .N), by = list(BrthYear)] %>%
  .[, `:=` (rate = count/total)] %>%
  .[,.(BrthYear, total, rate, postph)] %>%
  unique()


dta1 <- dta1[, `:=` (postph = factor(postph,
                                   levels = c(0,1),
                                   labels = c("No",
                                              "Postpartum hemorrhage")))]

## getting color for plotting

pal <- c("No" = "#1b9e77",
         "Postpartum hemorrhage" = "#d95f02")

plot <- ggplot(data = dta1[!tolower(postph) %in% "no"],
               aes(x = BrthYear,
                   y = rate,
                   group = postph,
                   colour = postph,
                   fill = postph,
                   #shape = postph,
                   text =paste(paste0(postph," - ", BrthYear),
                               "<br>Total deliveries:", scales::comma(total, accuracy = 1),
                               paste0("<br>",postph," (%):") , scales::percent(rate, accuracy = .1)
                   )
               )
) +
  geom_point(aes(color = postph, fill = postph)) +
  geom_line(aes(color = postph, fill = postph)) +
  # Specifying colors by age group
  scale_color_manual(name = "",
                     values = pal,
                     limits = names(pal)) +
  scale_fill_manual(name = "",
                    values = pal,
                    limits = names(pal)) +
  # scale_shape_manual(name = "",
  #                    values = c(15, 16)) +
  # scale_x_date(date_breaks = "2 years",
  #              date_labels = "%Y") +
  scale_x_continuous(#trans = 'log2',
    breaks = pretty) +
  scale_y_continuous(#trans = 'log2',
    breaks = pretty,
    labels = scales::percent_format(accuracy = 1),
    limits = c(0, NA)) +
  labs(y = 'Proportion of deliveries', x = '') +
  theme_classic()

plotly::ggplotly(plot, tooltip = c("text")) %>%
  plotly::style(hoverlabel = list(
    bgcolor  = "black",
    bordercolor = "transparent",
    align = "left",
    font = list(
      color = "white",
      size = 14,
      face = "bold"
    )
  )) %>%
  plotly::layout(legend = list(orientation = "h", y = -0.25, x = 0.5,
                               xanchor = "center"),
                 font = list(face = "bold", size = 10),
                 showlegend = FALSE)
```

```{r tbl58}
tbl <- dta[, .(BrthYear,
     R003_00400, R003_00900, R003_01300, R003_01400, R003_02200,
     R027_00400, 
     R029_00100, R029_00200, R029_00300, R029_00400, 
     MO72,
     M_1KT51, M_1RM13GQC2,
     M_1RM13GQW0, M_5PC91HR, M_5PC91HT,
     M_5PC91HU, M_5PC91HV, M_5PC91LA,
     CONTCTID)] %>%
  unique() %>%
  .[, `:=` (postph = fcase(
    R003_00400 >=1 | R003_00900 >=1 | R003_01300 >=1 | R003_01400 >=1 | R003_02200 >=1 |
     R027_00400 >=1 | 
     R029_00100 >=1 | R029_00200 >=1 | R029_00300 >=1 | R029_00400 >=1 | 
     MO72 >=1 |
     M_1KT51 >=1 | M_1RM13GQC2 >=1 |
     M_1RM13GQW0 >=1 | M_5PC91HR >=1 | M_5PC91HT >=1 |
     M_5PC91HU >=1 | M_5PC91HV >=1 | M_5PC91LA >=1, 1,
    default = 0
  ))] %>%
  #unique() %>%
  .[, `:=` (count = .N), by = list(BrthYear, postph)] %>%
  .[, `:=` (total = .N), by = list(BrthYear)] %>%
  .[, `:=` (rate = 100*count/total)] %>%
  unique() %>%
  melt(measure.vars = c("count", "total", "rate")) %>%
  .[!tolower(variable) %in% "count" &
    !postph %in% 0] %>%
  .[, `:=` (postph = fifelse(tolower(variable) %in% "total",
                          0, postph))] %>%
  .[, .(BrthYear, postph, value)] %>%
  unique() %>%
  dcast(postph ~ BrthYear, value.var = "value")

# coln <- names(tbl)
# coln[1] <- ""

tbl <- as.data.frame(sapply(tbl, sprintf, fmt = c("%.0f", "%1.1f%%")))

#names(tbl) <- coln

tbl[,1] <- c("Total deliveries", "Postpartum hemorrhage (%)")

# add comma separator to total count
tbl[1,-1] <- as.character(sapply(as.numeric(tbl[1,-1]), scales::comma_format(accuracy = 1)))

reactable(tbl,
          rownames = FALSE,
          resizable = TRUE, 
          onClick = "expand",
          highlight = TRUE, 
          compact = TRUE,
          defaultPageSize = nrow(tbl),
          defaultColDef = colDef(align = "right",
                                 vAlign = "center",
                                 format = colFormat(locales = "en-CA", separators = TRUE),
                                 style = list(fontSize = 14),
                                 headerStyle = list(fontSize = 14),
                                 minWidth = 50,
                                 maxWidth = 180),
          # rowStyle = function(index){
          #   if (index == nrow(tbl)) list(color = "black",
          #                                backgroundColor = "#e1e1e1",
          #                                fontWeight = "bold",
          #                                align = "center",
          #                                borderBottom = "#dddddd")
          # },
          columns = list(
            postph = colDef(name = "",
                             align = "left",
                             width = 180)
          )
) %>%
  reactablefmtr::add_source("Note: Postpartum hemorrhage is diagnosed if, after the delivery of the fetus, excessive maternal bleeding occurs from the genital tract with an estimated blood loss of greater than 500 mL for vaginal deliveries or 1000 mL for Caesarean section deliveries.",
                            font_size = 12,
                            #font_weight = "bold"
  )

# knitr::kable(tbl,
#              format = "html",
#              booktabs = TRUE) %>%
#   # add an hline before the total
#   #kableExtra::row_spec(nrow(tbl) - 1, hline_after = TRUE) %>%  #kableExtra::row_spec(nrow(tbl), bold = TRUE) %>%
# kableExtra::row_spec(0, bold = TRUE) %>%
# #kableExtra::column_spec(1, border_right = TRUE, bold  = TRUE) %>%
# kableExtra::kable_styling(font_size = 12, full_width = FALSE,
#                             bootstrap_options = c("hover","condensed", "responsive")) %>%
#   kableExtra::add_footnote(label = c("Note: Postpartum hemorrhage is diagnosed if, after the delivery of the fetus, excessive maternal bleeding occurs from the genital tract with an estimated blood loss of greater than 500 mL for vaginal deliveries or 1000 mL for Caesarean section deliveries."))
```

## Maternal blood transfusion by year, Nova Scotia, `r paste0(params$year1,"-",params$year2)` {#section-59}
  
```{r plot59, fig.cap= paste('Maternal blood transfusion by year, Nova Scotia,', paste0(params$year1,"-",params$year2))}

dta1 <- dta[, .(BrthYear,
     RBC,
     CONTCTID)] %>%
  unique() %>%
  .[, `:=` (transf = fcase(
    RBC >=1 , 1,
    default = 0
  ))] %>%
  #unique() %>%
  .[, `:=` (count = .N), by = list(BrthYear, transf)] %>%
  .[, `:=` (total = .N), by = list(BrthYear)] %>%
  .[, `:=` (rate = count/total)] %>%
  .[,.(BrthYear, rate, total, transf)] %>%
  unique()


dta1 <- dta1[, `:=` (transf = factor(transf,
                                   levels = c(0,1),
                                   labels = c("No",
                                              "Transfusion")))]

## getting color for plotting

pal <- c("No" = "#1b9e77",
         "Transfusion" = "#d95f02")

plot <- ggplot(data = dta1[!tolower(transf) %in% "no"],
               aes(x = BrthYear,
                   y = rate,
                   group = transf,
                   colour = transf,
                   fill = transf,
                   #shape = transf,
                   text =paste(paste0(transf," - ",BrthYear),
                               "<br>Total deliveries:", scales::comma(total, accuracy = 1),
                               paste0("<br>",transf, " (%):") , scales::percent(rate, accuracy = .01)
                   )
               )
) +
  geom_point(aes(color = transf, fill = transf)) +
  geom_line(aes(color = transf, fill = transf)) +
  # Specifying colors by age group
  scale_color_manual(name = "",
                     values = pal,
                     limits = names(pal)) +
  scale_fill_manual(name = "",
                    values = pal,
                    limits = names(pal)) +
  # scale_shape_manual(name = "",
  #                    values = c(15, 16)) +
  # scale_x_date(date_breaks = "2 years",
  #              date_labels = "%Y") +
  scale_x_continuous(#trans = 'log2',
    breaks = pretty) +
  scale_y_continuous(#trans = 'log2',
    breaks = pretty,
    labels = scales::percent_format(accuracy = .1),
    limits = c(0, NA)) +
  labs(y = 'Proportion of deliveries', x = '') +
  theme_classic()

plotly::ggplotly(plot, tooltip = c("text")) %>%
  plotly::style(hoverlabel = list(
    bgcolor  = "black",
    bordercolor = "transparent",
    align = "left",
    font = list(
      color = "white",
      size = 14,
      face = "bold"
    )
  )) %>%
  plotly::layout(legend = list(orientation = "h", y = -0.25, x = 0.5,
                               xanchor = "center"),
                 font = list(face = "bold", size = 10),
                 showlegend = FALSE)
```

```{r tbl59}
tbl <- dta[, .(BrthYear,
     RBC,
     CONTCTID)] %>%
  unique() %>%
  .[, `:=` (transf = fcase(
    RBC >=1 , 1,
    default = 0
  ))] %>%
  #unique() %>%
  .[, `:=` (count = .N), by = list(BrthYear, transf)] %>%
  .[, `:=` (total = .N), by = list(BrthYear)] %>%
  .[, `:=` (rate = 100*count/total)] %>%
  unique() %>%
  melt(measure.vars = c("count", "total", "rate")) %>%
  .[!tolower(variable) %in% "count" &
    !transf %in% 0] %>%
  .[, `:=` (transf = fifelse(tolower(variable) %in% "total",
                          0, transf))] %>%
  .[, .(BrthYear, transf, value)] %>%
  unique() %>%
  dcast(transf ~ BrthYear, value.var = "value")

# coln <- names(tbl)
# coln[1] <- ""

tbl <- as.data.frame(sapply(tbl, sprintf, fmt = c("%.0f", "%1.2f%%")))

#names(tbl) <- coln

tbl[,1] <- c("Total deliveries", "Transfusion (%)")

# add comma separator to total count
tbl[1,-1] <- as.character(sapply(as.numeric(tbl[1,-1]), scales::comma_format(accuracy = 1)))

reactable(tbl,
          rownames = FALSE,
          resizable = TRUE, 
          onClick = "expand",
          highlight = TRUE, 
          compact = TRUE,
          defaultPageSize = nrow(tbl),
          defaultColDef = colDef(align = "right",
                                 vAlign = "center",
                                 format = colFormat(locales = "en-CA", separators = TRUE),
                                 style = list(fontSize = 14),
                                 headerStyle = list(fontSize = 14),
                                 minWidth = 50,
                                 maxWidth = 180),
          # rowStyle = function(index){
          #   if (index == nrow(tbl)) list(color = "black",
          #                                backgroundColor = "#e1e1e1",
          #                                fontWeight = "bold",
          #                                align = "center",
          #                                borderBottom = "#dddddd")
          # },
          columns = list(
            transf = colDef(name = "",
                             align = "left",
                             width = 180)
          )
) %>%
  reactablefmtr::add_source("Note: One or more maternal transfusions of red blood cells in the antepartum, intrapartum, or postpartum periods.",
                            font_size = 12,
                            #font_weight = "bold"
  )

# knitr::kable(tbl,
#              format = "html",
#              booktabs = TRUE) %>%
#   # add an hline before the total
#   #kableExtra::row_spec(nrow(tbl) - 1, hline_after = TRUE) %>%  #kableExtra::row_spec(nrow(tbl), bold = TRUE) %>%
# kableExtra::row_spec(0, bold = TRUE) %>%
# #kableExtra::column_spec(1, border_right = TRUE, bold  = TRUE) %>%
# kableExtra::kable_styling(font_size = 12, full_width = FALSE,
#                             bootstrap_options = c("hover","condensed", "responsive")) %>%
#   kableExtra::add_footnote(label = c("Note: One or more maternal transfusions of red blood cells in the antepartum, intrapartum, or postpartum periods."))
```

## Maternal antepartum hospital length of stay (in hours) by year, Nova Scotia, `r paste0(params$year1,"-",params$year2)` {#section-510}
  
```{r plot510, fig.cap= paste('Maternal antepartum hospital length of stay (in hours) by year, Nova Scotia,', paste0(params$year1,"-",params$year2))}

dta1 <- dta[, .(BrthYear,
     APLoS,
     MOTHERID)] %>%
  unique() %>%
  .[!is.na(APLoS)] %>%
  .[, `:=` (total = .N,
            mean = mean(APLoS, na.rm = TRUE),
            med = median(APLoS, na.rm = TRUE)), by = list(BrthYear)] %>%
  .[,.(BrthYear, total, med)] %>%
  unique()

plot <- ggplot(data = dta1,
               aes(x = BrthYear,
                   y = med,
                   group = 1,
                   #shape = BrthYear,
                   text =paste(BrthYear,
                               "<br>Total deliveries:", scales::comma(total, accuracy = 1),
                               "<br>% Median lenght of stay (in hours):" , scales::comma(med, accuracy = .1)
                   )
               )
) +
  geom_point(aes(color = "#d95f02", fill = "#d95f02")) +
  geom_line(aes(color = "#d95f02", fill = "#d95f02")) +
  # Specifying colors by age group
  scale_color_manual(name = "",
                     values = "#d95f02") +
  scale_fill_manual(name = "",
                    values = "#d95f02") +
  # scale_shape_manual(name = "",
  #                    values = c(15, 16)) +
  # scale_x_date(date_breaks = "2 years",
  #              date_labels = "%Y") +
  scale_x_continuous(#trans = 'log2',
    breaks = pretty) +
  scale_y_continuous(#trans = 'log2',
    breaks = pretty,
    labels = scales::comma_format(accuracy = 1),
    limits = c(0, NA)) +
  labs(y = 'Median antepartum LoS (in hours)', x = '') +
  theme_classic()

plotly::ggplotly(plot, tooltip = c("text")) %>%
  plotly::style(hoverlabel = list(
    bgcolor  = "black",
    bordercolor = "transparent",
    align = "left",
    font = list(
      color = "white",
      size = 14,
      face = "bold"
    )
  )) %>%
  plotly::layout(legend = list(orientation = "h", y = -0.25, x = 0.5,
                               xanchor = "center"),
                 font = list(face = "bold", size = 10),
                 showlegend = FALSE)
```

```{r tbl510}
tbl <- dta[, .(BrthYear,
     APLoS,
     MOTHERID)] %>%
  unique() %>%
  .[!is.na(APLoS)] %>%
  .[, `:=` (total = .N,
            mean = mean(APLoS, na.rm = TRUE),
            med = median(APLoS, na.rm = TRUE)), by = list(BrthYear)] %>%
  unique() %>%
  melt(measure.vars = c("total", "mean", "med")) %>%
  # .[!tolower(variable) %in% "count" &
  #   !transf %in% 0] %>%
  # .[, `:=` (transf = fifelse(tolower(variable) %in% "total",
  #                         0, transf))] %>%
  .[, .(BrthYear, variable, value)] %>%
  unique() %>%
  dcast(variable ~ BrthYear, value.var = "value")

# coln <- names(tbl)
# coln[1] <- ""

tbl <- as.data.frame(sapply(tbl, sprintf, fmt = c("%.0f", "%.1f", "%.1f")))

#names(tbl) <- coln

tbl[,1] <- c("Total deliveries", "Mean (in hours)", "Median (in hours)")

# add comma separator to total count
tbl[1,-1] <- as.character(sapply(as.numeric(tbl[1,-1]), scales::comma_format(accuracy = 1)))

reactable(tbl,
          rownames = FALSE,
          resizable = TRUE, 
          onClick = "expand",
          highlight = TRUE, 
          compact = TRUE,
          defaultPageSize = nrow(tbl),
          defaultColDef = colDef(align = "right",
                                 vAlign = "center",
                                 format = colFormat(locales = "en-CA", separators = TRUE),
                                 style = list(fontSize = 14),
                                 headerStyle = list(fontSize = 14),
                                 minWidth = 50,
                                 maxWidth = 180),
          # rowStyle = function(index){
          #   if (index == nrow(tbl)) list(color = "black",
          #                                backgroundColor = "#e1e1e1",
          #                                fontWeight = "bold",
          #                                align = "center",
          #                                borderBottom = "#dddddd")
          # },
          columns = list(
            variable = colDef(name = "Lenght of stay",
                             align = "left",
                             width = 180)
          )
)

# knitr::kable(tbl,
#              format = "html",
#              booktabs = TRUE) %>%
#   # add an hline before the total
#   #kableExtra::row_spec(nrow(tbl) - 1, hline_after = TRUE) %>%  #kableExtra::row_spec(nrow(tbl), bold = TRUE) %>%
# kableExtra::row_spec(0, bold = TRUE) %>%
# #kableExtra::column_spec(1, border_right = TRUE, bold  = TRUE) %>%
# kableExtra::kable_styling(font_size = 12, full_width = FALSE,
#                             bootstrap_options = c("hover","condensed", "responsive"))
```

## Maternal postpartum hospital length of stay (in hours) by type of delivery and year, Nova Scotia, `r paste0(params$year1,"-",params$year2)` {#section-511}
  
```{r plot511, fig.cap= paste('Maternal postpartum hospital length of stay (in hours) by type of delivery and year, Nova Scotia,', paste0(params$year1,"-",params$year2))}

dta1 <- dta[, .(BrthYear,
     PPLoS,
     DMMETHOD,
     MOTHERID)] %>%
  unique() %>%
  .[!is.na(PPLoS)] %>%
  .[, `:=` (deliv = fcase(
      # spontaneous vaginal
      tolower(DMMETHOD) %in% c("biv", "spt"), 1,
      # assisted vaginal
      tolower(DMMETHOD) %in% c("vac", "low", "mid"), 2,
      # c-section
      tolower(DMMETHOD) %in% c("cst"), 3,
      default = 4
    ))] %>%
  .[!deliv %in% 4] %>%
  .[, `:=` (total = .N,
            mean = mean(PPLoS, na.rm = TRUE),
            med = median(PPLoS, na.rm = TRUE)), by = list(BrthYear, deliv)] %>%
  .[,.(BrthYear, total, med, deliv)] %>%
  unique()

dta1 <- dta1[, `:=` (deliv = factor(deliv,
                                   levels = c(1,2,3),
                                   labels = c("Spontaneous vaginal",
                                              "Assisted vaginal",
                                              "Caesarean section")))]

## getting color for plotting

pal <- c("Spontaneous vaginal" = "#1b9e77",
         "Assisted vaginal" = "#d95f02",
         "Caesarean section" = "#7570b3")

plot <- ggplot(data = dta1,
               aes(x = BrthYear,
                   y = med,
                   group = deliv,
                   colour = deliv,
                   fill = deliv,
                   shape = deliv,
                   text =paste(paste0(deliv," - ",BrthYear),
                               "<br>Total deliveries:", scales::comma(total, accuracy = 1),
                               "<br> Median lenght of stay (in hours):" , scales::comma(med, accuracy = .1)
                   )
               )
) +
  geom_point(aes(color = deliv, fill = deliv, shape = deliv)) +
  geom_line(aes(color = deliv, fill = deliv)) +
  # Specifying colors by age group
  scale_color_manual(name = "",
                     values = pal,
                     limits = names(pal)) +
  scale_fill_manual(name = "",
                    values = pal,
                     limits = names(pal)) +
  scale_shape_manual(name = "",
                     values = c(21, 22, 23)) +
  # scale_x_date(date_breaks = "2 years",
  #              date_labels = "%Y") +
  scale_x_continuous(#trans = 'log2',
    breaks = pretty) +
  scale_y_continuous(#trans = 'log2',
    breaks = pretty,
    labels = scales::comma_format(accuracy = 1),
    limits = c(0, NA)) +
  labs(y = 'Median postpartum LoS (in hours)', x = '') +
  theme_classic()

plotly::ggplotly(plot, tooltip = c("text")) %>%
  plotly::style(hoverlabel = list(
    bgcolor  = "black",
    bordercolor = "transparent",
    align = "left",
    font = list(
      color = "white",
      size = 14,
      face = "bold"
    )
  )) %>%
  plotly::layout(legend = list(orientation = "h", y = -0.25, x = 0.5,
                               xanchor = "center"),
                 font = list(face = "bold", size = 10),
                 showlegend = TRUE)
```

```{r tbl511}
tbl <- dta[, .(BrthYear,
     PPLoS,
     DMMETHOD,
     MOTHERID)] %>%
  unique() %>%
  .[!is.na(PPLoS)] %>%
  .[, `:=` (deliv = fcase(
      # spontaneous vaginal
      tolower(DMMETHOD) %in% c("biv", "spt"), 1,
      # assisted vaginal
      tolower(DMMETHOD) %in% c("vac", "low", "mid"), 2,
      # c-section
      tolower(DMMETHOD) %in% c("cst"), 3,
      default = 4
    ))] %>%
  .[!deliv %in% 4] %>%
  .[, `:=` (total = .N,
            mean = mean(PPLoS, na.rm = TRUE),
            med = median(PPLoS, na.rm = TRUE)), by = list(BrthYear, deliv)] %>%
  unique()

tbl1 <- tbl %>%
  .[deliv %in% 1] %>%
  melt(measure.vars = c("total", "mean", "med")) %>%
  .[, .(BrthYear, variable, value)] %>%
  unique() %>%
  dcast(variable ~ BrthYear, value.var = "value")

# coln <- names(tbl1)
# coln[1] <- ""

tbl1 <- as.data.frame(sapply(tbl1, sprintf, fmt = c("%.0f", "%.1f", "%.1f")))

#names(tbl1) <- coln
tbl1[,1] <- c("Total deliveries", "Mean (in hours)", "Median (in hours)")

# add comma separator to total count
tbl1[1,-1] <- as.character(sapply(as.numeric(tbl1[1,-1]), scales::comma_format(accuracy = 1)))

temp <- as.data.frame(t(c("Spontenaous vaginal", rep(NA, ncol(tbl1)-1))))
names(temp) <- names(tbl1)

tbl1 <- rbind(temp,tbl1)

tbl2 <- tbl %>%
  .[deliv %in% 2] %>%
  melt(measure.vars = c("total", "mean", "med")) %>%
  .[, .(BrthYear, variable, value)] %>%
  unique() %>%
  dcast(variable ~ BrthYear, value.var = "value")

# coln <- names(tbl2)
# coln[1] <- ""

tbl2 <- as.data.frame(sapply(tbl2, sprintf, fmt = c("%.0f", "%.1f", "%.1f")))

#names(tbl2) <- coln
tbl2[,1] <- c("Total deliveries", "Mean (in hours)", "Median (in hours)")

# add comma separator to total count
tbl2[1,-1] <- as.character(sapply(as.numeric(tbl2[1,-1]), scales::comma_format(accuracy = 1)))

temp <- as.data.frame(t(c("Assisted vaginal", rep(NA, ncol(tbl2)-1))))
names(temp) <- names(tbl2)

tbl2 <- rbind(temp,tbl2)

tbl3 <- tbl %>%
  .[deliv %in% 3] %>%
  melt(measure.vars = c("total", "mean", "med")) %>%
  .[, .(BrthYear, variable, value)] %>%
  unique() %>%
  dcast(variable ~ BrthYear, value.var = "value")

# coln <- names(tbl3)
# coln[1] <- ""

tbl3 <- as.data.frame(sapply(tbl3, sprintf, fmt = c("%.0f", "%.1f", "%.1f")))

#names(tbl3) <- coln
tbl3[,1] <- c("Total deliveries", "Mean (in hours)", "Median (in hours)")

# add comma separator to total count
tbl3[1,-1] <- as.character(sapply(as.numeric(tbl3[1,-1]), scales::comma_format(accuracy = 1)))

temp <- as.data.frame(t(c("Caesarean section", rep(NA, ncol(tbl3)-1))))
names(temp) <- names(tbl3)

tbl3 <- rbind(temp,tbl3)

tbl <- setDT(rbind(tbl1,tbl2,tbl3))

#names(tbl)[1] <- "Type of delivery"

reactable(tbl,
          rownames = FALSE,
          resizable = TRUE, 
          onClick = "expand",
          highlight = TRUE, 
          compact = TRUE,
          defaultPageSize = nrow(tbl),
          defaultColDef = colDef(align = "right",
                                 vAlign = "center",
                                 format = colFormat(locales = "en-CA", separators = TRUE),
                                 style = list(fontSize = 14),
                                 headerStyle = list(fontSize = 14),
                                 minWidth = 50,
                                 maxWidth = 180),
          rowStyle = function(index){
            if (index %in% (seq(1, nrow(tbl), by = 4))) list(color = "black",
                                         backgroundColor = "#e1e1e1",
                                         fontWeight = "bold",
                                         align = "center",
                                         borderBottom = "#dddddd")
          },
          columns = list(
            variable = colDef(name = "",
                             align = "left",
                             width = 180)
          )
)

# knitr::kable(tbl,
#              format = "html",
#              booktabs = TRUE) %>%
#   # add an hline before the total
#   #kableExtra::row_spec(nrow(tbl) - 1, hline_after = TRUE) %>%  #kableExtra::row_spec(nrow(tbl), bold = TRUE) %>%
# kableExtra::row_spec(0, bold = TRUE) %>%
# #kableExtra::column_spec(1, border_right = TRUE, bold  = TRUE) %>%
# kableExtra::group_rows("Spontaneous vaginal", start_row = 1, end_row = 3) %>%
# kableExtra::group_rows("Assited vaginal", start_row = 4, end_row = 6) %>%
# kableExtra::group_rows("Caesarean section", start_row = 7, end_row = 9) %>%
# kableExtra::kable_styling(font_size = 12, full_width = FALSE,
#                             bootstrap_options = c("hover","condensed", "responsive"))
```