---
params:
  year1: 2010
  year2: 2019
fig.caption: yes
---
  
```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE, 
  message = FALSE, 
  warning = FALSE,
  fig.pos = "h", 
  out.extra = "",
  out.width = "90%",
  fig.align = "center",
  eval.after = "fig.cap")

pckgs <- c("tidyverse", "knitr", "scales","janitor",
           "kableExtra", "formattable","data.table",
           "forcats","lubridate", "anytime","reactable",
           "htmltools", "plotly","purrr")
# for (pp in `pckgs`)00/870 { if (!require(pp)) install.packages(pp); library(pp, character.only = T)  }

if (!require("pacman")) install.packages("pacman")
pacman::p_load(pckgs, character.only = TRUE)
## data input

# add data
dta <- readr::read_csv("data/NSAtleePD.csv") %>%
  setDT() %>%
  .[data.table::between(BrthYear, params$year1, params$year2), ]
```

<!-- # (PART) Deliveries and Births {-} -->
  
# Maternal Health Outcomes {#section-5}
  
## Labour induction by parity and year, Nova Scotia, `r paste0(params$year1,"-",params$year2)` {#section-51}
  
```{r plot51, fig.cap= paste('Labour induction by parity and year, Nova Scotia,', paste0(params$year1,"-",params$year2))}

dta1 <- dta[, .(BrthYear,
                # pre-pregnancy diabetes
                 R014_00900,
                 R014_01000,
                 R014_01100,
                 R014_01200,
                 MO240,
                 MO241,
                 MO242,
                 MO243,
                 MO245, MO24501, MO24502, MO24504,
                 MO246, MO24601, MO24602, MO24604,
                 MO247, MO24701, MO24704,
                 ME10, ME11, ME13, ME14,
                 MH360,
                 MG632,
                 IP701,
                # gestational diabetes
                 R014_01300,
                 R014_01400,
                 MO244, MO24401, MO24404, MO24409,
                 MO248, MO24801, MO24802, MO24803, MO24804,
                 MO249,
                 IP700
                )] %>%
  .[, idx_244 := rowSums(.SD, na.rm = TRUE), .SDcols = names(dta)[which(grepl("MO2440.", names(dta)))]] %>%
  .[, idx_245 := rowSums(.SD, na.rm = TRUE), .SDcols = names(dta)[which(grepl("MO2450.", names(dta)))]] %>%
  .[, idx_246 := rowSums(.SD, na.rm = TRUE), .SDcols = names(dta)[which(grepl("MO2460.", names(dta)))]] %>%
  .[, idx_247 := rowSums(.SD, na.rm = TRUE), .SDcols = names(dta)[which(grepl("MO2470.", names(dta)))]] %>%
  .[, idx_248 := rowSums(.SD, na.rm = TRUE), .SDcols = names(dta)[which(grepl("MO2480.", names(dta)))]] %>%
  .[, `:=` (predm_nsapd = fcase(
    R014_00900 >=1 | R014_01000 >=1 | 
    R014_01100 >=1 | R014_01200 >=1, 1,
   default = 0
    ),
   predm_icd = fcase(
    MO240 >=1 | MO241 >=1 | MO242 >=1 | MO243 >=1 |
    MO245 >=1 | idx_245 >=1 |
    MO246 >=1 | idx_246 >=1 |
    MO247 >=1 | idx_247 >=1 |
    ME10 >=1 | ME11 >=1 | ME13 >=1 | ME14 >=1 |
    MH360 >=1 | MG632 >=1 | IP701  >=1, 1,
   default = 0
    ),
   gdm_nsapd = fcase(
    R014_01300 >=1 | R014_01400 >=1, 1,
   default = 0
    ),
   gdm_icd = fcase(
    MO244 >=1 | MO249 >=1 | MO248 >=1 |
    idx_244 >=1 | idx_248 >=1 |
    IP700  >=1, 1,
   default = 0
    ))] %>%
  .[, `:=` (prexdiab = fcase(
    predm_nsapd %in% 1 | predm_icd %in% 1, 1,
   default = 0
    ),
   gestdiab = fcase(
     gdm_nsapd %in% 1 | gdm_icd %in% 1, 1,
     default = 0
   ))] %>%
  #unique() %>%
  .[, `:=` (count = .N), by = list(BrthYear, gestdiab)] %>%
  .[, `:=` (total = .N), by = list(BrthYear)] %>%
  .[, `:=` (rate = count/total)] %>%
  unique()

# age category using UNICODE codes
dta1 <- dta1[, `:=` (gestdiab = factor(gestdiab,
                                  levels = c(0,1),
                                  labels = c("No",
                                             "Gestational diabetes")))]

## getting color for plotting

pal <- c("No" = "#1b9e77",
         "Gestational diabetes" = "#d95f02")

plot <- ggplot(data = dta1[!gestdiab %in% c("No")],
               aes(x = BrthYear,
                   y = rate,
                   group = gestdiab,
                   colour = gestdiab,
                   fill = gestdiab,
                   #shape = gestdiab,
                   text =paste(gestdiab,
                               "<br>Year:", BrthYear,
                               "<br># births:", scales::comma(total, accuracy = 1),
                               "<br>% of deliveries:" , scales::percent(rate, accuracy = .1)
                   )
               )
) +
  geom_point(aes(color = gestdiab, fill = gestdiab)) +
  geom_line(aes(color = gestdiab, fill = gestdiab)) +
  # Specifying colors by age group
  scale_color_manual(name = "",
                     values = pal,
                     limits = names(pal)) +
  scale_fill_manual(name = "",
                    values = pal,
                    limits = names(pal)) +
  scale_shape_manual(name = "",
                     values = c(15, 16, 17)) +
  # scale_x_date(date_breaks = "2 years",
  #              date_labels = "%Y") +
  scale_x_continuous(#trans = 'log2',
    breaks = pretty) +
  scale_y_continuous(#trans = 'log2',
    breaks = pretty,
    labels = scales::percent_format(accuracy = 1),
    limits = c(0, NA)) +
  labs(y = '% of deliveries', x = '') +
  theme_minimal()

plotly::ggplotly(plot, tooltip = c("text")) %>%
  plotly::style(hoverlabel = list(
    bgcolor  = "black",
    bordercolor = "transparent",
    font = list(
      color = "white",
      size = 14,
      face = "bold"
    )
  )) %>%
  plotly::layout(legend = list(orientation = "h", y = -0.25, x = 0.5,
                               xanchor = "center"),
                 font = list(face = "bold", size = 10),
                 showlegend = FALSE)
```

```{r tbl51}
tbl <- dta[, .(BrthYear,
                # pre-pregnancy diabetes
                 R014_00900,
                 R014_01000,
                 R014_01100,
                 R014_01200,
                 MO240,
                 MO241,
                 MO242,
                 MO243,
                 MO245, MO24501, MO24502, MO24504,
                 MO246, MO24601, MO24602, MO24604,
                 MO247, MO24701, MO24704,
                 ME10, ME11, ME13, ME14,
                 MH360,
                 MG632,
                 IP701,
                # gestational diabetes
                 R014_01300,
                 R014_01400,
                 MO244, MO24401, MO24404, MO24409,
                 MO248, MO24801, MO24802, MO24803, MO24804,
                 MO249,
                 IP700
                )] %>%
  .[, idx_244 := rowSums(.SD, na.rm = TRUE), .SDcols = names(dta)[which(grepl("MO2440.", names(dta)))]] %>%
  .[, idx_245 := rowSums(.SD, na.rm = TRUE), .SDcols = names(dta)[which(grepl("MO2450.", names(dta)))]] %>%
  .[, idx_246 := rowSums(.SD, na.rm = TRUE), .SDcols = names(dta)[which(grepl("MO2460.", names(dta)))]] %>%
  .[, idx_247 := rowSums(.SD, na.rm = TRUE), .SDcols = names(dta)[which(grepl("MO2470.", names(dta)))]] %>%
  .[, idx_248 := rowSums(.SD, na.rm = TRUE), .SDcols = names(dta)[which(grepl("MO2480.", names(dta)))]] %>%
  .[, `:=` (predm_nsapd = fcase(
    R014_00900 >=1 | R014_01000 >=1 | 
    R014_01100 >=1 | R014_01200 >=1, 1,
   default = 0
    ),
   predm_icd = fcase(
    MO240 >=1 | MO241 >=1 | MO242 >=1 | MO243 >=1 |
    MO245 >=1 | idx_245 >=1 |
    MO246 >=1 | idx_246 >=1 |
    MO247 >=1 | idx_247 >=1 |
    ME10 >=1 | ME11 >=1 | ME13 >=1 | ME14 >=1 |
    MH360 >=1 | MG632 >=1 | IP701  >=1, 1,
   default = 0
    ),
   gdm_nsapd = fcase(
    R014_01300 >=1 | R014_01400 >=1, 1,
   default = 0
    ),
   gdm_icd = fcase(
    MO244 >=1 | MO249 >=1 | MO248 >=1 |
    idx_244 >=1 | idx_248 >=1 |
    IP700  >=1, 1,
   default = 0
    ))] %>%
  .[, `:=` (prexdiab = fcase(
    predm_nsapd %in% 1 | predm_icd %in% 1, 1,
   default = 0
    ),
   gestdiab = fcase(
     gdm_nsapd %in% 1 | gdm_icd %in% 1, 1,
     default = 0
   ))] %>%
  #unique() %>%
  .[, `:=` (count = .N), by = list(BrthYear, gestdiab)] %>%
  .[, `:=` (total = .N), by = list(BrthYear)] %>%
  .[, `:=` (rate = 100*count/total)] %>%
  melt(measure.vars = c("rate", "total", "count")) %>%
  .[!tolower(variable) %in% "count" &
    !gestdiab %in% 0] %>%
  .[, `:=` (gestdiab = fifelse(tolower(variable) %in% "total",
                          0, gestdiab))] %>%
  .[, .(BrthYear, gestdiab, value)] %>%
  unique() %>%
  dcast(gestdiab ~ BrthYear, value.var = "value")

coln <- names(tbl)
coln[1] <- ""

tbl <- as.data.frame(sapply(tbl, sprintf, fmt = c("%.0f", "%1.1f%%")))
names(tbl) <- coln

tbl[,1] <- c("# deliveries", "Gestational diabetes")

knitr::kable(tbl,
             format = "html",
             booktabs = TRUE) %>%
  # add an hline before the total
  #kableExtra::row_spec(nrow(tbl) - 1, hline_after = TRUE) %>%  #kableExtra::row_spec(nrow(tbl), bold = TRUE) %>%
kableExtra::row_spec(0, bold = TRUE) %>%
#kableExtra::column_spec(1, border_right = TRUE, bold  = TRUE) %>%
kableExtra::kable_styling(font_size = 12, full_width = FALSE,
                            bootstrap_options = c("hover","condensed", "responsive")) %>%
  kableExtra::add_footnote(label = c("Among women without pre-existing diabetes.",
                                     "Note: Diabetes mellitus first detected in pregnancy as recorded in the medical record. Please note that the criteria for the diagnosis of gestational diabetes were revised by Diabetes Canada (formerly the Canadian Diabetes Association) in 2013. Therefore, the rates of gestational diabetes were expected to increase as the new criteria are adopted across Nova Scotia, starting approximately in late 2014."))
```

## Gestational hypertension by year, Nova Scotia, `r paste0(params$year1,"-",params$year2)` {#section-52}
  
```{r plot52, fig.cap= paste('Gestational hypertension by year, Nova Scotia,', paste0(params$year1,"-",params$year2))}

dta1 <- dta[, .(BrthYear,
                # pre-eclampsia
                R014_00600, R014_00800,
                MO14, MO11, MO15,
                # gh any
                MO13,
                R014_00500, R014_01900,
                # pre htn
                R014_00700, R014_00800, 
                MO10,
                MI10, MI11, MI12, MI13, MI15, MCHTD
                )] %>%
  .[, `:=` (preeclamp = fcase(
    R014_00600 >=1 | R014_00800 >=1 |
    MO14 >=1 | MO11 >=1 | MO15 >=1, 1,
   default = 0
    ),
   ghany = fcase(
    MO13 >=1 | MO14 >=1 | MO15 >=1 |
    R014_00500 >=1 | R014_01900 >=1, 1,
   default = 0
   ),
   prehtn = fcase(
    R014_00700 >=1 | R014_00800 >=1 | 
    MO10 >=1 | MO11 >=1 |
    MI10 >=1 | MI11 >=1 | MI12 >=1 | MI13 >=1 | MI15 >=1 | MCHTD >=1, 1,
   default = 0
    ))] %>%
  .[, `:=` (ghinsig = fcase(
    ghany >=1 & !preeclamp >=1 & !prehtn >=1 , 1,
   default = 0
    ))] %>%
  .[, `:=` (ghcat = fcase(
    !preeclamp >=1 & !ghinsig >=1, 1,
    !preeclamp >=1 & ghinsig >=1, 2,
    preeclamp >=1 , 3,
   default = 0
    ))] %>%
  # remove prehtn
  .[!prehtn >= 1] %>%
  #unique() %>%
  .[, `:=` (count = .N), by = list(BrthYear, ghcat)] %>%
  .[, `:=` (total = .N), by = list(BrthYear)] %>%
  .[, `:=` (rate = count/total)] %>%
  .[!ghcat %in% 1, .(BrthYear, ghcat, count, total, rate)] %>% 
  unique()

temp <- unique(dta1) %>%
  .[,`:=` (ghcat = 1,
           count = sum(count),
           total = total,
           rate = sum(rate)), by = list(BrthYear)]


dta1 <- unique(rbind(dta1, temp))[,
            `:=` (ghcat = factor(ghcat,
                                   levels = c(1,2,3),
                                   labels = c("GH-total",
                                              "GH-no sig proteinuria",
                                              "GH-sig proteinuria")))]

## getting color for plotting

pal <- c("GH-total" = "#1b9e77",
         "GH-sig proteinuria" = "#d95f02",
         "GH-no sig proteinuria" = "#7570b3")

plot <- ggplot(data = dta1,
               aes(x = BrthYear,
                   y = rate,
                   group = ghcat,
                   colour = ghcat,
                   fill = ghcat,
                   shape = ghcat,
                   text =paste(ghcat,
                               "<br>Year:", BrthYear,
                               "<br># births:", scales::comma(total, accuracy = 1),
                               "<br>% of deliveries:" , scales::percent(rate, accuracy = .1)
                   )
               )
) +
  geom_point(aes(color = ghcat, fill = ghcat, shape = ghcat)) +
  geom_line(aes(color = ghcat, fill = ghcat)) +
  # Specifying colors by age group
  scale_color_manual(name = "",
                     values = pal,
                     limits = names(pal)) +
  scale_fill_manual(name = "",
                    values = pal,
                    limits = names(pal)) +
  scale_shape_manual(name = "",
                     values = c(15, 16, 17)) +
  # scale_x_date(date_breaks = "2 years",
  #              date_labels = "%Y") +
  scale_x_continuous(#trans = 'log2',
    breaks = pretty) +
  scale_y_continuous(#trans = 'log2',
    breaks = pretty,
    labels = scales::percent_format(accuracy = 1),
    limits = c(0, NA)) +
  labs(y = '% of deliveries', x = '') +
  theme_minimal()

plotly::ggplotly(plot, tooltip = c("text")) %>%
  plotly::style(hoverlabel = list(
    bgcolor  = "black",
    bordercolor = "transparent",
    font = list(
      color = "white",
      size = 14,
      face = "bold"
    )
  )) %>%
  plotly::layout(legend = list(orientation = "h", y = -0.25, x = 0.5,
                               xanchor = "center"),
                 font = list(face = "bold", size = 10),
                 showlegend = TRUE)
```

```{r tbl52}
tbl <- dta[, .(BrthYear,
                # pre-eclampsia
                R014_00600, R014_00800,
                MO14, MO11, MO15,
                # gh any
                MO13,
                R014_00500, R014_01900,
                # pre htn
                R014_00700, R014_00800, 
                MO10,
                MI10, MI11, MI12, MI13, MI15, MCHTD
                )] %>%
  .[, `:=` (preeclamp = fcase(
    R014_00600 >=1 | R014_00800 >=1 |
    MO14 >=1 | MO11 >=1 | MO15 >=1, 1,
   default = 0
    ),
   ghany = fcase(
    MO13 >=1 | MO14 >=1 | MO15 >=1 |
    R014_00500 >=1 | R014_01900 >=1, 1,
   default = 0
   ),
   prehtn = fcase(
    R014_00700 >=1 | R014_00800 >=1 | 
    MO10 >=1 | MO11 >=1 |
    MI10 >=1 | MI11 >=1 | MI12 >=1 | MI13 >=1 | MI15 >=1 | MCHTD >=1, 1,
   default = 0
    ))] %>%
  .[, `:=` (ghinsig = fcase(
    ghany >=1 & !preeclamp >=1 & !prehtn >=1 , 1,
   default = 0
    ))] %>%
  .[, `:=` (ghcat = fcase(
    !preeclamp >=1 & !ghinsig >=1, 1,
    !preeclamp >=1 & ghinsig >=1, 2,
    preeclamp >=1 , 3,
   default = 0
    ))] %>%
  # remove prehtn
  .[!prehtn >= 1] %>%
  #unique() %>%
  .[, `:=` (count = .N), by = list(BrthYear, ghcat)] %>%
  .[, `:=` (total = .N), by = list(BrthYear)] %>%
  .[, `:=` (rate = 100*count/total)] %>%
  .[!ghcat %in% 1, .(BrthYear, ghcat, count, total, rate)] %>% 
  unique()

temp <- unique(tbl) %>%
  .[,`:=` (ghcat = 4,
           count = sum(count),
           total = total,
           rate = sum(rate)), by = list(BrthYear)]


tbl <- unique(rbind(tbl, temp)) %>%
  melt(measure.vars = c("count", "total", "rate")) %>%
  .[!tolower(variable) %in% "count"] %>%
  .[, `:=` (ghcat = fifelse(tolower(variable) %in% "total",
                          0, ghcat))] %>%
  .[, .(BrthYear, ghcat, value)] %>%
  unique() %>%
  dcast(ghcat ~ BrthYear, value.var = "value")

coln <- names(tbl)
coln[1] <- ""

tbl <- as.data.frame(sapply(tbl, sprintf, fmt = c("%.0f", "%1.1f%%", "%1.1f%%", "%1.1f%%")))

names(tbl) <- coln

tbl[,1] <- c("# deliveries", "GH-no sig proteinuria", "GH-sig proteinuria", "GH-total")

knitr::kable(tbl,
             format = "html",
             booktabs = TRUE) %>%
  # add an hline before the total
  #kableExtra::row_spec(nrow(tbl) - 1, hline_after = TRUE) %>%  #kableExtra::row_spec(nrow(tbl), bold = TRUE) %>%
kableExtra::row_spec(0, bold = TRUE) %>%
#kableExtra::column_spec(1, border_right = TRUE, bold  = TRUE) %>%
kableExtra::kable_styling(font_size = 12, full_width = FALSE,
                            bootstrap_options = c("hover","condensed", "responsive")) %>%
  kableExtra::add_footnote(label = c("Among women without pre-existing hypertension.",
                                     "Note: Gestational hypertension is hypertension that is first detected after the 20th week of gestation. Gestational hypertension with significant proteinuria includes those cases denoted as such; severe pre-eclampsia; HELLP syndrome (Hemolysis, Elevated Liver Enzymes, Low Platelets); and eclampsia."))
```

## Pre-eclampsia by year, Nova Scotia, `r paste0(params$year1,"-",params$year2)` {#section-53}
  
```{r plot53, fig.cap= paste('Pre-eclampsia by year, Nova Scotia,', paste0(params$year1,"-",params$year2))}

dta1 <- dta[, .(BrthYear,
        # pre-existing hypertension
        MO10, MO11,
        MI10, MI11, MI12 , MI13, MI15, 
        R014_00700, R014_00800, MCHTD,
        # HELLP
        MO14201, MO14202, MO14204,
        MPIHTC, R014_01900,
        # severe PIH
        MO14101, MO14102,
        R014_00600, R014_01800,
        # eclampsia
        MECLP, MO15,
        # gestational hypertension
        MO13, MO14,
        R014_00500, MPIHT,
        # anti hypertensive
        R003_00700, R003_01700, R003_02300, R004_00400,
        # any hypertension
        MO16, R014_00550, IINDUCTN, INDICCS1, DLADMREA, MMSO4,
        # proteinuria
        MO12
)] %>%
  .[, `:=` (prehtn = fcase(
    MO10 >=1 | MO11 >=1 | 
      MI10 >=1 | MI11 >=1 | MI12 >=1 | MI13 >=1 | MI15 >=1 |
      R014_00700 >=1 | R014_00800 >=1 | MCHTD >=1, 1, 
    default = 0
  ),
  hellp = fcase(
    MO14201 >=1 | MO14202 >=1 | MO14204 >=1 | 
      tolower(MPIHTC) %in% "hlp" |
      R014_01900 >=1, 1,
    default = 0
  ),
  eclamp = fcase(
    MECLP >=1 | MO15 >=1, 1,
    default = 0
  ),
  antihtn = fcase(
    R003_00700 >=1 | R003_01700 >=1 | R003_02300 >=1 |
      R004_00400 >=1, 1,
    default = 0
  ))] %>%
  .[, `:=` (sevpih = fcase(
    hellp >=1 | MO14101>=1 | MO14102>=1 |
      R014_00600>=1 | R014_01800 >=1 | tolower(MPIHTC) %in% "sev", 1,
    default = 0
  ),
  gesthtn = fcase(
    MO11>=1 | MO13>=1 | MO14>=1 |
      R014_00500 >=1 | R014_00800>=1 | MPIHT >=1 | 
      hellp >=1 | MO14101>=1 | MO14102>=1 |
      R014_00600>=1 | R014_01800 >=1 | tolower(MPIHTC) %in% "sev" | eclamp >=1 , 1,
    default = 0
  ))] %>%
  .[, `:=` (anyhtn = fcase(
    prehtn >=1 | gesthtn >=1 | MO16 >=1 |
      R014_00550>=1 | antihtn >=1 | 
      tolower(INDICCS1) %in% "htd" | IINDUCTN %in% 12 |
      DLADMREA %in% c(8,17) |
      MMSO4 >= 1, 1,
    default = 0
  ),
  protein = fcase(
    eclamp >=1 |
      R014_00800 >=1 | R014_00600 >=1 | R014_01800 >=1 |
      hellp >=1 | MO14 >=1 | MO12 >=1, 1,
    default = 0
  ))] %>%
  .[, `:=` (preeclamp = fcase(
    anyhtn >=1 & protein >=1, 1,
    default = 0
  ))] %>%
  #unique() %>%
  .[, `:=` (count = .N), by = list(BrthYear, preeclamp)] %>%
  .[, `:=` (total = .N), by = list(BrthYear)] %>%
  .[, `:=` (rate = count/total)] %>%
  unique()


dta1 <- dta1[, `:=` (preeclamp = factor(preeclamp,
                                   levels = c(0,1),
                                   labels = c("No",
                                              "Pre-eclampsia")))]

## getting color for plotting

pal <- c("No" = "#1b9e77",
         "Pre-eclampsia" = "#d95f02")

plot <- ggplot(data = dta1[!tolower(preeclamp) %in% "no"],
               aes(x = BrthYear,
                   y = rate,
                   group = preeclamp,
                   colour = preeclamp,
                   fill = preeclamp,
                   #shape = preeclamp,
                   text =paste(preeclamp,
                               "<br>Year:", BrthYear,
                               "<br># births:", scales::comma(total, accuracy = 1),
                               "<br>% of deliveries:" , scales::percent(rate, accuracy = .1)
                   )
               )
) +
  geom_point(aes(color = preeclamp, fill = preeclamp)) +
  geom_line(aes(color = preeclamp, fill = preeclamp)) +
  # Specifying colors by age group
  scale_color_manual(name = "",
                     values = pal,
                     limits = names(pal)) +
  scale_fill_manual(name = "",
                    values = pal,
                    limits = names(pal)) +
  # scale_shape_manual(name = "",
  #                    values = c(15, 16, 17)) +
  # scale_x_date(date_breaks = "2 years",
  #              date_labels = "%Y") +
  scale_x_continuous(#trans = 'log2',
    breaks = pretty) +
  scale_y_continuous(#trans = 'log2',
    breaks = pretty,
    labels = scales::percent_format(accuracy = 1),
    limits = c(0, NA)) +
  labs(y = '% of deliveries', x = '') +
  theme_minimal()

plotly::ggplotly(plot, tooltip = c("text")) %>%
  plotly::style(hoverlabel = list(
    bgcolor  = "black",
    bordercolor = "transparent",
    font = list(
      color = "white",
      size = 14,
      face = "bold"
    )
  )) %>%
  plotly::layout(legend = list(orientation = "h", y = -0.25, x = 0.5,
                               xanchor = "center"),
                 font = list(face = "bold", size = 10),
                 showlegend = FALSE)
```

```{r tbl53}
tbl <- dta[, .(BrthYear,
        # pre-existing hypertension
        MO10, MO11,
        MI10, MI11, MI12 , MI13, MI15, 
        R014_00700, R014_00800, MCHTD,
        # HELLP
        MO14201, MO14202, MO14204,
        MPIHTC, R014_01900,
        # severe PIH
        MO14101, MO14102,
        R014_00600, R014_01800,
        # eclampsia
        MECLP, MO15,
        # gestational hypertension
        MO13, MO14,
        R014_00500, MPIHT,
        # anti hypertensive
        R003_00700, R003_01700, R003_02300, R004_00400,
        # any hypertension
        MO16, R014_00550, IINDUCTN, INDICCS1, DLADMREA, MMSO4,
        # proteinuria
        MO12
)] %>%
  .[, `:=` (prehtn = fcase(
    MO10 >=1 | MO11 >=1 | 
      MI10 >=1 | MI11 >=1 | MI12 >=1 | MI13 >=1 | MI15 >=1 |
      R014_00700 >=1 | R014_00800 >=1 | MCHTD >=1, 1, 
    default = 0
  ),
  hellp = fcase(
    MO14201 >=1 | MO14202 >=1 | MO14204 >=1 | 
      tolower(MPIHTC) %in% "hlp" |
      R014_01900 >=1, 1,
    default = 0
  ),
  eclamp = fcase(
    MECLP >=1 | MO15 >=1, 1,
    default = 0
  ),
  antihtn = fcase(
    R003_00700 >=1 | R003_01700 >=1 | R003_02300 >=1 |
      R004_00400 >=1, 1,
    default = 0
  ))] %>%
  .[, `:=` (sevpih = fcase(
    hellp >=1 | MO14101>=1 | MO14102>=1 |
      R014_00600>=1 | R014_01800 >=1 | tolower(MPIHTC) %in% "sev", 1,
    default = 0
  ),
  gesthtn = fcase(
    MO11>=1 | MO13>=1 | MO14>=1 |
      R014_00500 >=1 | R014_00800>=1 | MPIHT >=1 | 
      hellp >=1 | MO14101>=1 | MO14102>=1 |
      R014_00600>=1 | R014_01800 >=1 | tolower(MPIHTC) %in% "sev" | eclamp >=1 , 1,
    default = 0
  ))] %>%
  .[, `:=` (anyhtn = fcase(
    prehtn >=1 | gesthtn >=1 | MO16 >=1 |
      R014_00550>=1 | antihtn >=1 | 
      tolower(INDICCS1) %in% "htd" | IINDUCTN %in% 12 |
      DLADMREA %in% c(8,17) |
      MMSO4 >= 1, 1,
    default = 0
  ),
  protein = fcase(
    eclamp >=1 |
      R014_00800 >=1 | R014_00600 >=1 | R014_01800 >=1 |
      hellp >=1 | MO14 >=1 | MO12 >=1, 1,
    default = 0
  ))] %>%
  .[, `:=` (preeclamp = fcase(
    anyhtn >=1 & protein >=1, 1,
    default = 0
  ))] %>%
  #unique() %>%
  .[, `:=` (count = .N), by = list(BrthYear, preeclamp)] %>%
  .[, `:=` (total = .N), by = list(BrthYear)] %>%
  .[, `:=` (rate = 100*count/total)] %>%
  unique() %>%
  melt(measure.vars = c("count", "total", "rate")) %>%
  .[!tolower(variable) %in% "count" &
    !preeclamp %in% 0] %>%
  .[, `:=` (preeclamp = fifelse(tolower(variable) %in% "total",
                          0, preeclamp))] %>%
  .[, .(BrthYear, preeclamp, value)] %>%
  unique() %>%
  dcast(preeclamp ~ BrthYear, value.var = "value")

coln <- names(tbl)
coln[1] <- ""

tbl <- as.data.frame(sapply(tbl, sprintf, fmt = c("%.0f", "%1.1f%%")))

names(tbl) <- coln

tbl[,1] <- c("# deliveries", "Pre-eclampsia")

knitr::kable(tbl,
             format = "html",
             booktabs = TRUE) %>%
  # add an hline before the total
  #kableExtra::row_spec(nrow(tbl) - 1, hline_after = TRUE) %>%  #kableExtra::row_spec(nrow(tbl), bold = TRUE) %>%
kableExtra::row_spec(0, bold = TRUE) %>%
#kableExtra::column_spec(1, border_right = TRUE, bold  = TRUE) %>%
kableExtra::kable_styling(font_size = 12, full_width = FALSE,
                            bootstrap_options = c("hover","condensed", "responsive")) %>%
  kableExtra::add_footnote(label = c("Note: Pre-eclampsia includes women coded as having gestational hypertension with significant proteinuria, moderate or severe pre-eclampsia, HELLP syndrome (Hemolysis, Elevated Liver Enzymes, Low Platelets), eclampsia, or pre-existing hypertension with superimposed proteinuria."))
```

## Placenta previa by year, Nova Scotia, `r paste0(params$year1,"-",params$year2)` {#section-54}
  
```{r plot54, fig.cap= paste('Placenta previa by year, Nova Scotia,', paste0(params$year1,"-",params$year2))}

dta1 <- dta[, .(BrthYear,
        MO44
)] %>%
  .[, `:=` (plaprev = fcase(
    MO44 >=1, 1, 
    default = 0
  ))] %>%
  #unique() %>%
  .[, `:=` (count = .N), by = list(BrthYear, plaprev)] %>%
  .[, `:=` (total = .N), by = list(BrthYear)] %>%
  .[, `:=` (rate = count/total)] %>%
  unique()


dta1 <- dta1[, `:=` (plaprev = factor(plaprev,
                                   levels = c(0,1),
                                   labels = c("No",
                                              "Placenta previa")))]

## getting color for plotting

pal <- c("No" = "#1b9e77",
         "Placenta previa" = "#d95f02")

plot <- ggplot(data = dta1[!tolower(plaprev) %in% "no"],
               aes(x = BrthYear,
                   y = rate,
                   group = plaprev,
                   colour = plaprev,
                   fill = plaprev,
                   #shape = plaprev,
                   text =paste(plaprev,
                               "<br>Year:", BrthYear,
                               "<br># births:", scales::comma(total, accuracy = 1),
                               "<br>% of deliveries:" , scales::percent(rate, accuracy = .1)
                   )
               )
) +
  geom_point(aes(color = plaprev, fill = plaprev)) +
  geom_line(aes(color = plaprev, fill = plaprev)) +
  # Specifying colors by age group
  scale_color_manual(name = "",
                     values = pal,
                     limits = names(pal)) +
  scale_fill_manual(name = "",
                    values = pal,
                    limits = names(pal)) +
  # scale_shape_manual(name = "",
  #                    values = c(15, 16, 17)) +
  # scale_x_date(date_breaks = "2 years",
  #              date_labels = "%Y") +
  scale_x_continuous(#trans = 'log2',
    breaks = pretty) +
  scale_y_continuous(#trans = 'log2',
    breaks = pretty,
    labels = scales::percent_format(accuracy = .1),
    limits = c(0, NA)) +
  labs(y = '% of deliveries', x = '') +
  theme_minimal()

plotly::ggplotly(plot, tooltip = c("text")) %>%
  plotly::style(hoverlabel = list(
    bgcolor  = "black",
    bordercolor = "transparent",
    font = list(
      color = "white",
      size = 14,
      face = "bold"
    )
  )) %>%
  plotly::layout(legend = list(orientation = "h", y = -0.25, x = 0.5,
                               xanchor = "center"),
                 font = list(face = "bold", size = 10),
                 showlegend = FALSE)
```

```{r tbl54}
tbl <- dta[, .(BrthYear,
        MO44
)] %>%
  .[, `:=` (plaprev = fcase(
    MO44 >=1, 1, 
    default = 0
  ))] %>%
  #unique() %>%
  .[, `:=` (count = .N), by = list(BrthYear, plaprev)] %>%
  .[, `:=` (total = .N), by = list(BrthYear)] %>%
  .[, `:=` (rate = 100*count/total)] %>%
  unique() %>%
  melt(measure.vars = c("count", "total", "rate")) %>%
  .[!tolower(variable) %in% "count" &
    !plaprev %in% 0] %>%
  .[, `:=` (plaprev = fifelse(tolower(variable) %in% "total",
                          0, plaprev))] %>%
  .[, .(BrthYear, plaprev, value)] %>%
  unique() %>%
  dcast(plaprev ~ BrthYear, value.var = "value")

coln <- names(tbl)
coln[1] <- ""

tbl <- as.data.frame(sapply(tbl, sprintf, fmt = c("%.0f", "%1.1f%%")))

names(tbl) <- coln

tbl[,1] <- c("# deliveries", "Placenta previa")

knitr::kable(tbl,
             format = "html",
             booktabs = TRUE) %>%
  # add an hline before the total
  #kableExtra::row_spec(nrow(tbl) - 1, hline_after = TRUE) %>%  #kableExtra::row_spec(nrow(tbl), bold = TRUE) %>%
kableExtra::row_spec(0, bold = TRUE) %>%
#kableExtra::column_spec(1, border_right = TRUE, bold  = TRUE) %>%
kableExtra::kable_styling(font_size = 12, full_width = FALSE,
                            bootstrap_options = c("hover","condensed", "responsive")) %>%
  kableExtra::add_footnote(label = c("Note: Placenta previa is diagnosed when the placenta entirely or partially covers the opening of the uterus (cervix). The diagnosis is not made on ultrasound alone and must be confirmed clinically"))
```

## Placental abruption by year, Nova Scotia, `r paste0(params$year1,"-",params$year2)` {#section-55}
  
```{r plot55, fig.cap= paste('Placental abruption by year, Nova Scotia,', paste0(params$year1,"-",params$year2))}

dta1 <- dta[, .(BrthYear,
        MO45
)] %>%
  .[, `:=` (plaabrp = fcase(
    MO45 >=1, 1, 
    default = 0
  ))] %>%
  #unique() %>%
  .[, `:=` (count = .N), by = list(BrthYear, plaabrp)] %>%
  .[, `:=` (total = .N), by = list(BrthYear)] %>%
  .[, `:=` (rate = count/total)] %>%
  unique()


dta1 <- dta1[, `:=` (plaabrp = factor(plaabrp,
                                   levels = c(0,1),
                                   labels = c("No",
                                              "Placenta abruption")))]

## getting color for plotting

pal <- c("No" = "#1b9e77",
         "Placenta abruption" = "#d95f02")

plot <- ggplot(data = dta1[!tolower(plaabrp) %in% "no"],
               aes(x = BrthYear,
                   y = rate,
                   group = plaabrp,
                   colour = plaabrp,
                   fill = plaabrp,
                   #shape = plaabrp,
                   text =paste(plaabrp,
                               "<br>Year:", BrthYear,
                               "<br># births:", scales::comma(total, accuracy = 1),
                               "<br>% of deliveries:" , scales::percent(rate, accuracy = .1)
                   )
               )
) +
  geom_point(aes(color = plaabrp, fill = plaabrp)) +
  geom_line(aes(color = plaabrp, fill = plaabrp)) +
  # Specifying colors by age group
  scale_color_manual(name = "",
                     values = pal,
                     limits = names(pal)) +
  scale_fill_manual(name = "",
                    values = pal,
                    limits = names(pal)) +
  # scale_shape_manual(name = "",
  #                    values = c(15, 16, 17)) +
  # scale_x_date(date_breaks = "2 years",
  #              date_labels = "%Y") +
  scale_x_continuous(#trans = 'log2',
    breaks = pretty) +
  scale_y_continuous(#trans = 'log2',
    breaks = pretty,
    labels = scales::percent_format(accuracy = .1),
    limits = c(0, NA)) +
  labs(y = '% of deliveries', x = '') +
  theme_minimal()

plotly::ggplotly(plot, tooltip = c("text")) %>%
  plotly::style(hoverlabel = list(
    bgcolor  = "black",
    bordercolor = "transparent",
    font = list(
      color = "white",
      size = 14,
      face = "bold"
    )
  )) %>%
  plotly::layout(legend = list(orientation = "h", y = -0.25, x = 0.5,
                               xanchor = "center"),
                 font = list(face = "bold", size = 10),
                 showlegend = FALSE)
```

```{r tbl55}
tbl <- dta[, .(BrthYear,
        MO45
)] %>%
  .[, `:=` (plaabrp = fcase(
    MO45 >=1, 1, 
    default = 0
  ))] %>%
  #unique() %>%
  .[, `:=` (count = .N), by = list(BrthYear, plaabrp)] %>%
  .[, `:=` (total = .N), by = list(BrthYear)] %>%
  .[, `:=` (rate = 100*count/total)] %>%
  unique() %>%
  melt(measure.vars = c("count", "total", "rate")) %>%
  .[!tolower(variable) %in% "count" &
    !plaabrp %in% 0] %>%
  .[, `:=` (plaabrp = fifelse(tolower(variable) %in% "total",
                          0, plaabrp))] %>%
  .[, .(BrthYear, plaabrp, value)] %>%
  unique() %>%
  dcast(plaabrp ~ BrthYear, value.var = "value")

coln <- names(tbl)
coln[1] <- ""

tbl <- as.data.frame(sapply(tbl, sprintf, fmt = c("%.0f", "%1.1f%%")))

names(tbl) <- coln

tbl[,1] <- c("# deliveries", "Placenta previa")

knitr::kable(tbl,
             format = "html",
             booktabs = TRUE) %>%
  # add an hline before the total
  #kableExtra::row_spec(nrow(tbl) - 1, hline_after = TRUE) %>%  #kableExtra::row_spec(nrow(tbl), bold = TRUE) %>%
kableExtra::row_spec(0, bold = TRUE) %>%
#kableExtra::column_spec(1, border_right = TRUE, bold  = TRUE) %>%
kableExtra::kable_styling(font_size = 12, full_width = FALSE,
                            bootstrap_options = c("hover","condensed", "responsive")) %>%
  kableExtra::add_footnote(label = c("Note: Placenta previa is diagnosed when the placenta entirely or partially covers the opening of the uterus (cervix). The diagnosis is not made on ultrasound alone and must be confirmed clinically"))
```

## Perineal laceration among primiparous vaginal deliveries by year, Nova Scotia, `r paste0(params$year1,"-",params$year2)` {#section-56}
  
```{r plot56, fig.cap= paste('Perineal laceration among primiparous vaginal deliveries by year, Nova Scotia,', paste0(params$year1,"-",params$year2))}

dta1 <- dta[, .(BrthYear,
       MO701,
       MO702, MO70201, MO70204, MO70211, MO70214, MO70221, MO70231, MO70234, MO70281, MO70291,
       MO703, MO70301, MO70304,
       DLPARA,
       DMMETHOD
)] %>%
  # Primiparous women
  # vaginal delivery
  .[DLPARA %in% 0 &
    # vaginal
    tolower(DMMETHOD) %in% c("biv", "spt", "vac", "low", "mid")] %>%
  .[, `:=` (lacer = fcase(
    # second degree
    MO701 >=1, 1,
    # 3rd degree
    MO702 >=1 | MO70201 >=1 | MO70204 >=1 | MO70211 >=1 | MO70214 >=1 | MO70221 >=1 | MO70231 >=1 | MO70234 >=1 | MO70281 >=1 | MO70291 >=1 |
    MO703 >=1 | MO70301 >=1 | MO70304 >=1, 2, 
    default = 0
  ))] %>%
  #unique() %>%
  .[, `:=` (count = .N), by = list(BrthYear, lacer)] %>%
  .[, `:=` (total = .N), by = list(BrthYear)] %>%
  .[, `:=` (rate = count/total)] %>%
  unique()


dta1 <- dta1[, `:=` (lacer = factor(lacer,
                                   levels = c(0,1,2),
                                   labels = c("No",
                                              "2\u207f\u1d48 degree laceration",
                                              "3\u02b3\u1d48 or 4\u1d57\u02b0 degree laceration")))]

## getting color for plotting

pal <- c("No" = "#1b9e77",
         "2\u207f\u1d48 degree laceration" = "#d95f02",
         "3\u02b3\u1d48 or 4\u1d57\u02b0 degree laceration" = "#7570b3")

plot <- ggplot(data = dta1[!tolower(lacer) %in% "no"],
               aes(x = BrthYear,
                   y = rate,
                   group = lacer,
                   colour = lacer,
                   fill = lacer,
                   shape = lacer,
                   text =paste(lacer,
                               "<br>Year:", BrthYear,
                               "<br># births:", scales::comma(total, accuracy = 1),
                               "<br>% of deliveries:" , scales::percent(rate, accuracy = .1)
                   )
               )
) +
  geom_point(aes(color = lacer, fill = lacer, shape = lacer)) +
  geom_line(aes(color = lacer, fill = lacer)) +
  # Specifying colors by age group
  scale_color_manual(name = "",
                     values = pal,
                     limits = names(pal)) +
  scale_fill_manual(name = "",
                    values = pal,
                    limits = names(pal)) +
  scale_shape_manual(name = "",
                     values = c(15, 16)) +
  # scale_x_date(date_breaks = "2 years",
  #              date_labels = "%Y") +
  scale_x_continuous(#trans = 'log2',
    breaks = pretty) +
  scale_y_continuous(#trans = 'log2',
    breaks = pretty,
    labels = scales::percent_format(accuracy = .1),
    limits = c(0, NA)) +
  labs(y = '% of primiparous vaginal deliveries', x = '') +
  theme_minimal()

plotly::ggplotly(plot, tooltip = c("text")) %>%
  plotly::style(hoverlabel = list(
    bgcolor  = "black",
    bordercolor = "transparent",
    font = list(
      color = "white",
      size = 14,
      face = "bold"
    )
  )) %>%
  plotly::layout(legend = list(orientation = "h", y = -0.25, x = 0.5,
                               xanchor = "center"),
                 font = list(face = "bold", size = 10),
                 showlegend = TRUE)
```

```{r tbl56}
tbl <- dta[, .(BrthYear,
       MO701,
       MO702, MO70201, MO70204, MO70211, MO70214, MO70221, MO70231, MO70234, MO70281, MO70291,
       MO703, MO70301, MO70304,
       DLPARA,
       DMMETHOD
)] %>%
  # Primiparous women
  # vaginal delivery
  .[DLPARA %in% 0 &
    # vaginal
    tolower(DMMETHOD) %in% c("biv", "spt", "vac", "low", "mid")] %>%
  .[, `:=` (lacer = fcase(
    # second degree
    MO701 >=1, 1,
    # 3rd degree
    MO702 >=1 | MO70201 >=1 | MO70204 >=1 | MO70211 >=1 | MO70214 >=1 | MO70221 >=1 | MO70231 >=1 | MO70234 >=1 | MO70281 >=1 | MO70291 >=1 |
    MO703 >=1 | MO70301 >=1 | MO70304 >=1, 2, 
    default = 0
  ))] %>%
  #unique() %>%
  .[, `:=` (count = .N), by = list(BrthYear, lacer)] %>%
  .[, `:=` (total = .N), by = list(BrthYear)] %>%
  .[, `:=` (rate = 100*count/total)] %>%
  unique() %>%
  melt(measure.vars = c("count", "total", "rate")) %>%
  .[!tolower(variable) %in% "count" &
    !lacer %in% 0] %>%
  .[, `:=` (lacer = fifelse(tolower(variable) %in% "total",
                          0, lacer))] %>%
  .[, .(BrthYear, lacer, value)] %>%
  unique() %>%
  dcast(lacer ~ BrthYear, value.var = "value")

coln <- names(tbl)
coln[1] <- ""

tbl <- as.data.frame(sapply(tbl, sprintf, fmt = c("%.0f", "%1.1f%%", "%1.1f%%")))

names(tbl) <- coln

tbl[,1] <- c("# primiparous vaginal deliveries", "2\u207f\u1d48 degree laceration", "3\u02b3\u1d48 or 4\u1d57\u02b0 degree laceration")

knitr::kable(tbl,
             format = "html",
             booktabs = TRUE) %>%
  # add an hline before the total
  #kableExtra::row_spec(nrow(tbl) - 1, hline_after = TRUE) %>%  #kableExtra::row_spec(nrow(tbl), bold = TRUE) %>%
kableExtra::row_spec(0, bold = TRUE) %>%
#kableExtra::column_spec(1, border_right = TRUE, bold  = TRUE) %>%
kableExtra::kable_styling(font_size = 12, full_width = FALSE,
                            bootstrap_options = c("hover","condensed", "responsive")) %>%
  kableExtra::add_footnote(label = c("Note: Maternal perineal laceration, rupture or tear during delivery involving the pelvic floor, perineal muscles, or vaginal muscles (2$^{nd}$ degree), anal sphincter (3$^{rd}$ degree), or rectal mucosa (4$^{th}$ degree)."), escape = FALSE)
```

## Perineal laceration among multiparous vaginal deliveries by year, Nova Scotia, `r paste0(params$year1,"-",params$year2)` {#section-57}
  
```{r plot57, fig.cap= paste('Perineal laceration among multiparous vaginal deliveries by year, Nova Scotia,', paste0(params$year1,"-",params$year2))}

dta1 <- dta[, .(BrthYear,
       MO701,
       MO702, MO70201, MO70204, MO70211, MO70214, MO70221, MO70231, MO70234, MO70281, MO70291,
       MO703, MO70301, MO70304,
       DLPARA,
       DMMETHOD
)] %>%
  # Primiparous women
  # vaginal delivery
  .[DLPARA >=1 &
    # vaginal
    tolower(DMMETHOD) %in% c("biv", "spt", "vac", "low", "mid")] %>%
  .[, `:=` (lacer = fcase(
    # second degree
    MO701 >=1, 1,
    # 3rd degree
    MO702 >=1 | MO70201 >=1 | MO70204 >=1 | MO70211 >=1 | MO70214 >=1 | MO70221 >=1 | MO70231 >=1 | MO70234 >=1 | MO70281 >=1 | MO70291 >=1 |
    MO703 >=1 | MO70301 >=1 | MO70304 >=1, 2, 
    default = 0
  ))] %>%
  #unique() %>%
  .[, `:=` (count = .N), by = list(BrthYear, lacer)] %>%
  .[, `:=` (total = .N), by = list(BrthYear)] %>%
  .[, `:=` (rate = count/total)] %>%
  unique()


dta1 <- dta1[, `:=` (lacer = factor(lacer,
                                   levels = c(0,1,2),
                                   labels = c("No",
                                              "2\u207f\u1d48 degree laceration",
                                              "3\u1d63\u1d48 or 4\u1d57\u1d78 degree laceration")))]

## getting color for plotting

pal <- c("No" = "#1b9e77",
         "2\u207f\u1d48 degree laceration" = "#d95f02",
         "3\u1d63\u1d48 or 4\u1d57\u1d78 degree laceration" = "#7570b3")

plot <- ggplot(data = dta1[!tolower(lacer) %in% "no"],
               aes(x = BrthYear,
                   y = rate,
                   group = lacer,
                   colour = lacer,
                   fill = lacer,
                   shape = lacer,
                   text =paste(lacer,
                               "<br>Year:", BrthYear,
                               "<br># births:", scales::comma(total, accuracy = 1),
                               "<br>% of deliveries:" , scales::percent(rate, accuracy = .1)
                   )
               )
) +
  geom_point(aes(color = lacer, fill = lacer, shape = lacer)) +
  geom_line(aes(color = lacer, fill = lacer)) +
  # Specifying colors by age group
  scale_color_manual(name = "",
                     values = pal,
                     limits = names(pal)) +
  scale_fill_manual(name = "",
                    values = pal,
                    limits = names(pal)) +
  scale_shape_manual(name = "",
                     values = c(15, 16)) +
  # scale_x_date(date_breaks = "2 years",
  #              date_labels = "%Y") +
  scale_x_continuous(#trans = 'log2',
    breaks = pretty) +
  scale_y_continuous(#trans = 'log2',
    breaks = pretty,
    labels = scales::percent_format(accuracy = .1),
    limits = c(0, NA)) +
  labs(y = '% of multiparous vaginal deliveries', x = '') +
  theme_minimal()

plotly::ggplotly(plot, tooltip = c("text")) %>%
  plotly::style(hoverlabel = list(
    bgcolor  = "black",
    bordercolor = "transparent",
    font = list(
      color = "white",
      size = 14,
      face = "bold"
    )
  )) %>%
  plotly::layout(legend = list(orientation = "h", y = -0.25, x = 0.5,
                               xanchor = "center"),
                 font = list(face = "bold", size = 10),
                 showlegend = TRUE)
```

```{r tbl57}
tbl <- dta[, .(BrthYear,
       MO701,
       MO702, MO70201, MO70204, MO70211, MO70214, MO70221, MO70231, MO70234, MO70281, MO70291,
       MO703, MO70301, MO70304,
       DLPARA,
       DMMETHOD
)] %>%
  # Primiparous women
  # vaginal delivery
  .[DLPARA >= 1 &
    # vaginal
    tolower(DMMETHOD) %in% c("biv", "spt", "vac", "low", "mid")] %>%
  .[, `:=` (lacer = fcase(
    # second degree
    MO701 >=1, 1,
    # 3rd degree
    MO702 >=1 | MO70201 >=1 | MO70204 >=1 | MO70211 >=1 | MO70214 >=1 | MO70221 >=1 | MO70231 >=1 | MO70234 >=1 | MO70281 >=1 | MO70291 >=1 |
    MO703 >=1 | MO70301 >=1 | MO70304 >=1, 2, 
    default = 0
  ))] %>%
  #unique() %>%
  .[, `:=` (count = .N), by = list(BrthYear, lacer)] %>%
  .[, `:=` (total = .N), by = list(BrthYear)] %>%
  .[, `:=` (rate = 100*count/total)] %>%
  unique() %>%
  melt(measure.vars = c("count", "total", "rate")) %>%
  .[!tolower(variable) %in% "count" &
    !lacer %in% 0] %>%
  .[, `:=` (lacer = fifelse(tolower(variable) %in% "total",
                          0, lacer))] %>%
  .[, .(BrthYear, lacer, value)] %>%
  unique() %>%
  dcast(lacer ~ BrthYear, value.var = "value")

coln <- names(tbl)
coln[1] <- ""

tbl <- as.data.frame(sapply(tbl, sprintf, fmt = c("%.0f", "%1.1f%%", "%1.1f%%")))

names(tbl) <- coln

tbl[,1] <- c("# multiparous vaginal deliveries", "2\u207f\u1d48 degree laceration", "3\u1d63\u1d48 or 4\u1d57\u1d78 degree laceration")

knitr::kable(tbl,
             format = "html",
             booktabs = TRUE) %>%
  # add an hline before the total
  #kableExtra::row_spec(nrow(tbl) - 1, hline_after = TRUE) %>%  #kableExtra::row_spec(nrow(tbl), bold = TRUE) %>%
kableExtra::row_spec(0, bold = TRUE) %>%
#kableExtra::column_spec(1, border_right = TRUE, bold  = TRUE) %>%
kableExtra::kable_styling(font_size = 12, full_width = FALSE,
                            bootstrap_options = c("hover","condensed", "responsive")) %>%
  kableExtra::add_footnote(label = c("Note: Maternal perineal laceration, rupture or tear during delivery involving the pelvic floor, perineal muscles, or vaginal muscles (2$^{nd}$ degree), anal sphincter (3$^{rd}$ degree), or rectal mucosa (4$^{th}$ degree)."), escape = FALSE)
```

## Postpartum hemorrhage by year, Nova Scotia, `r paste0(params$year1,"-",params$year2)` {#section-58}
  
```{r plot58, fig.cap= paste('Postpartum hemorrhage by year, Nova Scotia,', paste0(params$year1,"-",params$year2))}

dta1 <- dta[, .(BrthYear,
     R003_00400, R003_00900, R003_01300, R003_01400, R003_02200,
     R027_00400, 
     R029_00100, R029_00200, R029_00300, R029_00400, 
     MO72,
     M_1KT51, M_1RM13GQC2,
     M_1RM13GQW0, M_5PC91HR, M_5PC91HT,
     M_5PC91HU, M_5PC91HV, M_5PC91LA
)] %>%
  .[, `:=` (postph = fcase(
    R003_00400 >=1 | R003_00900 >=1 | R003_01300 >=1 | R003_01400 >=1 | R003_02200 >=1 |
     R027_00400 >=1 | 
     R029_00100 >=1 | R029_00200 >=1 | R029_00300 >=1 | R029_00400 >=1 | 
     MO72 >=1 |
     M_1KT51 >=1 | M_1RM13GQC2 >=1 |
     M_1RM13GQW0 >=1 | M_5PC91HR >=1 | M_5PC91HT >=1 |
     M_5PC91HU >=1 | M_5PC91HV >=1 | M_5PC91LA >=1, 1,
    default = 0
  ))] %>%
  #unique() %>%
  .[, `:=` (count = .N), by = list(BrthYear, postph)] %>%
  .[, `:=` (total = .N), by = list(BrthYear)] %>%
  .[, `:=` (rate = count/total)] %>%
  unique()


dta1 <- dta1[, `:=` (postph = factor(postph,
                                   levels = c(0,1),
                                   labels = c("No",
                                              "Postpartum hemorrhage")))]

## getting color for plotting

pal <- c("No" = "#1b9e77",
         "Postpartum hemorrhage" = "#d95f02")

plot <- ggplot(data = dta1[!tolower(postph) %in% "no"],
               aes(x = BrthYear,
                   y = rate,
                   group = postph,
                   colour = postph,
                   fill = postph,
                   #shape = postph,
                   text =paste(postph,
                               "<br>Year:", BrthYear,
                               "<br># births:", scales::comma(total, accuracy = 1),
                               "<br>% of deliveries:" , scales::percent(rate, accuracy = .1)
                   )
               )
) +
  geom_point(aes(color = postph, fill = postph)) +
  geom_line(aes(color = postph, fill = postph)) +
  # Specifying colors by age group
  scale_color_manual(name = "",
                     values = pal,
                     limits = names(pal)) +
  scale_fill_manual(name = "",
                    values = pal,
                    limits = names(pal)) +
  # scale_shape_manual(name = "",
  #                    values = c(15, 16)) +
  # scale_x_date(date_breaks = "2 years",
  #              date_labels = "%Y") +
  scale_x_continuous(#trans = 'log2',
    breaks = pretty) +
  scale_y_continuous(#trans = 'log2',
    breaks = pretty,
    labels = scales::percent_format(accuracy = 1),
    limits = c(0, NA)) +
  labs(y = '% of deliveries', x = '') +
  theme_minimal()

plotly::ggplotly(plot, tooltip = c("text")) %>%
  plotly::style(hoverlabel = list(
    bgcolor  = "black",
    bordercolor = "transparent",
    font = list(
      color = "white",
      size = 14,
      face = "bold"
    )
  )) %>%
  plotly::layout(legend = list(orientation = "h", y = -0.25, x = 0.5,
                               xanchor = "center"),
                 font = list(face = "bold", size = 10),
                 showlegend = FALSE)
```

```{r tbl58}
tbl <- dta[, .(BrthYear,
     R003_00400, R003_00900, R003_01300, R003_01400, R003_02200,
     R027_00400, 
     R029_00100, R029_00200, R029_00300, R029_00400, 
     MO72,
     M_1KT51, M_1RM13GQC2,
     M_1RM13GQW0, M_5PC91HR, M_5PC91HT,
     M_5PC91HU, M_5PC91HV, M_5PC91LA
)] %>%
  .[, `:=` (postph = fcase(
    R003_00400 >=1 | R003_00900 >=1 | R003_01300 >=1 | R003_01400 >=1 | R003_02200 >=1 |
     R027_00400 >=1 | 
     R029_00100 >=1 | R029_00200 >=1 | R029_00300 >=1 | R029_00400 >=1 | 
     MO72 >=1 |
     M_1KT51 >=1 | M_1RM13GQC2 >=1 |
     M_1RM13GQW0 >=1 | M_5PC91HR >=1 | M_5PC91HT >=1 |
     M_5PC91HU >=1 | M_5PC91HV >=1 | M_5PC91LA >=1, 1,
    default = 0
  ))] %>%
  #unique() %>%
  .[, `:=` (count = .N), by = list(BrthYear, postph)] %>%
  .[, `:=` (total = .N), by = list(BrthYear)] %>%
  .[, `:=` (rate = 100*count/total)] %>%
  unique() %>%
  melt(measure.vars = c("count", "total", "rate")) %>%
  .[!tolower(variable) %in% "count" &
    !postph %in% 0] %>%
  .[, `:=` (postph = fifelse(tolower(variable) %in% "total",
                          0, postph))] %>%
  .[, .(BrthYear, postph, value)] %>%
  unique() %>%
  dcast(postph ~ BrthYear, value.var = "value")

coln <- names(tbl)
coln[1] <- ""

tbl <- as.data.frame(sapply(tbl, sprintf, fmt = c("%.0f", "%1.1f%%")))

names(tbl) <- coln

tbl[,1] <- c("#deliveries", "Postpartum hemorrhage")

knitr::kable(tbl,
             format = "html",
             booktabs = TRUE) %>%
  # add an hline before the total
  #kableExtra::row_spec(nrow(tbl) - 1, hline_after = TRUE) %>%  #kableExtra::row_spec(nrow(tbl), bold = TRUE) %>%
kableExtra::row_spec(0, bold = TRUE) %>%
#kableExtra::column_spec(1, border_right = TRUE, bold  = TRUE) %>%
kableExtra::kable_styling(font_size = 12, full_width = FALSE,
                            bootstrap_options = c("hover","condensed", "responsive")) %>%
  kableExtra::add_footnote(label = c("Note: Postpartum hemorrhage is diagnosed if, after the delivery of the fetus, excessive maternal bleeding occurs from the genital tract with an estimated blood loss of greater than 500 mL for vaginal deliveries or 1000 mL for Caesarean section deliveries."))
```

## Maternal blood transfusion by year, Nova Scotia, `r paste0(params$year1,"-",params$year2)` {#section-59}
  
```{r plot59, fig.cap= paste('Maternal blood transfusion by year, Nova Scotia,', paste0(params$year1,"-",params$year2))}

dta1 <- dta[, .(BrthYear,
     RBC
)] %>%
  .[, `:=` (transf = fcase(
    RBC >=1 , 1,
    default = 0
  ))] %>%
  #unique() %>%
  .[, `:=` (count = .N), by = list(BrthYear, transf)] %>%
  .[, `:=` (total = .N), by = list(BrthYear)] %>%
  .[, `:=` (rate = count/total)] %>%
  unique()


dta1 <- dta1[, `:=` (transf = factor(transf,
                                   levels = c(0,1),
                                   labels = c("No",
                                              "Transfusion")))]

## getting color for plotting

pal <- c("No" = "#1b9e77",
         "Transfusion" = "#d95f02")

plot <- ggplot(data = dta1[!tolower(transf) %in% "no"],
               aes(x = BrthYear,
                   y = rate,
                   group = transf,
                   colour = transf,
                   fill = transf,
                   #shape = transf,
                   text =paste(transf,
                               "<br>Year:", BrthYear,
                               "<br># births:", scales::comma(total, accuracy = 1),
                               "<br>% of deliveries:" , scales::percent(rate, accuracy = .1)
                   )
               )
) +
  geom_point(aes(color = transf, fill = transf)) +
  geom_line(aes(color = transf, fill = transf)) +
  # Specifying colors by age group
  scale_color_manual(name = "",
                     values = pal,
                     limits = names(pal)) +
  scale_fill_manual(name = "",
                    values = pal,
                    limits = names(pal)) +
  # scale_shape_manual(name = "",
  #                    values = c(15, 16)) +
  # scale_x_date(date_breaks = "2 years",
  #              date_labels = "%Y") +
  scale_x_continuous(#trans = 'log2',
    breaks = pretty) +
  scale_y_continuous(#trans = 'log2',
    breaks = pretty,
    labels = scales::percent_format(accuracy = .1),
    limits = c(0, NA)) +
  labs(y = '% of deliveries', x = '') +
  theme_minimal()

plotly::ggplotly(plot, tooltip = c("text")) %>%
  plotly::style(hoverlabel = list(
    bgcolor  = "black",
    bordercolor = "transparent",
    font = list(
      color = "white",
      size = 14,
      face = "bold"
    )
  )) %>%
  plotly::layout(legend = list(orientation = "h", y = -0.25, x = 0.5,
                               xanchor = "center"),
                 font = list(face = "bold", size = 10),
                 showlegend = FALSE)
```

```{r tbl59}
tbl <- dta[, .(BrthYear,
     RBC
)] %>%
  .[, `:=` (transf = fcase(
    RBC >=1 , 1,
    default = 0
  ))] %>%
  #unique() %>%
  .[, `:=` (count = .N), by = list(BrthYear, transf)] %>%
  .[, `:=` (total = .N), by = list(BrthYear)] %>%
  .[, `:=` (rate = 100*count/total)] %>%
  unique() %>%
  melt(measure.vars = c("count", "total", "rate")) %>%
  .[!tolower(variable) %in% "count" &
    !transf %in% 0] %>%
  .[, `:=` (transf = fifelse(tolower(variable) %in% "total",
                          0, transf))] %>%
  .[, .(BrthYear, transf, value)] %>%
  unique() %>%
  dcast(transf ~ BrthYear, value.var = "value")

coln <- names(tbl)
coln[1] <- ""

tbl <- as.data.frame(sapply(tbl, sprintf, fmt = c("%.0f", "%1.1f%%")))

names(tbl) <- coln

tbl[,1] <- c("# deliveries", "Transfusion")

knitr::kable(tbl,
             format = "html",
             booktabs = TRUE) %>%
  # add an hline before the total
  #kableExtra::row_spec(nrow(tbl) - 1, hline_after = TRUE) %>%  #kableExtra::row_spec(nrow(tbl), bold = TRUE) %>%
kableExtra::row_spec(0, bold = TRUE) %>%
#kableExtra::column_spec(1, border_right = TRUE, bold  = TRUE) %>%
kableExtra::kable_styling(font_size = 12, full_width = FALSE,
                            bootstrap_options = c("hover","condensed", "responsive")) %>%
  kableExtra::add_footnote(label = c("Note: One or more maternal transfusions of red blood cells in the antepartum, intrapartum, or postpartum periods."))
```

## Maternal antepartum hospital length of stay (in hours) by year, Nova Scotia, `r paste0(params$year1,"-",params$year2)` {#section-510}
  
```{r plot510, fig.cap= paste('Maternal antepartum hospital length of stay (in hours) by year, Nova Scotia,', paste0(params$year1,"-",params$year2))}

dta1 <- dta[, .(BrthYear,
     APLoS
)] %>%
  .[!is.na(APLoS)] %>%
  .[, `:=` (total = .N,
            mean = mean(APLoS, na.rm = TRUE),
            med = median(APLoS, na.rm = TRUE)), by = list(BrthYear)] %>%
  unique()

plot <- ggplot(data = dta1,
               aes(x = BrthYear,
                   y = med,
                   group = 1,
                   #shape = BrthYear,
                   text =paste(BrthYear,
                               "<br># births:", scales::comma(total, accuracy = 1),
                               "<br>% Median lenght of stay (in hours):" , scales::comma(med, accuracy = .1)
                   )
               )
) +
  geom_point(aes(color = "#d95f02", fill = "#d95f02")) +
  geom_line(aes(color = "#d95f02", fill = "#d95f02")) +
  # Specifying colors by age group
  scale_color_manual(name = "",
                     values = "#d95f02") +
  scale_fill_manual(name = "",
                    values = "#d95f02") +
  # scale_shape_manual(name = "",
  #                    values = c(15, 16)) +
  # scale_x_date(date_breaks = "2 years",
  #              date_labels = "%Y") +
  scale_x_continuous(#trans = 'log2',
    breaks = pretty) +
  scale_y_continuous(#trans = 'log2',
    breaks = pretty,
    labels = scales::comma_format(accuracy = 1),
    limits = c(0, NA)) +
  labs(y = 'Median antepartum LoS (in hours)', x = '') +
  theme_minimal()

plotly::ggplotly(plot, tooltip = c("text")) %>%
  plotly::style(hoverlabel = list(
    bgcolor  = "black",
    bordercolor = "transparent",
    font = list(
      color = "white",
      size = 14,
      face = "bold"
    )
  )) %>%
  plotly::layout(legend = list(orientation = "h", y = -0.25, x = 0.5,
                               xanchor = "center"),
                 font = list(face = "bold", size = 10),
                 showlegend = FALSE)
```

```{r tbl510}
tbl <- dta[, .(BrthYear,
     APLoS
)] %>%
  .[!is.na(APLoS)] %>%
  .[, `:=` (total = .N,
            mean = mean(APLoS, na.rm = TRUE),
            med = median(APLoS, na.rm = TRUE)), by = list(BrthYear)] %>%
  unique() %>%
  melt(measure.vars = c("total", "mean", "med")) %>%
  # .[!tolower(variable) %in% "count" &
  #   !transf %in% 0] %>%
  # .[, `:=` (transf = fifelse(tolower(variable) %in% "total",
  #                         0, transf))] %>%
  .[, .(BrthYear, variable, value)] %>%
  unique() %>%
  dcast(variable ~ BrthYear, value.var = "value")

coln <- names(tbl)
coln[1] <- ""

tbl <- as.data.frame(sapply(tbl, sprintf, fmt = c("%.0f", "%.1f", "%.1f")))

names(tbl) <- coln

tbl[,1] <- c("# deliveries", "Mean (in hours)", "Median (in hours)")

knitr::kable(tbl,
             format = "html",
             booktabs = TRUE) %>%
  # add an hline before the total
  #kableExtra::row_spec(nrow(tbl) - 1, hline_after = TRUE) %>%  #kableExtra::row_spec(nrow(tbl), bold = TRUE) %>%
kableExtra::row_spec(0, bold = TRUE) %>%
#kableExtra::column_spec(1, border_right = TRUE, bold  = TRUE) %>%
kableExtra::kable_styling(font_size = 12, full_width = FALSE,
                            bootstrap_options = c("hover","condensed", "responsive"))
```

## Maternal postpartum hospital length of stay (in hours) by type of delivery and year, Nova Scotia, `r paste0(params$year1,"-",params$year2)` {#section-511}
  
```{r plot511, fig.cap= paste('Maternal postpartum hospital length of stay (in hours) by type of delivery and year, Nova Scotia,', paste0(params$year1,"-",params$year2))}

dta1 <- dta[, .(BrthYear,
     PPLoS,
     DMMETHOD
)] %>%
  .[!is.na(PPLoS)] %>%
  .[, `:=` (deliv = fcase(
      # spontaneous vaginal
      tolower(DMMETHOD) %in% c("biv", "spt"), 1,
      # assisted vaginal
      tolower(DMMETHOD) %in% c("vac", "low", "mid"), 2,
      # c-section
      tolower(DMMETHOD) %in% c("cst"), 3,
      default = 4
    ))] %>%
  .[!deliv %in% 4] %>%
  .[, `:=` (total = .N,
            mean = mean(PPLoS, na.rm = TRUE),
            med = median(PPLoS, na.rm = TRUE)), by = list(BrthYear, deliv)] %>%
  unique()

dta1 <- dta1[, `:=` (deliv = factor(deliv,
                                   levels = c(1,2,3),
                                   labels = c("Spontaneous vaginal",
                                              "Assisted vaginal",
                                              "Caesarean section")))]

## getting color for plotting

pal <- c("Spontaneous vaginal" = "#1b9e77",
         "Assisted vaginal" = "#d95f02",
         "Caesarean section" = "#7570b3")

plot <- ggplot(data = dta1,
               aes(x = BrthYear,
                   y = med,
                   group = deliv,
                   colour = deliv,
                   fill = deliv,
                   shape = deliv,
                   text =paste(deliv,
                               "<br>Year:",BrthYear,
                               "<br># births:", scales::comma(total, accuracy = 1),
                               "<br>% Median lenght of stay (in hours):" , scales::percent(med, accuracy = .1)
                   )
               )
) +
  geom_point(aes(color = deliv, fill = deliv, shape = deliv)) +
  geom_line(aes(color = deliv, fill = deliv)) +
  # Specifying colors by age group
  scale_color_manual(name = "",
                     values = pal,
                     limits = names(pal)) +
  scale_fill_manual(name = "",
                    values = pal,
                     limits = names(pal)) +
  scale_shape_manual(name = "",
                     values = c(15, 16, 17)) +
  # scale_x_date(date_breaks = "2 years",
  #              date_labels = "%Y") +
  scale_x_continuous(#trans = 'log2',
    breaks = pretty) +
  scale_y_continuous(#trans = 'log2',
    breaks = pretty,
    labels = scales::comma_format(accuracy = 1),
    limits = c(0, NA)) +
  labs(y = 'Median postpartum LoS (in hours)', x = '') +
  theme_minimal()

plotly::ggplotly(plot, tooltip = c("text")) %>%
  plotly::style(hoverlabel = list(
    bgcolor  = "black",
    bordercolor = "transparent",
    font = list(
      color = "white",
      size = 14,
      face = "bold"
    )
  )) %>%
  plotly::layout(legend = list(orientation = "h", y = -0.25, x = 0.5,
                               xanchor = "center"),
                 font = list(face = "bold", size = 10),
                 showlegend = TRUE)
```

```{r tbl511}
tbl <- dta[, .(BrthYear,
     PPLoS,
     DMMETHOD
)] %>%
  .[!is.na(PPLoS)] %>%
  .[, `:=` (deliv = fcase(
      # spontaneous vaginal
      tolower(DMMETHOD) %in% c("biv", "spt"), 1,
      # assisted vaginal
      tolower(DMMETHOD) %in% c("vac", "low", "mid"), 2,
      # c-section
      tolower(DMMETHOD) %in% c("cst"), 3,
      default = 4
    ))] %>%
  .[!deliv %in% 4] %>%
  .[, `:=` (total = .N,
            mean = mean(PPLoS, na.rm = TRUE),
            med = median(PPLoS, na.rm = TRUE)), by = list(BrthYear, deliv)] %>%
  unique()

tbl1 <- tbl %>%
  .[deliv %in% 1] %>%
  melt(measure.vars = c("total", "mean", "med")) %>%
  .[, .(BrthYear, variable, value)] %>%
  unique() %>%
  dcast(variable ~ BrthYear, value.var = "value")

coln <- names(tbl1)
coln[1] <- ""

tbl1 <- as.data.frame(sapply(tbl1, sprintf, fmt = c("%.0f", "%.1f", "%.1f")))

names(tbl1) <- coln
tbl1[,1] <- c("# deliveries", "Mean (in hours)", "Median (in hours)")

tbl2 <- tbl %>%
  .[deliv %in% 2] %>%
  melt(measure.vars = c("total", "mean", "med")) %>%
  .[, .(BrthYear, variable, value)] %>%
  unique() %>%
  dcast(variable ~ BrthYear, value.var = "value")

coln <- names(tbl2)
coln[1] <- ""

tbl2 <- as.data.frame(sapply(tbl2, sprintf, fmt = c("%.0f", "%.1f", "%.1f")))

names(tbl2) <- coln
tbl2[,1] <- c("# deliveries", "Mean (in hours)", "Median (in hours)")

tbl3 <- tbl %>%
  .[deliv %in% 3] %>%
  melt(measure.vars = c("total", "mean", "med")) %>%
  .[, .(BrthYear, variable, value)] %>%
  unique() %>%
  dcast(variable ~ BrthYear, value.var = "value")

coln <- names(tbl3)
coln[1] <- ""

tbl3 <- as.data.frame(sapply(tbl3, sprintf, fmt = c("%.0f", "%.1f", "%.1f")))

names(tbl3) <- coln
tbl3[,1] <- c("# deliveries", "Mean (in hours)", "Median (in hours)")

tbl <- setDT(rbind(tbl1,tbl2,tbl3))

names(tbl)[1] <- "Type of delivery"



knitr::kable(tbl,
             format = "html",
             booktabs = TRUE) %>%
  # add an hline before the total
  #kableExtra::row_spec(nrow(tbl) - 1, hline_after = TRUE) %>%  #kableExtra::row_spec(nrow(tbl), bold = TRUE) %>%
kableExtra::row_spec(0, bold = TRUE) %>%
#kableExtra::column_spec(1, border_right = TRUE, bold  = TRUE) %>%
kableExtra::group_rows("Spontaneous vaginal", start_row = 1, end_row = 3) %>%
kableExtra::group_rows("Assited vaginal", start_row = 4, end_row = 6) %>%
kableExtra::group_rows("Caesarean section", start_row = 7, end_row = 9) %>%
kableExtra::kable_styling(font_size = 12, full_width = FALSE,
                            bootstrap_options = c("hover","condensed", "responsive"))
```